---
title: Fifth Tutorial for R4DS
author: David Kane
tutorial:
  id: r4ds-5
output:
  learnr::tutorial:
    df_print: default
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: 'Working with hierarchical JSON data, web scraping with rvest and httr2, and custom functions, using earthquake and IMDb movie datasets.'
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)

library(tidyverse)
library(leaflet)
library(rvest)
library(ggrepel)
library(httr2)

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(out.width = '90%')
options(tutorial.exercise.timelimit = 60,
        tutorial.storage = "local")

convert_earthquake_time <- function(milliseconds) {
  as.POSIXct(milliseconds / 1000, origin = "1970-01-01", tz = "UTC")
}
categorize_magnitude <- function(mag) {
  case_when(
    mag < 3.0 ~ "Minor",
    mag < 4.0 ~ "Light",
    mag < 5.0 ~ "Moderate",
    mag < 6.0 ~ "Strong",
    mag < 7.0 ~ "Major",
    TRUE ~ "Great"
  )
}

earthquake_df <- jsonlite::read_json("data/earthquakes.geojson") |>
  pluck("features") |>
  tibble(data = _) |>
  unnest_wider(data) |>
  unnest_wider(properties, names_sep = "_") |>
  hoist(geometry, coordinates = "coordinates") |>
  unnest_wider(coordinates, names_sep = "_") |>
  rename(
    longitude = coordinates_1,
    latitude  = coordinates_2,
    depth     = coordinates_3
  ) |>
  mutate(across(where(is.character), ~ifelse(.x == "null", NA, .x))) |>
  filter(!if_any(c(properties_mag, longitude, latitude), is.na)) |>
  mutate(
    datetime = convert_earthquake_time(as.numeric(properties_time)),
    mag_category = categorize_magnitude(as.numeric(properties_mag))
  ) |>
  mutate(across(c(properties_mag, depth), as.numeric)) |>
  filter(datetime >= Sys.Date() - 365) |>
  filter(longitude >= -125, longitude <= -66, latitude >= 20, latitude <= 50) |>
  mutate(
    hour = hour(datetime),
    day_of_week = wday(datetime, label = TRUE),
    depth_category = case_when(
      depth < 70 ~ "Shallow",
      depth < 300 ~ "Intermediate",
      depth < 600 ~ "Deep"
    )
  ) |>
  select(longitude, latitude, properties_mag, properties_place, datetime, depth, mag_category, depth_category) |>
  mutate(
    tooltip_text = paste0(
      "Magnitude: ", properties_mag, " (", mag_category, ")<br>",
      "Location: ", properties_place, "<br>",
      "Time: ", format(datetime, "%Y-%m-%d %H:%M"), "<br>",
      "Depth: ", round(depth, 1), " km (", depth_category, ")"
    )
  ) |>
  filter(
    longitude >= -180, longitude <= 180,
    latitude >= -90, latitude <= 90,
    properties_mag >= 0, properties_mag <= 10
  ) |>
  mutate(
    color = case_when(
      properties_mag >= 6 ~ "red",
      properties_mag >= 5 ~ "orange",
      properties_mag >= 4 ~ "yellow",
      properties_mag >= 3 ~ "green",
      TRUE ~ "blue"
    )
  ) |>
  arrange(desc(properties_mag))

top_movies <- request("https://web.archive.org/web/20220201012049/https://www.imdb.com/chart/top/") |>
  req_perform() |>
  resp_body_html() |>
  html_element("table") |>
  html_table() |>
  select(
    rank_title_year = `Rank & Title`,
    rating = `IMDb Rating`
  ) |>
  mutate(
    rank_title_year = str_replace_all(rank_title_year, "\n +", " "),
    rating_details = request("https://web.archive.org/web/20220201012049/https://www.imdb.com/chart/top/") |>
      req_perform() |>
      resp_body_html() |>
      html_elements("td strong") |>
      html_attr("title")
  ) |>
  separate_wider_regex(
    rank_title_year,
    patterns = c(
      rank = "\\d+", "\\. ",
      title = ".+", " +\\(",
      year = "\\d+", "\\)"
    )
  ) |>
  separate_wider_regex(
    rating_details,
    patterns = c(
      "[0-9.]+ based on ",
      number = "[0-9,]+",
      " user ratings"
    )
  ) |>
  mutate(
    number = parse_number(number),
    across(c(rank, year), as.numeric),
    decade = paste0(floor(year/10)*10, "s"),
    rating_category = case_when(
      rating >= 9.0 ~ "Masterpiece (9.0+)",
      rating >= 8.5 ~ "Excellent (8.5-8.9)",
      rating >= 8.0 ~ "Great (8.0-8.4)",
      TRUE ~ "Good (<8.0)"
    ),
    popularity = case_when(
      number >= 2000000 ~ "Very Popular (2M+)",
      number >= 1000000 ~ "Popular (1M-2M)",
      number >= 500000 ~ "Moderate (500K-1M)",
      TRUE ~ "Niche (<500K)"
    )
  ) |>
  select(rank, title, year, rating, number) |>
  arrange(rank)
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

## Introduction
###

This section covers key concepts from [Chapter 23: Hierarchical Data](https://r4ds.hadley.nz/hierarchical), [Chapter 24: Web Scraping](https://r4ds.hadley.nz/webscraping), [Chapter 25: Functions](https://r4ds.hadley.nz/functions), and [Chapter 26: Iterations](https://r4ds.hadley.nz/iteration) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund.

You will learn about working with nested data structures, web scraping techniques, and creating custom functions using core packages such as **[jsonlite](https://cran.r-project.org/package=jsonlite)**, **[rvest](https://rvest.tidyverse.org/)**, **[httr2](https://httr2.r-lib.org/)**, and **[purrr](https://purrr.tidyverse.org/)**. Key concepts include extracting nested elements from lists, expanding list-columns into separate columns, extracting tables from HTML pages, and splitting text into structured columns using regex patterns.

We recommend using an agentic coding tool such as [Gemini CLI](https://github.com/google-gemini/gemini-cli) or [Claude Code](https://claude.ai/code). Our instructions are written with these tools in mind. You may also use a chat-based AI, but you will need to copy/paste code and data context manually.

### Exercise 1

You should be connected to a repo named `r4ds-5`. If you are not, create one and connect to it.

Create a new file, `analysis.qmd`, with the title `"Earthquakes and Movies"` and your name as the author. In a bash Terminal, render it:

```
quarto render analysis.qmd
```

Open `analysis.html` with Live Server (right-click it in the Explorer → **Open with Live Server**) and keep the tab open. It refreshes on every render.

Create a `.gitignore` with `analysis_files` on its own line. Commit and push.

In the R Terminal, run:

```
show_file(".gitignore")
```

If that fails, it is probably because you have not yet loaded `library(tutorial.helpers)` in the R Terminal.

CP/CR.

```{r introduction-1}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

###

```
analysis_files
```

###

List-columns come in two basic forms, named and unnamed. A **named** list labels each element, like a record:

```r
list(place = "Loma Prieta", year = 1989, magnitude = 6.9)
```

An **unnamed** list is just an ordered sequence — `list(6.9, 7.1, 5.4)` — whose length can vary from row to row.

### Exercise 2

In your QMD, paste the following `library()` commands in a new code chunk with `#| message: false`:

```
library(tidyverse)
library(leaflet)
library(rvest)
library(httr2)
library(ggrepel)
```

Also add the following to the YAML header to remove all code from the HTML output:

```
execute:
  echo: false
```

In a bash Terminal, run `quarto render analysis.qmd`. Check your Live Server tab.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r introduction-2}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 6)
```

<pre><code>```{r}
#| message: false
library(tidyverse)
library(leaflet)
library(rvest)
library(httr2)
library(ggrepel)
```</code></pre>

###

JSON is a machine-readable format with six data types: null (like R's `NA`), string, number, boolean (lowercase `true`/`false`), array (like unnamed R lists), and object (like named R lists).

```json
{
  "place": "Northern California",
  "magnitude": 4.2,
  "felt": true,
  "tags": ["minor", "shallow"]
}
```

### Exercise 3

In the R Terminal, run:

```
dir.create("data")
list.files()
```

This creates a `data` directory in your project for storing downloaded files.

CP/CR.

```{r introduction-3}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 5)
```

###

```
[1] "analysis.html"  "analysis.qmd"   "analysis_files" "data"
```

###

**jsonlite** can also read JSON with automatic simplification, which often works well for simple, regular JSON. Reading it *without* simplification keeps the full nested structure, giving you more control when you need to rectangle something complex by hand.

## Earthquakes
###

This section focuses on working with hierarchical JSON data and writing your own transformation functions, using **[jsonlite](https://cran.r-project.org/package=jsonlite)**, **[tidyr](https://tidyr.tidyverse.org/)**, and **[leaflet](https://rstudio.github.io/leaflet/)**. You will learn how to extract nested data from GeoJSON files, convert Unix timestamps to datetimes, and build interactive maps. Key concepts include reaching into deeply nested lists, *rectangling* them into a tidy table, and writing small custom functions.

We use `earthquakes.geojson` from the **USGS real-time earthquake feed**, a live GeoJSON feed maintained by the U.S. Geological Survey providing seismic activity data including magnitude, location, depth, and timestamps.

Here is the tibble we will build:

```{r earthquakes-data}
earthquake_df
```

### Exercise 1

Download the GeoJSON file to your `data` directory. In the R Terminal, run:

```
download.file(
  "https://github.com/PPBDS/misc.tutorials/raw/refs/heads/main/inst/tutorials/r4ds-5/data/earthquakes.geojson",
  destfile = "data/earthquakes.geojson",
  mode = "wb"
)
```

CP/CR.

```{r earthquakes-1}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 5)
```

###

GeoJSON is a JSON-based format for representing geographic features. A `FeatureCollection` object contains a `features` array — each feature has `type`, `geometry` (with coordinates), and `properties` (metadata like magnitude and place).

### Exercise 2

In a new code chunk in your `analysis.qmd`, paste the following code to explore the raw JSON structure:

```
jsonlite::read_json("data/earthquakes.geojson") |>
  pluck("features") |>
  tibble(data = _) |>
  unnest_wider(data) |>
  unnest_wider(properties, names_sep = "_") |>
  select(type, id, properties_mag, properties_place, properties_time) |>
  head(5)
```

In a bash Terminal, run `quarto render analysis.qmd`. Check your Live Server tab.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r earthquakes-2}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 6)
```

###

```{r earthquakes-2-test}
jsonlite::read_json("data/earthquakes.geojson") |>
  pluck("features") |>
  tibble(data = _) |>
  unnest_wider(data) |>
  unnest_wider(properties, names_sep = "_") |>
  select(type, id, properties_mag, properties_place, properties_time) |>
  head(5)
```

###

Pulling a named element out of a deeply nested list, then spreading a named list-column so each element becomes its own column, are the core *rectangling* moves — they turn hierarchical JSON into a flat table. When you unnest several levels, prefixing the new column names with their parent keeps them from colliding.

### Exercise 3

In your `analysis.qmd`, replace the code from the previous exercise with a new code chunk. Turn the raw GeoJSON into a tidy table called `earthquake_df` with one row per earthquake. From each quake's nested record, pull out longitude, latitude, depth, and magnitude, and turn the timestamp into a real datetime. Keep only quakes from the last year and within the continental US (longitude −125 to −66, latitude 20 to 50). Add a magnitude category (Minor, Light, Moderate, Strong, Major, Great), a depth category (Shallow, Intermediate, Deep), the hour and day of week, a color keyed to magnitude, and a `tooltip_text` string for the map popups. Sort by descending magnitude.

In a bash Terminal, run `quarto render analysis.qmd`. Check your Live Server tab.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r earthquakes-3}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 8)
```

###

```{r earthquakes-3-test}
earthquake_df
```

###

If your result does not match ours, replace your code with the following:

<pre><code>```{r}
convert_earthquake_time <- function(milliseconds) {
  as.POSIXct(milliseconds / 1000, origin = "1970-01-01", tz = "UTC")
}
categorize_magnitude <- function(mag) {
  case_when(
    mag < 3.0 ~ "Minor",
    mag < 4.0 ~ "Light",
    mag < 5.0 ~ "Moderate",
    mag < 6.0 ~ "Strong",
    mag < 7.0 ~ "Major",
    TRUE ~ "Great"
  )
}
```</code></pre>

<pre><code>```{r}
#| cache: true
earthquake_df <- jsonlite::read_json("data/earthquakes.geojson") |>
  pluck("features") |>
  tibble(data = _) |>
  unnest_wider(data) |>
  unnest_wider(properties, names_sep = "_") |>
  hoist(geometry, coordinates = "coordinates") |>
  unnest_wider(coordinates, names_sep = "_") |>
  rename(
    longitude = coordinates_1,
    latitude  = coordinates_2,
    depth     = coordinates_3
  ) |>
  mutate(across(where(is.character), ~ifelse(.x == "null", NA, .x))) |>
  filter(!if_any(c(properties_mag, longitude, latitude), is.na)) |>
  mutate(
    datetime = convert_earthquake_time(as.numeric(properties_time)),
    mag_category = categorize_magnitude(as.numeric(properties_mag))
  ) |>
  mutate(across(c(properties_mag, depth), as.numeric)) |>
  filter(datetime >= Sys.Date() - 365) |>
  filter(longitude >= -125, longitude <= -66, latitude >= 20, latitude <= 50) |>
  mutate(
    hour = hour(datetime),
    day_of_week = wday(datetime, label = TRUE),
    depth_category = case_when(
      depth < 70 ~ "Shallow",
      depth < 300 ~ "Intermediate",
      depth < 600 ~ "Deep"
    )
  ) |>
  select(longitude, latitude, properties_mag, properties_place,
         datetime, depth, mag_category, depth_category) |>
  mutate(
    tooltip_text = paste0(
      "Magnitude: ", properties_mag, " (", mag_category, ")<br>",
      "Location: ", properties_place, "<br>",
      "Time: ", format(datetime, "%Y-%m-%d %H:%M"), "<br>",
      "Depth: ", round(depth, 1), " km (", depth_category, ")"
    )
  ) |>
  filter(
    longitude >= -180, longitude <= 180,
    latitude >= -90, latitude <= 90,
    properties_mag >= 0, properties_mag <= 10
  ) |>
  mutate(
    color = case_when(
      properties_mag >= 6 ~ "red",
      properties_mag >= 5 ~ "orange",
      properties_mag >= 4 ~ "yellow",
      properties_mag >= 3 ~ "green",
      TRUE ~ "blue"
    )
  ) |>
  arrange(desc(properties_mag))
```</code></pre>

You can reach into a list-column and lift out just the named values you need — here, the coordinates inside the `geometry` object — without fully unnesting everything around them. That is handy when a nested structure carries far more fields than you care about.

### Exercise 4

Within the `earthquake_df` code chunk, add `#| cache: true` at the top. Add `analysis_cache` to your `.gitignore` if you have not already done so, since cache files should never go to GitHub.

In a bash Terminal, run `quarto render analysis.qmd`. This first render will take some time — Quarto runs the code and saves the result to disk — but every later render loads `earthquake_df` from the cache instead of re-rectangling the GeoJSON. Check your Live Server tab.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r earthquakes-4}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 8)
```

###

Filtering to the last 365 days and to the continental US trims the global feed down to a few thousand quakes — small enough to map smoothly. The vast majority are minor; the few above magnitude 5 are the ones a reader's eye should land on.

### Exercise 5

In a new code chunk, create an interactive **leaflet** map of `earthquake_df`. It should display circle markers sized by magnitude and colored by the `color` column, with popups showing the `tooltip_text`, plus a legend. Do not assign the plot to a variable.

In a bash Terminal, run `quarto render analysis.qmd`. Check your Live Server tab.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r earthquakes-5}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 12)
```

###

```{r earthquakes-5-test}
#| echo: true
leaflet(earthquake_df) |>
  addTiles() |>
  addCircleMarkers(
    lng = ~longitude,
    lat = ~latitude,
    radius = ~properties_mag * 2,
    color = ~color,
    popup = ~tooltip_text,
    fillOpacity = 0.7,
    stroke = TRUE,
    weight = 1
  ) |>
  addLegend(
    position = "bottomright",
    colors = c("red", "orange", "yellow", "green", "blue"),
    labels = c("Major (6.0+)", "Strong (5.0-5.9)", "Moderate (4.0-4.9)",
              "Light (3.0-3.9)", "Minor (<3.0)"),
    title = "Earthquake Magnitude"
  )
```

###

**[leaflet](https://rstudio.github.io/leaflet/)** builds interactive HTML maps: you start from a base layer of background map tiles, then add a layer of circle markers at each latitude/longitude. Marker popups accept HTML, which is why `tooltip_text` uses `<br>` for line breaks.

## Top Movies
###

This section covers web scraping and data extraction from HTML tables using **[rvest](https://rvest.tidyverse.org/)**, **[httr2](https://httr2.r-lib.org/)**, and **[stringr](https://stringr.tidyverse.org/)**. You will scrape the archived IMDb Top 250 movies list, clean the resulting text, and extract structured data. Key concepts include selecting elements from an HTML page, reading an HTML table into a data frame, pulling attributes off HTML nodes, and splitting one messy text column into several clean ones.

We use a archived version of **The IMDb Top 250 Movies list** from February 2022, accessed via the Wayback Machine for a stable, reproducible URL.

Here is the tibble we will build:

```{r top-movies-data}
top_movies
```

### Exercise 1

In a new code chunk in your `analysis.qmd`, paste the following code to fetch and display the raw IMDb table:

```
request("https://web.archive.org/web/20220201012049/https://www.imdb.com/chart/top/") |>
  req_perform() |>
  resp_body_html() |>
  html_element("table") |>
  html_table()
```

In a bash Terminal, run `quarto render analysis.qmd`. Check your Live Server tab.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r top-movies-1}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 6)
```

###

A `robots.txt` file tells automated bots which pages the site owner permits them to access. Always check `https://example.com/robots.txt` before scraping a new site. The Wayback Machine archives are generally permitted for research scraping.

### Exercise 2

In a new code chunk in your `analysis.qmd`, paste the following code to extract the rating details — the number of user votes — stored as an HTML attribute:

```
request("https://web.archive.org/web/20220201012049/https://www.imdb.com/chart/top/") |>
  req_perform() |>
  resp_body_html() |>
  html_elements("td strong") |>
  html_attr("title") |>
  head(5)
```

In a bash Terminal, run `quarto render analysis.qmd`. Check your Live Server tab.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r top-movies-2}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 6)
```

###

Selecting all the HTML nodes that match a CSS selector, then pulling a named attribute off each one, is how you reach data that isn't in the visible text — here the `title` attribute on the `<strong>` elements inside the table cells holds the vote-count string.

### Exercise 3

In your `analysis.qmd`, replace the code from the previous two exercises with a single new code chunk. Scrape the archived IMDb Top 250 list at `https://web.archive.org/web/20220201012049/https://www.imdb.com/chart/top/` into a tidy table called `top_movies`, with one row per film carrying its rank, title, year, rating, and number of votes, sorted by rank. (The rank, title, and year arrive mashed together in one cell, and the vote count is tucked into an HTML attribute rather than the visible text — so part of the job is prising those apart.)

In a bash Terminal, run `quarto render analysis.qmd`. Check your Live Server tab.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r top-movies-3}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 8)
```

###

```{r top-movies-3-test}
top_movies
```

###

If your result does not match ours, replace your code with the following:

<pre><code>```{r}
#| cache: true
top_movies <- request("https://web.archive.org/web/20220201012049/https://www.imdb.com/chart/top/") |>
  req_perform() |>
  resp_body_html() |>
  html_element("table") |>
  html_table() |>
  select(
    rank_title_year = `Rank & Title`,
    rating = `IMDb Rating`
  ) |>
  mutate(
    rank_title_year = str_replace_all(rank_title_year, "\n +", " "),
    rating_details = request("https://web.archive.org/web/20220201012049/https://www.imdb.com/chart/top/") |>
      req_perform() |>
      resp_body_html() |>
      html_elements("td strong") |>
      html_attr("title")
  ) |>
  separate_wider_regex(
    rank_title_year,
    patterns = c(
      rank = "\\d+", "\\. ",
      title = ".+", " +\\(",
      year = "\\d+", "\\)"
    )
  ) |>
  separate_wider_regex(
    rating_details,
    patterns = c(
      "[0-9.]+ based on ",
      number = "[0-9,]+",
      " user ratings"
    )
  ) |>
  mutate(
    number = parse_number(number),
    across(c(rank, year), as.numeric),
    decade = paste0(floor(year/10)*10, "s"),
    rating_category = case_when(
      rating >= 9.0 ~ "Masterpiece (9.0+)",
      rating >= 8.5 ~ "Excellent (8.5-8.9)",
      rating >= 8.0 ~ "Great (8.0-8.4)",
      TRUE ~ "Good (<8.0)"
    ),
    popularity = case_when(
      number >= 2000000 ~ "Very Popular (2M+)",
      number >= 1000000 ~ "Popular (1M-2M)",
      number >= 500000 ~ "Moderate (500K-1M)",
      TRUE ~ "Niche (<500K)"
    )
  ) |>
  select(rank, title, year, rating, number) |>
  arrange(rank)
```</code></pre>

One messy text column can be split into several clean ones with a regular expression made of *named* pieces: each named piece becomes a new column, while the unnamed pieces are matched and discarded. It is the workhorse for turning a string like "1. The Shawshank Redemption (1994)" into separate rank, title, and year columns.

### Exercise 4

Within the `top_movies` code chunk, add `#| cache: true` at the top.

In a bash Terminal, run `quarto render analysis.qmd`. This first render will take some time. Check your Live Server tab.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r top-movies-4}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 8)
```

###

### Exercise 5

In a new code chunk, create an interesting plot of `top_movies`. Give it a proper title, subtitle, and caption. Do not assign the plot to a variable.

In a bash Terminal, run `quarto render analysis.qmd`. Check your Live Server tab.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r top-movies-5}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 12)
```

###

```{r top-movies-5-test}
#| echo: true
top_movies |>
  filter(year >= 1990, year <= 2010) |>
  ggplot(aes(x = year, y = rating)) +
  geom_point(aes(size = number / 1000), alpha = 0.8, stroke = 0.5) +
  geom_smooth(method = "loess", se = TRUE, color = "navy",
              linewidth = 1.2, alpha = 0.3) +
  geom_text_repel(aes(label = ifelse(rating >= 9.0, str_wrap(title, 15), "")),
                  size = 3, max.overlaps = 8,
                  box.padding = 0.5, point.padding = 0.3) +
  scale_size_continuous(name = "Votes\n(thousands)", range = c(2, 12),
                        breaks = c(500, 1000, 1500, 2000, 2500),
                        labels = c("500", "1,000", "1,500", "2,000", "2,500")) +
  scale_x_continuous(breaks = seq(1990, 2010, 2)) +
  scale_y_continuous(breaks = seq(8.0, 9.2, 0.2), limits = c(7.9, 9.3)) +
  labs(
    title = "IMDb Top 250: The Golden Age (1990–2010)",
    subtitle = "Masterpiece films labeled · Size shows popularity",
    x = "Release Year",
    y = "IMDb Rating",
    caption = "Data: IMDb Top 250 archived February 2022"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 18, face = "bold", color = "navy"),
    plot.subtitle = element_text(size = 12, color = "gray40"),
    plot.caption = element_text(size = 9, color = "gray50"),
    panel.grid.minor = element_blank()
  )
```

###

**[ggrepel](https://ggrepel.slowkow.com/)** places text labels that automatically nudge apart so they don't overlap each other or the data points — the standard fix for a crowded scatter plot — with a control for how many overlapping labels to allow before some are hidden.

## Summary
###

This section covered key concepts from [Chapter 23: Hierarchical Data](https://r4ds.hadley.nz/hierarchical), [Chapter 24: Web Scraping](https://r4ds.hadley.nz/webscraping), [Chapter 25: Functions](https://r4ds.hadley.nz/functions), and [Chapter 26: Iterations](https://r4ds.hadley.nz/iteration) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund.

You learned about working with nested data structures, web scraping techniques, and creating custom functions using core packages such as **[jsonlite](https://cran.r-project.org/package=jsonlite)**, **[rvest](https://rvest.tidyverse.org/)**, **[httr2](https://httr2.r-lib.org/)**, and **[purrr](https://purrr.tidyverse.org/)**. Key concepts included extracting nested elements from lists, expanding list-columns, extracting HTML tables, and parsing complex text into structured columns using regex.

### Exercise 1

In a bash Terminal, run `quarto render analysis.qmd` to ensure that everything works. Check your Live Server tab.

In the R Terminal, run:

```
show_file("analysis.qmd")
```

CP/CR.

```{r summary-1}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 30)
```

###

**[purrr](https://purrr.tidyverse.org/)** lets you apply the same function to each element of a list and gather the results — either as a new list, or row-bound into a single data frame. That is the natural way to run one scraping function across many URLs.

### Exercise 2

Publish your rendered QMD to GitHub Pages. In a bash Terminal, run:

```
quarto publish gh-pages analysis.qmd
```

If the bash Terminal is still running from rendering, stop it first using `Ctrl/Cmd + C`.

Copy/paste the resulting URL below.

```{r summary-2}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 1)
```

###

**ggplot2** can save a plot to a file: by default it saves the most recently displayed plot, at the current device size, and infers the format (PNG, PDF, …) from the file extension.

### Exercise 3

Commit and push all your files. Copy/paste the URL to your GitHub repo.

```{r summary-3}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

###

There is one special type of missing value you may encounter: `NaN` (**n**ot **a** **n**umber), which arises from undefined operations like `0/0`. It behaves like `NA` in most respects, and on the rare occasion you must tell the two apart, there is a dedicated test for it.

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
