---
title: Second Tutorial for R4DS
author: Sruthi Gandhi and David Kane
tutorial:
  id: r4ds-2
output:
  learnr::tutorial:
    df_print: default
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: 'This tutorial for "R for Data Science" covers Chapter 9: Layers, Chapter 10: Exploratory Data Analysis, Chapter 11: Communication, and Chapter 20: Spreadsheets. Exercises practice ggplot2 layers, faceting, themes, and reading Excel files with readxl.'
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)

library(tidyverse)
library(readxl)

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(out.width = '90%')
options(tutorial.exercise.timelimit = 60,
        tutorial.storage = "local")


cheese_tibble <- read_excel("data/cheeses.xlsx") |>
  filter(!is.na(country)) |>
  add_count(country, name = "country_total") |>
  filter(country_total >= sort(unique(country_total), decreasing = TRUE)[3]) |>
  filter(!is.na(milk)) |>
  count(country, milk) |>
  group_by(country) |>
  arrange(country, desc(n))

wine_tibble <- read_excel("data/wine.xlsx")|>
  select(Phenols, Color, Flavanoids)
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

## Introduction
###

This section covers key concepts from [Chapter 9: Layers](https://r4ds.hadley.nz/layers.html), [Chapter 10: Exploratory Data Analysis](https://r4ds.hadley.nz/eda.html), [Chapter 11: Communication](https://r4ds.hadley.nz/communication.html), and [Chapter 20: Spreadsheets](https://r4ds.hadley.nz/spreadsheets.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund.

You will learn about working with categorical and continuous variables using different types of variation and exploratory data analysis using core packages such as **[readxl](https://readxl.tidyverse.org/)** and **[ggplot2](https://ggplot2.tidyverse.org/)**. Key **ggplot2** techniques include faceting plots into panels, flipping axes, and adding jitter to avoid overplotting.

We recommend using an agentic coding tool such as [Gemini CLI](https://github.com/google-gemini/gemini-cli) or [Claude Code](https://claude.ai/code), which can read and edit your files directly. Our instructions are written with these tools in mind. You may also use a chat-based AI, but you will need to copy/paste code and data context manually.


### Exercise 1

You should be in a repo named `r4ds-2`. If you are not, create one and move to it.

You need two R Terminals: one for running this tutorial and one for your exercises. Open a second R Terminal using the dropdown button.

Create a new file and save it as `analysis.qmd`. Add a YAML header at the top with a title (`"Analyzing Cheese and Wine"`) and your name as author. In a bash Terminal, run:

```
quarto render analysis.qmd
```

Open `analysis.html` in a browser by right-clicking it in the Explorer and selecting **Open with Live Server**. Keep this tab open — it will refresh automatically whenever you re-render.

Create a `.gitignore` file with `analysis_files` on the first line and then a blank line. Save and push.

In the R Terminal, run:

```
show_file(".gitignore")
```

If that fails, it is probably because you have not yet loaded `library(tutorial.helpers)` in the R Terminal.

CP/CR.

```{r introduction-1}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

###

The `.gitignore` file tells Git which files to skip. `analysis_files/` contains auto-generated files that Quarto creates during rendering — tracking them on GitHub wastes space and creates noisy diffs.

### Exercise 2

In your QMD, put `library(tidyverse)` and `library(readxl)` in a new code chunk. In a bash Terminal, run `quarto render analysis.qmd`. Check your Live Server tab — it should refresh automatically.

Notice that the file does not look good because the code is visible and there are annoying messages. To take care of this, add `#| message: false` to remove all the messages in this code chunk. Also add the following to the YAML header to remove all code echoes from the HTML:

```
execute:
  echo: false
```

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r introduction-2}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 6)
```

###

Adding `execute: echo: false` to the YAML header hides all code output by default. `#| message: false` on the setup chunk suppresses the package-loading messages that **tidyverse** prints. Together they give you a clean rendered page with only your results visible.

### Exercise 3

From the R Terminal, run these three commands:

`getwd()`. 
`dir.create("data")`. 
`list.files()`. 

This will create a `data` directory in your project. This is a good place to store any data that you are working with.

CP/CR.

```{r introduction-3}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 5)
```

###

Your answer should look something like this, although your path will be different.

````
> getwd()
[1] "/workspaces/r4ds-2"
> dir.create("data")
> list.files()
 [1] "analysis_files" "analysis.html"  "analysis.qmd"   "data"
>
````

## Cheese
###

The [cheese.csv](https://github.com/rfordatascience/tidytuesday/blob/main/data/2024/2024-06-04/readme.md) dataset is a [TidyTuesday](https://github.com/rfordatascience/tidytuesday) dataset on Github which contains data scraped from [cheese.com](https://www.cheese.com) about data from 2046 different types of cheese. Our spreadsheet `data/cheeses.xlsx` contains this data in a spreadsheet format.

### Exercise 1

We begin by downloading our spreadsheet `cheeses.xlsx` directly from GitHub into the `data` directory.

In the R Terminal, run:

```
download.file(
  "https://github.com/PPBDS/misc.tutorials/raw/refs/heads/main/inst/tutorials/r4ds-2/data/cheeses.xlsx",
  destfile = "data/cheeses.xlsx",
  mode = "wb"
)
```

CP/CR.

```{r cheese-1}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 8)
```

###

Downloading a file means copying it from a URL to a local path in your project. Spreadsheets like `.xlsx` are *binary* files, so they must be downloaded in binary mode — otherwise text-style line-ending conversion can quietly corrupt the file. The thing to carry away is the distinction between binary and text data, not the mechanics; your AI agent handles those.

### Exercise 2

Add a code chunk to your QMD that reads the cheese spreadsheet `data/cheeses.xlsx` and shows its structure — the column types with a few example values. In a bash Terminal, run `quarto render analysis.qmd`. Check your Live Server tab.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r cheese-2}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 5)
```

###

Our result:

```{r cheese-2-test}
glimpse(read_excel("data/cheeses.xlsx"))
```

The **[readxl](https://readxl.tidyverse.org/)** package reads both `.xls` and `.xlsx` spreadsheets, guessing the format from the file extension. A quick look at the column types and sample values helps you decide what to filter or transform before plotting.

### Exercise 3

Replace the code in the previous chunk with a pipeline that reads `data/cheeses.xlsx`, filters to the three most common cheese-producing countries, removes rows with missing `milk` values, and returns a count of milk types per country sorted by country and descending count. In a bash Terminal, run `quarto render analysis.qmd`.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r cheese-3}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 6)
```

###

Our result:

```{r cheese-3-test}
read_excel("data/cheeses.xlsx") |>
  filter(!is.na(country)) |>
  add_count(country, name = "country_total") |>
  filter(country_total >= sort(unique(country_total), decreasing = TRUE)[3]) |>
  filter(!is.na(milk)) |>
  count(country, milk) |>
  group_by(country) |>
  arrange(country, desc(n))
```

If your result doesn't match ours, replace your chunk code with the following before moving on:

<pre><code>```{r}
read_excel("data/cheeses.xlsx") |>
  filter(!is.na(country)) |>
  add_count(country, name = "country_total") |>
  filter(country_total >= sort(unique(country_total), decreasing = TRUE)[3]) |>
  filter(!is.na(milk)) |>
  count(country, milk) |>
  group_by(country) |>
  arrange(country, desc(n))
```</code></pre>

Sometimes you want a group count attached to every row without collapsing the data — for example, to keep each cheese's row while also knowing how many cheeses its country has, so you can filter to the busiest countries. Adding a per-group count as a new column does exactly that.

### Exercise 4

Within the recent code chunk, add the option `#| cache: true` and assign the result to `cheese_tibble`. If you have not done so already, add `analysis_cache` to your `.gitignore`. In a bash Terminal, run `quarto render analysis.qmd`. This first render takes noticeably longer — Quarto runs the code and writes the cache to disk; subsequent renders skip that code and load from cache instead.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r cheese-4}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 6)
```

###

Quarto's `#| cache: true` option saves a chunk's result to disk so subsequent renders load from cache instead of re-running. Counting combinations of values produces a tidy table with an `n` column of counts — and that count column can drive a plot's bar height, point size, or fill color in **[ggplot2](https://ggplot2.tidyverse.org/)**.

### Exercise 5

Pipe `cheese_tibble` into a plot that shows the number of cheeses by milk type, faceted by country, with flipped axes. Add this code to a new code chunk in your QMD. In a bash Terminal, run `quarto render analysis.qmd`. Check your Live Server tab.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r cheese-5}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 5)
```

###

Our plot:

```{r cheese-5-test}
cheese_tibble |>
  ggplot(aes(x = milk, y = n, fill = milk)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ country, scales = "fixed") +
  coord_flip() +
  scale_y_sqrt() +
  theme_gray() +
  labs(
    title = "Milk Types Used in Cheeses by Country",
    subtitle = "Italy has the most variety of cheese",
    caption = "Source: cheese.com",
    x = "Type of Milk",
    y = "Number of Cheeses (sqrt scale)"
  )
```

**Faceting** splits a plot into panels, one per level of a variable. Flipping the axes turns vertical bars into horizontal ones, which makes long category labels easier to read. Look at the France and USA panels — their bars are compressed because a single shared scale is set by Italy's much larger counts.

### Exercise 6

Look at the France and USA panels in our example plot above. Update the cheese plot so that each country's panel uses its own independent scale. In a bash Terminal, run `quarto render analysis.qmd`. Check your Live Server tab.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r cheese-6}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 5)
```

###

Our improved plot:

```{r cheese-6-test}
cheese_tibble |>
  ggplot(aes(x = milk, y = n, fill = milk)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ country, scales = "free_y") +
  coord_flip() +
  scale_y_sqrt() +
  theme_gray() +
  labs(
    title = "Milk Types Used in Cheeses by Country",
    subtitle = "Cow milk dominates across all three countries",
    caption = "Source: cheese.com",
    x = "Type of Milk",
    y = "Number of Cheeses (sqrt scale)"
  )
```

Giving each panel its own axis range — a *free* scale — lets every country's bars fill the available space instead of being dwarfed by the largest group. Use free scales when counts vary widely across groups and you care about the pattern *within* each group; keep a shared scale when the comparison *across* groups is the point.

### Exercise 7

Stage `analysis.qmd`, commit with the message `"Add cheese analysis"`, and push to GitHub.

CP/CR.

```{r cheese-7}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 5)
```

###

Committing regularly keeps your work safe and creates a clear history of what changed and when.

## Wine
###

The `wine` dataset contains the results of a chemical analysis of wines grown in a specific area of Italy. Three types of wine are represented in the 178 samples, with the results of 13 chemical analyses recorded for each sample. This data is from the UCI Machine Learning Repository. Our spreadsheet `data/wine.xlsx` contains this data in a spreadsheet format.

### Exercise 1

We begin by downloading our spreadsheet `wine.xlsx` directly from GitHub into the `data` directory.

In the R Terminal, run:

```
download.file(
  "https://github.com/PPBDS/misc.tutorials/raw/refs/heads/main/inst/tutorials/r4ds-2/data/wine.xlsx",
  destfile = "data/wine.xlsx",
  mode = "wb"
)
```

CP/CR.

```{r wine-1}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

###

The [UCI Machine Learning Repository](https://archive.ics.uci.edu/) hosts hundreds of benchmark datasets used in research and teaching. The wine dataset comes from a chemical analysis of wines from three cultivars grown in the same region of Italy — it is one of the most widely used classification datasets in machine learning.

### Exercise 2

Read the spreadsheet `data/wine.xlsx` into R. Add this code to a new code chunk in your QMD. In a bash Terminal, run `quarto render analysis.qmd`. Check your Live Server tab.

CP/CR.

```{r wine-2}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

###

Our result:

```{r wine-2-test}
read_excel("data/wine.xlsx")
```

Notice that the column names in this dataset are capitalized (`Phenols`, `Flavanoids`, `Color`) — unusual for tidy data. When asking AI to work with this dataset, use these exact column names and AI will match them correctly.

### Exercise 3

From `data/wine.xlsx`, keep only the `Phenols`, `Color`, and `Flavanoids` columns. Replace your current code in the same code chunk in your QMD with this. In a bash Terminal, run `quarto render analysis.qmd`.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r wine-3}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 6)
```

###

Our result:

```{r wine-3-test}
read_excel("data/wine.xlsx") |>
  select(Phenols, Color, Flavanoids)
```

If your result doesn't match ours, replace your chunk code with the following before moving on:

<pre><code>```{r}
read_excel("data/wine.xlsx") |>
  select(Phenols, Color, Flavanoids)
```</code></pre>

Selecting columns — keeping only the ones you name, or dropping the ones you don't — is one of the most basic data-wrangling moves, handled by **[dplyr](https://dplyr.tidyverse.org/)**. Narrowing to the columns you actually need keeps later steps and plots simple.

### Exercise 4

Within the recent code chunk, add the option `#| cache: true` and assign the result to `wine_tibble`. If you have not done so already, add `analysis_cache` to your `.gitignore`. In a bash Terminal, run `quarto render analysis.qmd`.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r wine-4}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 6)
```

###

Covariation is the tendency for the values of two or more variables to vary together in a related way. The best way to spot covariation is to visualize the relationship between variables using a scatter plot.

### Exercise 5

Create a scatter plot from `wine_tibble` with `Phenols` on the x-axis and `Flavanoids` on the y-axis, with color mapped to `Color`. Include jittered points, a LOESS smooth curve, and label the maximum `Flavanoids` value as "Outlier" in magenta. Use a blue-to-orange color gradient and a classic theme. Add this code to a new code chunk in your QMD. In a bash Terminal, run `quarto render analysis.qmd`. Check your Live Server tab.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r wine-5}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 12)
```

###

Our plot:

```{r wine-5-test}
wine_tibble |>
  ggplot(aes(x = Phenols, y = Flavanoids, color = Color)) +
  geom_jitter(size = 2, alpha = 0.5) +
  geom_smooth(method = "loess", se = TRUE, color = "black") +
  geom_text(
    data = \(d) d |> filter(Flavanoids == max(Flavanoids)),
    aes(label = "Outlier"),
    nudge_y = 0.2,
    color = "magenta",
    fontface = "bold"
  ) +
  scale_color_gradient(low = "blue", high = "orange") +
  theme_classic() +
  labs(
    title = "Wine: Flavanoids vs Phenols",
    x = "Phenols",
    y = "Flavanoids"
  )
```

**Jittering** adds a small amount of random noise to each point so that points which would otherwise land on top of one another become visible — the standard fix for *overplotting* in dense regions. The jittered points should replace the plain ones, not sit on top of them: drawing both puts every point on the plot twice.

### Exercise 6

Look at your rendered wine plot and compare it to our example above. If it draws each point twice — once at its true location and once with a random offset — it is layering plain points and jittered points on top of each other. Fix the wine plot so it uses a single jittered layer, choosing point size and transparency so dense regions stay readable. Render and confirm your plot matches our example.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r wine-6}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 5)
```

###

Layering plain points and jittered points draws every point twice — once at its exact location and once with a random offset. When you want jitter, use the jittered layer alone, and set point size and transparency so dense regions stay readable.

### Exercise 7

Stage `analysis.qmd`, commit with the message `"Add wine analysis"`, and push to GitHub.

CP/CR.

```{r wine-7}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 5)
```

###

Committing at natural section boundaries gives you checkpoints you can return to if something goes wrong later.

## Summary
###

This section covered key concepts from [Chapter 9: Layers](https://r4ds.hadley.nz/layers.html), [Chapter 10: Exploratory Data Analysis](https://r4ds.hadley.nz/eda.html), [Chapter 11: Communication](https://r4ds.hadley.nz/communication.html), and [Chapter 20: Spreadsheets](https://r4ds.hadley.nz/spreadsheets.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund.

You learned about working with categorical and continuous variables using different types of variation and exploratory data analysis with core packages such as **[readxl](https://readxl.tidyverse.org/)** and **[ggplot2](https://ggplot2.tidyverse.org/)**. Key **ggplot2** techniques included faceting plots into panels, flipping axes, and adding jitter to avoid overplotting.

### Exercise 1

In a bash Terminal, run `quarto render analysis.qmd` to ensure that everything works. Check your Live Server tab — the resulting HTML page should be attractive, showing clean versions of your plots.

In the R Terminal, run:

```
show_file("analysis.qmd")
```

CP/CR.

```{r summary-1}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 30)
```

###

Spreadsheet data in `.xls` and `.xlsx` formats is read into R with the **[readxl](https://readxl.tidyverse.org/)** package.

### Exercise 2

Publish your rendered QMD to GitHub Pages. In a bash Terminal, run:

````
quarto publish gh-pages analysis.qmd
````

If the bash Terminal is still running from rendering, stop it first using `Ctrl/Cmd + C`.

Copy/paste the resulting URL below.

```{r summary-2}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 1)
```

###

Publishing the QMD creates a public URL anyone can use to view the rendered project.

### Exercise 3

Commit and push any remaining changes. Copy/paste the URL to your GitHub repo.

```{r summary-3}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

###

GitHub repositories are where people can view the code you used to create your project.

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
