---
title: Fourth Tutorial for R4DS
author: David Kane
tutorial:
  id: r4ds-4
output:
  learnr::tutorial:
    df_print: default
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: 'This tutorial for "R for Data Science" covers Chapter 16: Factors, Chapter 17: Dates and Times, Chapter 18: Missing Values, Chapter 19: Joins, and Chapter 22: Arrow. Exercises use NBA and FIFA datasets to practice forcats, lubridate, table joins, and reading parquet files with arrow.'
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)

library(tidyverse)
library(arrow)
library(plotly)
library(viridis)
library(scales)

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(out.width = '90%')
options(tutorial.exercise.timelimit = 60,
        tutorial.storage = "local")

nba_clean <- open_dataset("data/game.parquet") |>
  collect() |>
  left_join(
    open_dataset("data/line_score.parquet") |> collect(),
    by = "game_id",
    relationship = "many-to-many"
  ) |>
  mutate(
    game_date = ymd_hms(game_date),
    game_date_est = ymd_hms(game_date_est)
  ) |>
  drop_na(pts_home.x) |>
  mutate(
    game_date = as_datetime(game_date),
    game_year = year(game_date),
    season_label = paste0(season_id %/% 10000, "-", str_sub(season_id + 1, 3, 4))
  ) |>
  filter(!is.na(game_date), game_year >= 1946) |>
  mutate(
    wl_home = factor(wl_home, levels = c("W", "L")),
    game_era = case_when(
      game_year <= 1979 ~ "Pre-Three Point",
      game_year <= 1999 ~ "Modern Era",
      game_year <= 2019 ~ "Analytics Era",
      TRUE ~ "Current Era"
    ),
    game_era = factor(game_era, levels = c("Pre-Three Point", "Modern Era", "Analytics Era", "Current Era"))
  ) |>
  mutate(
    across(contains("pct"), ~replace_na(.x, 0)),
    across(c(pts_home.x, pts_away.x), ~replace_na(.x, 0)),
    total_points = pts_home.x + pts_away.x
  ) |>
  filter(total_points > 0, !is.na(pts_home.x), !is.na(pts_away.x)) |>
  group_by(game_year, game_era, season_label) |>
  summarise(
    avg_points = mean(total_points, na.rm = TRUE),
    games_count = n(),
    home_win_pct = mean(wl_home == "W", na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(game_year)

final_fifa <- "data/fifa.parquet" |>
  open_dataset() |>
  collect() |>
  mutate(
    height_cm = as.numeric(str_extract(Height, "\\d+")),
    weight_kg = as.numeric(str_extract(Weight, "\\d+"))
  ) |>
  mutate(across(where(is.character), ~str_replace_all(.x, "\\n", ""))) |>
  mutate(across(where(is.character), ~str_trim(.x))) |>
  mutate(
    joined_date = mdy(Joined),
    years_at_club = as.numeric(today() - joined_date) / 365.25,
    long_tenure = years_at_club > 10
  ) |>
  drop_na(joined_date) |>
  mutate(
    value_numeric = case_when(
      is.na(Value) | Value == "" | Value == "0" ~ 0,
      str_detect(Value, "M") ~ parse_number(Value) * 1000000,
      str_detect(Value, "K") ~ parse_number(Value) * 1000,
      str_detect(Value, "^€?[0-9.]+$") ~ parse_number(Value),
      TRUE ~ 0
    ),
    wage_numeric = case_when(
      is.na(Wage) | Wage == "" | Wage == "0" ~ 0,
      str_detect(Wage, "K") ~ parse_number(Wage) * 1000,
      str_detect(Wage, "^€?[0-9.]+$") ~ parse_number(Wage),
      TRUE ~ 0
    ),
    release_clause_numeric = case_when(
      is.na(Release.Clause) | Release.Clause == "" | Release.Clause == "0" ~ 0,
      str_detect(Release.Clause, "M") ~ parse_number(Release.Clause) * 1000000,
      str_detect(Release.Clause, "K") ~ parse_number(Release.Clause) * 1000,
      str_detect(Release.Clause, "^€?[0-9.]+$") ~ parse_number(Release.Clause),
      TRUE ~ 0
    )
  )
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

## Introduction
###

This tutorial covers key concepts from [Chapter 16: Factors](https://r4ds.hadley.nz/factors.html), [Chapter 17: Dates and Times](https://r4ds.hadley.nz/datetimes.html), [Chapter 18: Missing Values](https://r4ds.hadley.nz/missing-values.html), [Chapter 19: Joins](https://r4ds.hadley.nz/joins.html), and [Chapter 22: Arrow](https://r4ds.hadley.nz/arrow.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund.

You will learn about working with parquet files using **[arrow](https://arrow.apache.org/docs/r/)**, manipulating categorical variables with **[forcats](https://forcats.tidyverse.org/)**, handling date/time data with **[lubridate](https://lubridate.tidyverse.org/)**, managing missing values, and performing joins. Key concepts include opening large datasets from disk without loading them fully into memory, reordering factor levels for plotting, parsing datetime strings, combining datasets with joins, and dropping rows with missing values.

We recommend using an agentic coding tool such as [Gemini CLI](https://github.com/google-gemini/gemini-cli) or [Claude Code](https://claude.ai/code). Our instructions are written with these tools in mind. You may also use a chat-based AI, but you will need to copy/paste code and data context manually.

### Exercise 1

Create a new repo using `codespace-starter` as a template and name it `r4ds-4`.

If you haven't already, open the `r4ds-4` repo on GitHub and create a Codespace from it.

You need two R Terminals: one for running this tutorial and one for your exercises. Open a second R Terminal using the dropdown button.

Create a new file and save it as `analysis.qmd`. Add a YAML header at the top with a title (`"NBA and FIFA"`) and your name as author. In a bash terminal, run:

```
quarto render analysis.qmd
```

Open `analysis.html` in a browser by right-clicking it in the File Explorer and selecting **Open with Live Server**. Keep this tab open — it will refresh automatically whenever you re-render.

Create a `.gitignore` file with `analysis_files` on the first line and then a blank line. Save and push.

In the R Terminal, run:

```
show_file(".gitignore")
```

If that fails, it is probably because you have not yet loaded `library(tutorial.helpers)` in the R Terminal.

CP/CR.

```{r introduction-1}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

###

Parquet files are "column-oriented" — organized column-by-column like R's data frame — which leads to better performance for data analysis than row-oriented CSV files. They also support efficient compression, making them much smaller than equivalent CSV files.

### Exercise 2

In your QMD, put the following in a new code chunk to load all required libraries:

```
suppressPackageStartupMessages({
  library(tidyverse)
  library(arrow)
  library(plotly)
  library(viridis)
  library(scales)
})
```

Also add the following to the YAML header to remove all code echoes from the HTML:

```
execute:
  echo: false
```

In a bash terminal, run `quarto render analysis.qmd`. Check your Live Server tab.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r introduction-2}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 6)
```

###

Like CSV, parquet is used for rectangular data, but it's a custom binary format designed for big data — not a text format you can read in any editor. Both **arrow** and **dbplyr** provide **dplyr** backends, so you can use the same **dplyr** code whether your data is in a parquet file or a database.

### Exercise 3

In a bash terminal, run `quarto render analysis.qmd`. Confirm that the Live Server tab auto-refreshes.

CP/CR.

```{r introduction-3}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

###

Opening an Arrow dataset scans only a few thousand rows to work out its structure, without loading the whole file into memory. You can then filter and select with ordinary **[dplyr](https://dplyr.tidyverse.org/)** verbs and pull only the data you actually need into R at the end.

### Exercise 4

From the R Terminal, run these three commands:

`getwd()`
`dir.create("data")`
`list.files()`

This will create a `data` directory in your project.

CP/CR.

```{r introduction-4}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 5)
```

###

Your answer should look something like this, although your path will be different.

```
> getwd()
[1] "/workspaces/r4ds-4"
> dir.create("data")
> list.files()
 [1] "analysis_files" "analysis.html"  "analysis.qmd"   "data"
>
```

Parquet relies on [efficient encodings](https://parquet.apache.org/docs/file-format/data-pages/encodings/) to keep file size down and supports file compression. This helps make parquet files fast because there's less data to move from disk to memory.


## NBA
###

The [NBA Database](https://www.kaggle.com/datasets/wyattowalsh/basketball) is a Kaggle dataset containing 64,000+ games, 4,800+ players, and all 30 teams. Our parquet files `data/game.parquet` and `data/line_score.parquet` contain game and scoring data converted from CSV to parquet format.

Key packages you'll use: **arrow** for reading parquet files, **lubridate** for datetime parsing, **dplyr** for joins and transformations, and **forcats** for ordered factor levels.

```{r nba-data}
open_dataset("data/game.parquet")
```

### Exercise 1

We begin by downloading our parquet files directly from GitHub.

In the R Terminal, run:

```
download.file(
  "https://github.com/PPBDS/misc.tutorials/raw/refs/heads/main/inst/tutorials/r4ds-4/data/game.parquet",
  destfile = "data/game.parquet",
  mode = "wb"
)
download.file(
  "https://github.com/PPBDS/misc.tutorials/raw/refs/heads/main/inst/tutorials/r4ds-4/data/line_score.parquet",
  destfile = "data/line_score.parquet",
  mode = "wb"
)
```

CP/CR.

```{r nba-1}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

###

Opening a parquet file gives you an **Arrow** dataset — a handle to data still sitting on disk, not loaded into R's memory. Only the final step, once you're ready, pulls the processed result into R.

### Exercise 2

Read the full `data/game.parquet` dataset into R. Don't assign the result to a variable. Add this code to a new code chunk in your QMD. In a bash terminal, run `quarto render analysis.qmd`. Check your Live Server tab.

CP/CR.

```{r nba-2}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

###

```{r nba-2-test}
open_dataset("data/game.parquet") |>
  collect()
```

A **left join** adds columns from a second table while keeping every row of the first (left) table. When a key value can match several rows on both sides, that is a *many-to-many* relationship — worth flagging explicitly, because it can multiply rows in ways you didn't intend.

### Exercise 3

Build a cleaned, summarized NBA dataset called `nba_clean`. You want one row per season, carrying the year, the era it falls in (Pre-Three Point through 1979, Modern Era through 1999, Analytics Era through 2019, Current Era after that), the average total points scored per game, the number of games, and the home-win percentage. Combine the game data with the line scores, keep seasons from 1946 on, and pull the result into R. Add the code to a new code chunk in your QMD. In a bash terminal, run `quarto render analysis.qmd`.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r nba-3}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 6)
```

###

Our result:

```{r nba-3-test}
nba_clean
```

If your result doesn't match ours, replace your chunk code with the following before moving on:

<pre><code>```{r}
open_dataset("data/game.parquet") |>
  collect() |>
  left_join(
    open_dataset("data/line_score.parquet") |> collect(),
    by = "game_id",
    relationship = "many-to-many"
  ) |>
  mutate(
    game_date = ymd_hms(game_date),
    game_date_est = ymd_hms(game_date_est)
  ) |>
  drop_na(pts_home.x) |>
  mutate(
    game_date = as_datetime(game_date),
    game_year = year(game_date),
    season_label = paste0(season_id %/% 10000, "-", str_sub(season_id + 1, 3, 4))
  ) |>
  filter(!is.na(game_date), game_year >= 1946) |>
  mutate(
    wl_home = factor(wl_home, levels = c("W", "L")),
    game_era = case_when(
      game_year <= 1979 ~ "Pre-Three Point",
      game_year <= 1999 ~ "Modern Era",
      game_year <= 2019 ~ "Analytics Era",
      TRUE ~ "Current Era"
    ),
    game_era = factor(game_era, levels = c("Pre-Three Point", "Modern Era", "Analytics Era", "Current Era"))
  ) |>
  mutate(
    across(contains("pct"), ~replace_na(.x, 0)),
    across(c(pts_home.x, pts_away.x), ~replace_na(.x, 0)),
    total_points = pts_home.x + pts_away.x
  ) |>
  filter(total_points > 0, !is.na(pts_home.x), !is.na(pts_away.x)) |>
  group_by(game_year, game_era, season_label) |>
  summarise(
    avg_points = mean(total_points, na.rm = TRUE),
    games_count = n(),
    home_win_pct = mean(wl_home == "W", na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(game_year)
```</code></pre>

**[lubridate](https://lubridate.tidyverse.org/)** parses date-times with functions named for the order of their components — year, month, day, then hour-minute-second. Building a categorical variable from a cascade of conditions (one era per range of years) is a common way to bin a continuous value, and giving the resulting **factor** an explicit level order keeps it in logical — not alphabetical — order in summaries and plots.

### Exercise 4

Within the latest code chunk, add the option: `#| cache: true`. Assign the result of the pipe to `nba_clean`. If you have not done so already, add `analysis_cache` to your `.gitignore`.

In a bash terminal, run `quarto render analysis.qmd`. This first render will take some time — by including `#| cache: true`, Quarto saves the result to disk so the next render loads from cache instead of re-running. Check your Live Server tab.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r nba-4}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 8)
```

###

Applying one transformation **across** many columns at once — selected by name, position, or a test like "every character column" — saves repeating yourself. Replacing missing values with a default is the right move when an NA really means zero (a count of none) rather than genuinely unknown; dropping rows with missing values is right when it doesn't.

### Exercise 5

Make a line plot from `nba_clean` showing average points per game over time, colored by `game_era`, with points sized by `games_count`. Give it a proper title, subtitle, and caption. Don't assign the plot. Add this code to a new code chunk in your QMD. In a bash terminal, run `quarto render analysis.qmd`. Check your Live Server tab.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r nba-5}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 12)
```

###

```{r nba-5-test}
nba_clean |>
  ggplot(aes(x = game_year, y = avg_points, color = game_era)) +
  geom_line(linewidth = 1.2, alpha = 0.8) +
  geom_point(aes(size = games_count), alpha = 0.7) +
  scale_color_viridis_d(option = "plasma") +
  scale_size_continuous(range = c(2, 6), guide = "none") +
  labs(
    title = "NBA Scoring Evolution: 75+ Years of Basketball",
    subtitle = "Average points per game by season with era classifications",
    x = "Season Year",
    y = "Average Points Per Game",
    color = "NBA Era"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 12, color = "gray60"),
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )
```

A **viridis** color scale for discrete categories is perceptually uniform and colorblind-accessible. And a static **ggplot2** chart can be made interactive — hover tooltips, zoom — by handing it to **[plotly](https://plotly.com/r/)**, a big payoff for very little extra work.

## FIFA
###

The [FIFA 21 dataset](https://www.kaggle.com/datasets/yagunnersya/fifa-21-messy-raw-dataset-for-cleaning-exploring/data) is scraped from [Sofifa.com](https://sofifa.com) and contains raw information about thousands of players including ratings, positions, value, and club. Our parquet file `data/fifa.parquet` contains this data.

Key packages you'll use: **stringr** for text cleaning, **lubridate** for date parsing, **dplyr** for conditional transformations, and **forcats** for factor manipulation.

```{r fifa-data}
"data/fifa.parquet" |>
  open_dataset() |>
  collect() |>
  glimpse()
```

### Exercise 1

We begin by downloading our parquet file `data/fifa.parquet` directly from GitHub.

In the R Terminal, run:

```
download.file(
  "https://github.com/PPBDS/misc.tutorials/raw/refs/heads/main/inst/tutorials/r4ds-4/data/fifa.parquet",
  destfile = "data/fifa.parquet",
  mode = "wb"
)
```

CP/CR.

```{r fifa-1}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

###

The FIFA dataset is "messy" — heights and weights include units as text, financial values use "M" and "K" suffixes, club names contain newline characters, and dates use a non-standard format. Real data often requires this kind of cleaning before analysis.

### Exercise 2

Read `data/fifa.parquet` into R and show its structure — the column types with a few example values. Don't assign the result. Add this code to a new code chunk in your QMD. In a bash terminal, run `quarto render analysis.qmd`. Check your Live Server tab.

CP/CR.

```{r fifa-2}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

###

```{r fifa-2-test}
"data/fifa.parquet" |>
  open_dataset() |>
  collect() |>
  glimpse()
```

**[lubridate](https://lubridate.tidyverse.org/)** can parse dates written in month-day-year order (e.g. "Jun 17, 2018"). And **[stringr](https://stringr.tidyverse.org/)** can pull the first run of digits out of a messy string like "185cm" or "76kg" — a quick way to turn text like that into a usable number.

### Exercise 3

Build a cleaned FIFA player dataset called `final_fifa`. You want numeric height and weight pulled out of the text columns; the character columns tidied up (newlines and stray spaces removed); the join date parsed and turned into the number of years at the club, plus a TRUE/FALSE flag for more than ten years; and the player's value, wage, and release clause converted from strings like "€110.5M" or "€565K" into plain euro numbers. Drop players with no join date. Add the code to a new code chunk in your QMD. In a bash terminal, run `quarto render analysis.qmd`.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r fifa-3}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 6)
```

###

Our result:

```{r fifa-3-test}
final_fifa
```

If your result doesn't match ours, replace your chunk code with the following before moving on:

<pre><code>```{r}
"data/fifa.parquet" |>
  open_dataset() |>
  collect() |>
  mutate(
    height_cm = as.numeric(str_extract(Height, "\\d+")),
    weight_kg = as.numeric(str_extract(Weight, "\\d+"))
  ) |>
  mutate(across(where(is.character), ~str_replace_all(.x, "\\n", ""))) |>
  mutate(across(where(is.character), ~str_trim(.x))) |>
  mutate(
    joined_date = mdy(Joined),
    years_at_club = as.numeric(today() - joined_date) / 365.25,
    long_tenure = years_at_club > 10
  ) |>
  drop_na(joined_date) |>
  mutate(
    value_numeric = case_when(
      is.na(Value) | Value == "" | Value == "0" ~ 0,
      str_detect(Value, "M") ~ parse_number(Value) * 1000000,
      str_detect(Value, "K") ~ parse_number(Value) * 1000,
      str_detect(Value, "^€?[0-9.]+$") ~ parse_number(Value),
      TRUE ~ 0
    ),
    wage_numeric = case_when(
      is.na(Wage) | Wage == "" | Wage == "0" ~ 0,
      str_detect(Wage, "K") ~ parse_number(Wage) * 1000,
      str_detect(Wage, "^€?[0-9.]+$") ~ parse_number(Wage),
      TRUE ~ 0
    ),
    release_clause_numeric = case_when(
      is.na(Release.Clause) | Release.Clause == "" | Release.Clause == "0" ~ 0,
      str_detect(Release.Clause, "M") ~ parse_number(Release.Clause) * 1000000,
      str_detect(Release.Clause, "K") ~ parse_number(Release.Clause) * 1000,
      str_detect(Release.Clause, "^€?[0-9.]+$") ~ parse_number(Release.Clause),
      TRUE ~ 0
    )
  )
```</code></pre>

Applying one transformation **across** many columns at once — say, all the character columns, or every column whose name contains "pct" — keeps the code short. The trickier step is the money: **[readr](https://readr.tidyverse.org/)** can pull the numeric part out of a messy string like `"€110.5M"`, ignoring the currency symbol and the suffix, which is exactly what turns these fields into clean numbers. (It also sidesteps a spurious "NAs introduced by coercion" warning that the more naive approach produces, because the cascade of conditions evaluates every branch over the whole column.)

### Exercise 4

Within the latest code chunk, add the option: `#| cache: true`. Assign the result of the pipe to `final_fifa`. If you have not done so already, add `analysis_cache` to your `.gitignore`.

In a bash terminal, run `quarto render analysis.qmd`. This first render will take some time. Check your Live Server tab.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r fifa-4}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 8)
```

###

**[forcats](https://forcats.tidyverse.org/)** is the tidyverse package for factors, and two moves come up constantly: reordering a factor's levels by another variable, so a bar or dot chart displays in a meaningful order rather than alphabetically; and lumping rare levels together into an "Other" category, keeping only the few most common ones.

### Exercise 5

From `final_fifa`, make a scatter plot of `Age` vs `X.OVA` (overall rating), with point size and color showing `value_numeric`. Filter to players with `value_numeric > 0` and `X.OVA > 60`. Use a viridis color scale and proper labels. Don't assign the plot. Add this code to a new code chunk in your QMD. In a bash terminal, run `quarto render analysis.qmd`. Check your Live Server tab.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r fifa-5}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 12)
```

###

```{r fifa-5-test}
final_fifa |>
  filter(value_numeric > 0, X.OVA > 60) |>
  ggplot(aes(x = Age, y = X.OVA, size = value_numeric, color = value_numeric)) +
  geom_point(alpha = 0.7) +
  scale_size_continuous(range = c(0.5, 8),
                        labels = label_currency(prefix = "€", suffix = "M", scale = 1e-6),
                        name = "Market Value") +
  scale_color_viridis_c(option = "plasma",
                        labels = label_currency(prefix = "€", suffix = "M", scale = 1e-6),
                        name = "Market Value") +
  labs(
    title = "FIFA Player Performance vs Age",
    subtitle = "Size and color represent market value",
    x = "Age",
    y = "Overall Rating",
    caption = "Data: FIFA 21 dataset"
  ) +
  theme_minimal()
```

An **ordered factor** tells **ggplot2** to use your chosen order for legends and axes instead of sorting alphabetically. And the **[scales](https://scales.r-lib.org/)** package formats axis numbers as currency — dollar signs, thousands separators, and scaling — so a chart of player values reads naturally.

## Summary
###

This tutorial covered key concepts from [Chapter 16: Factors](https://r4ds.hadley.nz/factors.html), [Chapter 17: Dates and Times](https://r4ds.hadley.nz/datetimes.html), [Chapter 18: Missing Values](https://r4ds.hadley.nz/missing-values.html), [Chapter 19: Joins](https://r4ds.hadley.nz/joins.html), and [Chapter 22: Arrow](https://r4ds.hadley.nz/arrow.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund.

You learned about working with parquet files using **[arrow](https://arrow.apache.org/docs/r/)**, manipulating categorical variables with **[forcats](https://forcats.tidyverse.org/)**, handling date/time data with **[lubridate](https://lubridate.tidyverse.org/)**, managing missing values, and performing joins.

### Exercise 1

In a bash terminal, run `quarto render analysis.qmd` to ensure that everything works. Check your Live Server tab — the resulting HTML page should be attractive, showing clean versions of your plots.

In the R Terminal, run:

```
show_file("analysis.qmd")
```

CP/CR.

```{r summary-1}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 30)
```

###

**arrow** only understands a subset of R expressions, so before you pull data into R you may not be able to write exactly the code you normally would — the rest has to wait until the data is in R. The **arrow** documentation lists which operations it can push down to the on-disk query.

### Exercise 2

Publish your rendered QMD to GitHub Pages. In a bash terminal, run:

```
quarto publish gh-pages analysis.qmd
```

If the bash terminal is still running from rendering, stop it first using `Ctrl/Cmd + C`.

Copy/paste the resulting URL below.

```{r summary-2}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 1)
```

###

Publishing the QMD creates a public URL anyone can use to view the rendered project.

### Exercise 3

Commit and push all your files. Copy/paste the URL to your GitHub repo.

```{r summary-3}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

###

GitHub repositories are where people can view the code you used to create your project.

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
