---
title: Fourth Tutorial for R4DS
author: David Kane
tutorial:
  id: r4ds-4
output:
  learnr::tutorial:
    df_print: default
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: "Arrow / Parquet files: cryptocurrency prices and prediction markets"
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(tidyverse)
library(arrow)
library(viridis)

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(out.width = '90%')
options(tutorial.exercise.timelimit = 60, tutorial.storage = "local")

# ŌöĆŌöĆ Prediction Markets ŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆŌöĆ

pm_markets <- read_parquet("../../extdata/r4ds-4/markets.parquet") |>
  mutate(
    created_at  = ymd_hms(created_at),
    closed_at   = ymd_hms(closed_at),
    resolved_at = ymd_hms(resolved_at)
  )

pm_history <- open_dataset("../../extdata/r4ds-4/history.parquet") |>
  collect()

pm_groups        <- read_parquet("../../extdata/r4ds-4/groups.parquet")
pm_market_groups <- read_parquet("../../extdata/r4ds-4/market_groups.parquet")

pm_joined <- pm_history |>
  left_join(pm_markets, by = "market_id") |>
  mutate(
    outcome_numeric    = if_else(outcome == "YES", 1, 0),
    prediction_error   = abs(probability - outcome_numeric),
    days_to_resolution = as.numeric(as.Date(resolved_at) - date),
    time_bucket = factor(
      case_when(
        days_to_resolution >  90 ~ "90+ days",
        days_to_resolution >  30 ~ "30ŌĆō90 days",
        days_to_resolution >   7 ~ "7ŌĆō30 days",
        days_to_resolution >=  1 ~ "1ŌĆō7 days",
        TRUE                      ~ "Final day"
      ),
      levels = c("90+ days", "30ŌĆō90 days", "7ŌĆō30 days",
                 "1ŌĆō7 days", "Final day"),
      ordered = TRUE
    )
  )

pm_accuracy <- pm_joined |>
  group_by(time_bucket) |>
  summarise(mean_error = mean(prediction_error), n_rows = n(), .groups = "drop")

pm_top_market <- pm_markets |>
  slice_max(total_traders, n = 1) |>
  pull(market_id)
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

## Introduction
###

This tutorial covers key concepts from [Chapter 16: Factors](https://r4ds.hadley.nz/factors.html), [Chapter 17: Dates and Times](https://r4ds.hadley.nz/datetimes.html), [Chapter 18: Missing Values](https://r4ds.hadley.nz/missing-values.html), [Chapter 19: Joins](https://r4ds.hadley.nz/joins.html), and [Chapter 22: Arrow](https://r4ds.hadley.nz/arrow.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine ├ćetinkaya-Rundel, and Garrett Grolemund.

You will learn about working with parquet files using **[arrow](https://arrow.apache.org/docs/r/)**, joining multiple datasets with **[dplyr](https://dplyr.tidyverse.org/)**, parsing dates with **[lubridate](https://lubridate.tidyverse.org/)**, cleaning messy strings with **[stringr](https://stringr.tidyverse.org/)**, and reordering categorical variables with **[forcats](https://forcats.tidyverse.org/)**. Key concepts include opening large datasets from disk without loading them fully into memory, joining tables through shared keys, parsing mixed-format date strings, and building an ordered factor for a final line chart.

We recommend using an agentic coding tool such as [Gemini CLI](https://github.com/google-gemini/gemini-cli) or [Claude Code](https://claude.ai/code). Our instructions are written with these tools in mind. You may also use a chat-based AI, but you will need to copy/paste code and data context manually.

### Exercise 1

You should be connected to a repo named `r4ds-4`. If you are not, create one and connect to it.

Create a new file, `analysis.qmd`, with the title `"Crypto Markets"` and your name as the author. In a bash Terminal, render it:

```
quarto render analysis.qmd
```

Open `analysis.html` with Live Server (right-click it in the Explorer ŌåÆ **Open with Live Server**) and keep the tab open. It refreshes on every render. Going forward, we will just tell you to "Render" when we want you to take these steps.

Create a `.gitignore` with `analysis_files` followed by a blank line. Commit and push.

In the R Terminal, run:

```
show_file(".gitignore")
```

If that fails, it is probably because you have not yet loaded `library(tutorial.helpers)` in the R Terminal.

CP/CR.

```{r introduction-1}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

###

```
analysis_files
```

###

Parquet files are column-oriented ŌĆö organized column-by-column like R's data frame ŌĆö which leads to better performance for data analysis than row-oriented CSV files. They also support efficient compression, making them much smaller than equivalent CSV files.

### Exercise 2

In your QMD, put the following in a new code chunk to load all required libraries:

```
suppressPackageStartupMessages({
  library(tidyverse)
  library(arrow)
  library(viridis)
})
```

Also add the following to the YAML header to remove all code echoes from the HTML:

```
execute:
  echo: false
```

Render. In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r introduction-2}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 6)
```

###

<pre><code>suppressPackageStartupMessages({
  library(tidyverse)
  library(arrow)
  library(viridis)
})
</code></pre>

###

Like CSV, parquet is used for rectangular data, but it is a custom binary format designed for big data ŌĆö not a text format you can read in any editor. Both **arrow** and **dbplyr** provide **dplyr** backends, so you can use the same **dplyr** code whether your data is in a parquet file or a database.

### Exercise 3

Create a `data` directory at the top level of the `r4ds-4` repo. In the bash Terminal, run:

```
ls
```

CP/CR.

```{r introduction-3}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 5)
```

###

```
analysis.html  analysis.qmd  analysis_files  data
```

###

Parquet relies on [efficient encodings](https://parquet.apache.org/docs/file-format/data-pages/encodings/) to keep file size down and supports file compression. This helps make parquet files fast because there is less data to move from disk to memory.


## Crypto markets
###

Bitcoin launched the cryptocurrency market in 2009. By 2013, it represented over 90% of all crypto by market value. But by 2021, Bitcoin's share had fallen to around 40%. Where did the other 60% go?

In this section you will join three parquet files ŌĆö daily prices, coin metadata, and category labels ŌĆö to build a line chart showing how each category's share of the total crypto market evolved from 2018 to 2024. Along the way you will use **[arrow](https://arrow.apache.org/docs/r/)** to open files lazily, **[lubridate](https://lubridate.tidyverse.org/)** to parse dates, **[stringr](https://stringr.tidyverse.org/)** to clean messy fields, and **[forcats](https://forcats.tidyverse.org/)** to reorder categories for the final visualization.

### Exercise 1

Download all three parquet files into your `data/` directory. The stable URLs follow the pattern `https://github.com/PPBDS/misc.tutorials/raw/refs/heads/main/inst/extdata/r4ds-4/<file>` for each of `daily_prices.parquet`, `coin_metadata.parquet`, and `categories.parquet`.

In the bash Terminal, run:

```
ls data
```

CP/CR.

```{r crypto-markets-1}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

###

```
categories.parquet    coin_metadata.parquet    daily_prices.parquet
```

###

Parquet files are column-oriented and compressed. The three files for this section total just over 2 MB ŌĆö small enough to keep in a CRAN package ŌĆö but `daily_prices.parquet` holds 59,436 rows of daily price and market cap data for 22 cryptocurrencies spanning 2018 to 2026.

### Exercise 2

Open `data/daily_prices.parquet` with `open_dataset()`. Don't assign it or collect it ŌĆö just print the result. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r crypto-markets-2}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

###

```{r crypto-markets-2-test}
#| echo: true
open_dataset("../../extdata/r4ds-4/daily_prices.parquet")
```

###

The result is not a tibble ŌĆö it is an `ArrowDataset`. Arrow is showing you the schema (column names and types) without loading any rows. No data has moved from disk to memory yet. `date_raw` is a character column, not a Date ŌĆö an intentional design choice so that you parse it explicitly in a later exercise.

### Exercise 3

Chain a `filter()` and a `select()` onto the `open_dataset()` call, then `collect()`. Filter to years 2020 and later, and select `date_raw`, `coin_id`, and `market_cap_usd`. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r crypto-markets-3}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 4)
```

###

```{r crypto-markets-3-test}
#| echo: true
open_dataset("../../extdata/r4ds-4/daily_prices.parquet") |>
  filter(as.integer(substr(date_raw, 1, 4)) >= 2020) |>
  select(date_raw, coin_id, market_cap_usd) |>
  collect()
```

###

`filter()` and `select()` on an Arrow dataset execute on disk before any rows transfer to R. Only `collect()` moves data into memory. Arrow's expression engine supports most base R string functions like `substr()` but does not support `lubridate::year()` in lazy evaluation ŌĆö which is why the filter uses `as.integer(substr(date_raw, 1, 4))`. For this dataset, filtering to 2020 and later reduces the transferred rows by roughly 30%.

### Exercise 4

Collect all three parquet files into tibbles. Assign them to `prices`, `metadata`, and `categories`. Print `metadata`. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r crypto-markets-4}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 5)
```

###

```{r crypto-markets-4-test}
#| echo: true
prices     <- open_dataset("../../extdata/r4ds-4/daily_prices.parquet") |> collect()
metadata   <- open_dataset("../../extdata/r4ds-4/coin_metadata.parquet") |> collect()
categories <- open_dataset("../../extdata/r4ds-4/categories.parquet") |> collect()
metadata
```

###

"January 3, 2009" is when Satoshi Nakamoto mined the Bitcoin genesis block ŌĆö the date that launched the entire cryptocurrency industry. The two date formats in `launch_date` ("January 3, 2009" and "2015-07-30") reflect a common real-world data quality issue: records pulled from different sources at different times use different date conventions. **lubridate** provides `mdy()` and `ymd()` to handle both.

### Exercise 5

Left-join `prices` to `metadata` by `coin_id`, then left-join the result to `categories` by `category_id`. Both tables have a `description` column ŌĆö drop the one from `categories` before joining (it is not needed). Assign the result to `crypto`. Print the first 5 rows. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r crypto-markets-5}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 5)
```

###

```{r crypto-markets-5-test}
#| echo: true
crypto <- prices |>
  left_join(metadata, by = "coin_id") |>
  left_join(select(categories, -description), by = "category_id")
head(crypto, 5)
```

###

A left join keeps every row from the left table and adds matching columns from the right. Here, each of the 59,436 daily observations in `prices` picks up its coin's name, category, launch date, and description from `metadata`, and its category display name from `categories`. Notice that `category_id` was not in `daily_prices.parquet` ŌĆö it came from the first join. That is why two joins were needed: prices ŌåÆ metadata adds `category_id`, then metadata ŌåÆ categories adds the human-readable category name.

### Exercise 6

`market_cap_str` contains values like `"$1.24T"`, `"$450.2B"`, `"$37.5M"`, and some `NA`. Add a new column `market_cap_parsed` to `crypto` that extracts the numeric value: detect the suffix letter, extract the number with `parse_number()`, and multiply by the appropriate scale (1e12 for T, 1e9 for B, etc.). Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r crypto-markets-6}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 8)
```

###

```{r crypto-markets-6-test}
#| echo: true
crypto <- crypto |>
  mutate(
    market_cap_parsed = case_when(
      str_detect(market_cap_str, "T") ~ parse_number(market_cap_str) * 1e12,
      str_detect(market_cap_str, "B") ~ parse_number(market_cap_str) * 1e9,
      str_detect(market_cap_str, "M") ~ parse_number(market_cap_str) * 1e6,
      str_detect(market_cap_str, "K") ~ parse_number(market_cap_str) * 1e3,
      !is.na(market_cap_str) ~ parse_number(market_cap_str),
      TRUE ~ NA_real_
    )
  )
```

###

`parse_number()` strips non-numeric characters like `$` but cannot handle a scale suffix (`T`, `B`, `M`). The `case_when()` + `parse_number()` pattern shown here is the standard approach for parsing formatted financial strings in R. About 5% of `market_cap_str` values are `NA` ŌĆö simulating the kind of gap that appears when a dashboard exports data with occasional rendering failures. The clean numeric column `market_cap_usd` is already in the dataset; `market_cap_parsed` exists solely to practice this parsing pattern.

### Exercise 7

Parse dates. In `metadata`, detect which `launch_date` values use the long-form "Month Day, Year" format and parse them accordingly ŌĆö the remaining values use "YYYY-MM-DD". Store the parsed result in a new column `launch_date_parsed`. Also parse `date_raw` in `crypto` into a proper date column called `date`. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r crypto-markets-7}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 8)
```

###

```{r crypto-markets-7-test}
#| echo: true
metadata <- metadata |>
  mutate(
    launch_date_parsed = case_when(
      str_detect(launch_date, ",") ~ mdy(launch_date),
      TRUE ~ ymd(launch_date)
    )
  )

crypto <- crypto |>
  mutate(date = ymd(date_raw))
```

###

Lubridate's `mdy()` and `ymd()` are forgiving about separators but strict about field order. When a dataset has two different date formats, conditional parsing with `case_when()` is safer than hoping one parser handles both. Bitcoin's `launch_date` parses to 2009-01-03; Ethereum's parses to 2015-07-30 ŌĆö both correct despite looking completely different as strings.

### Exercise 8

Clean the `description` column in `crypto` by replacing embedded newlines (`\n`) with a space. Confirm the fix by filtering to `btc` and `eth` and showing their distinct `coin_id` and `description` values. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r crypto-markets-8}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 6)
```

###

```{r crypto-markets-8-test}
#| echo: true
crypto <- crypto |>
  mutate(description = str_replace_all(description, "\\n", " "))

crypto |>
  filter(coin_id %in% c("btc", "eth")) |>
  distinct(coin_id, description)
```

###

Embedded newlines appear when data is scraped from a web page or copied from a document that used visual line breaks. They silently break `str_detect()`, `str_length()`, and CSV export ŌĆö the text looks fine in a print statement but causes unexpected behavior downstream. `str_replace_all()` with `"\\n"` (the regex for a literal newline) replaces every occurrence; the double backslash is required because R parses the first `\` as an escape character before the regex engine sees the string.

### Exercise 9

Add an `era` column to `crypto` using `case_when()` on `year(date)`: `"Bitcoin Era"` (2016 and earlier), `"ICO/Ethereum Era"` (2017ŌĆō2019), `"DeFi Summer"` (2020ŌĆō2021), `"Maturation Era"` (2022 onward). Convert it to an ordered factor with levels in chronological order. Count rows per era. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r crypto-markets-9}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 10)
```

###

```{r crypto-markets-9-test}
#| echo: true
crypto <- crypto |>
  mutate(
    era = factor(
      case_when(
        year(date) <= 2016 ~ "Bitcoin Era",
        year(date) <= 2019 ~ "ICO/Ethereum Era",
        year(date) <= 2021 ~ "DeFi Summer",
        TRUE ~ "Maturation Era"
      ),
      levels = c("Bitcoin Era", "ICO/Ethereum Era", "DeFi Summer", "Maturation Era")
    )
  )

count(crypto, era)
```

###

The ICO era (2017ŌĆō2018) saw hundreds of new coins raise billions through initial coin offerings ŌĆö mostly on the Ethereum network. Most of those projects failed. DeFi Summer (2020) was different: the protocols that launched ŌĆö Uniswap, Aave, Compound, MakerDAO ŌĆö created real financial primitives that still process billions of dollars today. The era labels here are interpretive, not standardized; different analysts draw the boundaries differently. What matters for the tutorial is that `factor()` with explicit `levels` gives you control over the order ggplot2 uses for color legends and facet panels.

### Exercise 10

The `crypto` chunk is getting long. Clean it up: keep only the code that defines `crypto` (the data loading, joins, and all the mutations), and remove the `count(crypto, era)` line at the bottom.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r crypto-markets-10}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 20)
```

###

<pre><code>prices     <- open_dataset("data/daily_prices.parquet") |> collect()
metadata   <- open_dataset("data/coin_metadata.parquet") |> collect()
categories <- open_dataset("data/categories.parquet") |> collect()

metadata <- metadata |>
  mutate(
    launch_date_parsed = case_when(
      str_detect(launch_date, ",") ~ mdy(launch_date),
      TRUE ~ ymd(launch_date)
    )
  )

crypto <- prices |>
  left_join(metadata, by = "coin_id") |>
  left_join(select(categories, -description), by = "category_id") |>
  mutate(
    market_cap_parsed = case_when(
      str_detect(market_cap_str, "T") ~ parse_number(market_cap_str) * 1e12,
      str_detect(market_cap_str, "B") ~ parse_number(market_cap_str) * 1e9,
      str_detect(market_cap_str, "M") ~ parse_number(market_cap_str) * 1e6,
      str_detect(market_cap_str, "K") ~ parse_number(market_cap_str) * 1e3,
      !is.na(market_cap_str) ~ parse_number(market_cap_str),
      TRUE ~ NA_real_
    ),
    date = ymd(date_raw),
    description = str_replace_all(description, "\\n", " "),
    era = factor(
      case_when(
        year(date) <= 2016 ~ "Bitcoin Era",
        year(date) <= 2019 ~ "ICO/Ethereum Era",
        year(date) <= 2021 ~ "DeFi Summer",
        TRUE ~ "Maturation Era"
      ),
      levels = c("Bitcoin Era", "ICO/Ethereum Era", "DeFi Summer", "Maturation Era")
    )
  )
</code></pre>

###

The fully joined and cleaned `crypto` tibble has one row per coin per day ŌĆö 59,436 rows covering 22 coins from 2018 through 2026. Bitcoin's dominance, Ethereum's rise, stablecoin growth, and the meme-coin spike of 2021 are all latent in this table, waiting to be drawn out by the grouping and plotting exercises ahead.

### Exercise 11

Add `#| cache: true` to the chunk. Render. This first render with caching will take a moment; subsequent renders will load from disk.

In the bash Terminal, run:

```
ls
```

CP/CR.

```{r crypto-markets-11}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 5)
```

###

```
analysis.html  analysis.qmd  analysis_cache  analysis_files  data
```

###

Stablecoins like USDT grew from roughly $4B market cap in early 2019 to over $80B by 2022 ŌĆö not because the token's price rose, but because more dollars flowed into the crypto ecosystem seeking a stable store of value. USDT, USDC, and DAI all maintained their $1 peg throughout; their rising market cap reflects real-dollar demand for crypto-denominated stability, not speculation.

### Exercise 12

Add `analysis_cache` to your `.gitignore`. The Source Control change count in VS Code jumped when the cache directory appeared; adding it to `.gitignore` drops it back. Render.

In the R Terminal, run:

```
show_file(".gitignore")
```

CP/CR.

```{r crypto-markets-12}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 4)
```

###

```
analysis_files
analysis_cache
```

###

The FTX exchange token FTT had a market cap of roughly $7.25B on November 8, 2022. By November 9 it had fallen to $1.82B ŌĆö a 75% single-day drop. FTX, one of the largest crypto exchanges, filed for bankruptcy three days later. The collapse is visible as a near-vertical line in the data; it is one of the clearest single-event signals in the entire dataset.

### Exercise 13

In a new code chunk, compute the total market cap per day across all coins in `crypto` (sum `market_cap_usd` by `date`). Plot it as a line chart with date on the x-axis and total market cap in trillions on the y-axis. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r crypto-markets-13}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 8)
```

###

```{r crypto-markets-13-test}
#| echo: true
crypto |>
  group_by(date) |>
  summarise(total_mcap = sum(market_cap_usd, na.rm = TRUE)) |>
  ggplot(aes(x = date, y = total_mcap / 1e12)) +
  geom_line() +
  labs(x = NULL, y = "Total market cap (trillions USD)",
       title = "Total crypto market capitalization, 2020ŌĆō2026")
```

###

The total crypto market cap grew from roughly $150B in early 2019 to over $2.5T at its November 2021 peak ŌĆö a 16├Ś increase in under three years. It then fell sharply through 2022 (the "crypto winter"), driven by the Luna/UST collapse in May and FTX in November. By 2024 it had recovered above $2T. This context is essential for the next chart: Bitcoin's *share* fell during the bull market while its *absolute value* rose.

### Exercise 14

Compute daily market cap per category by summing `market_cap_usd` across all coins in each `category_name`, per `date`. Plot a multi-line chart with date on the x-axis and category market cap on the y-axis, colored by `category_name`. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r crypto-markets-14}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 8)
```

###

```{r crypto-markets-14-test}
#| echo: true
crypto |>
  group_by(date, category_name) |>
  summarise(cat_mcap = sum(market_cap_usd, na.rm = TRUE), .groups = "drop") |>
  ggplot(aes(x = date, y = cat_mcap / 1e12, color = category_name)) +
  geom_line() +
  labs(x = NULL, y = "Market cap (trillions USD)", color = NULL,
       title = "Crypto market cap by category, 2020ŌĆō2026")
```

###

This chart is hard to read because Bitcoin dominates the y-axis. Its absolute market cap dwarfs all other categories, compressing the other lines near zero. When one category dominates the absolute scale, relative share ŌĆö each category as a percentage of the total ŌĆö is the standard way to make the other trajectories visible.

### Exercise 15

Compute each category's share as a percentage of the total daily market cap. Plot the same multi-line chart with share (%) on the y-axis. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r crypto-markets-15}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 10)
```

###

```{r crypto-markets-15-test}
#| echo: true
crypto |>
  group_by(date, category_name) |>
  summarise(cat_mcap = sum(market_cap_usd, na.rm = TRUE), .groups = "drop") |>
  group_by(date) |>
  mutate(total_mcap = sum(cat_mcap),
         share_pct = cat_mcap / total_mcap * 100) |>
  ungroup() |>
  ggplot(aes(x = date, y = share_pct, color = category_name)) +
  geom_line() +
  labs(x = NULL, y = "Market cap share (%)", color = NULL,
       title = "Crypto market cap share by category, 2020ŌĆō2026")
```

###

Now the story is visible. Bitcoin's line starts above 50% and shows a declining arc through 2021 before recovering. Stablecoins grow steadily ŌĆö especially through 2022 when investors fled volatile assets. DeFi protocols spike in 2020ŌĆō2021 then partially retreat. Meme coins (DOGE) cause a sharp spike in May 2021 ŌĆö Elon Musk tweets drove DOGE's market cap briefly above $80B ŌĆö then almost entirely collapse within weeks. The diversity of trajectories across categories is exactly the answer to the guiding mystery.

### Exercise 16

Reorder the `category_name` factor so that categories with the highest median share appear first in the legend. Apply a colorblind-safe color palette. Add a proper title, subtitle, axis labels, and a caption crediting the source. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r crypto-markets-16}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 14)
```

###

```{r crypto-markets-16-test}
#| echo: true
crypto |>
  group_by(date, category_name) |>
  summarise(cat_mcap = sum(market_cap_usd, na.rm = TRUE), .groups = "drop") |>
  group_by(date) |>
  mutate(total_mcap = sum(cat_mcap),
         share_pct = cat_mcap / total_mcap * 100) |>
  ungroup() |>
  mutate(category_name = fct_reorder(category_name, share_pct, .fun = median, .desc = TRUE)) |>
  ggplot(aes(x = date, y = share_pct, color = category_name)) +
  geom_line(linewidth = 0.8) +
  scale_color_viridis_d(option = "plasma", end = 0.9) +
  labs(x = NULL, y = "Share of total market cap (%)", color = NULL,
       title = "Who ate Bitcoin's lunch?",
       subtitle = "Crypto market cap share by category, 2020ŌĆō2026",
       caption = "Source: Coin Metrics Community Data (CC BY 4.0)")
```

###

`fct_reorder()` sorts factor levels by a summary statistic ŌĆö here, the median share across all dates. This puts Bitcoin at the top of the legend (it has the highest median), which aligns with its line being highest in the chart. When legend order matches visual order, readers scan less. `viridis` and `viridis_d` are colorblind-safe palettes that also print clearly in grayscale ŌĆö a good default for any chart that may be viewed in multiple contexts.

### Exercise 17

Add annotations to highlight two historical events: the Luna/UST crash (approximately May 7ŌĆō12, 2022) and the FTX collapse (approximately November 8ŌĆō11, 2022). Use `annotate("rect", ...)` or `geom_vline()` ŌĆö your choice. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r crypto-markets-17}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 12)
```

###

```{r crypto-markets-17-test}
#| echo: true
crypto |>
  group_by(date, category_name) |>
  summarise(cat_mcap = sum(market_cap_usd, na.rm = TRUE), .groups = "drop") |>
  group_by(date) |>
  mutate(total_mcap = sum(cat_mcap),
         share_pct = cat_mcap / total_mcap * 100) |>
  ungroup() |>
  mutate(category_name = fct_reorder(category_name, share_pct, .fun = median, .desc = TRUE)) |>
  ggplot(aes(x = date, y = share_pct, color = category_name)) +
  geom_line(linewidth = 0.8) +
  scale_color_viridis_d(option = "plasma", end = 0.9) +
  annotate("rect", xmin = as.Date("2022-05-07"), xmax = as.Date("2022-05-12"),
           ymin = -Inf, ymax = Inf, alpha = 0.15, fill = "red") +
  annotate("rect", xmin = as.Date("2022-11-08"), xmax = as.Date("2022-11-11"),
           ymin = -Inf, ymax = Inf, alpha = 0.15, fill = "red") +
  labs(x = NULL, y = "Share of total market cap (%)", color = NULL,
       title = "Who ate Bitcoin's lunch?",
       subtitle = "Crypto market cap share by category, 2020ŌĆō2026",
       caption = "Source: Coin Metrics Community Data (CC BY 4.0)")
```

###

Terra's algorithmic stablecoin UST maintained its $1 peg through a supply-demand mechanism rather than dollar reserves. In May 2022, a large sell-off broke the mechanism; both UST and its sister token Luna lost almost all their value within 72 hours ŌĆö the largest single wealth-destruction event in crypto history at that time. Fiat-backed stablecoins like USDT and USDC held their pegs throughout; their divergence from UST is visible in the stablecoin category line, which dips slightly in May 2022 then resumes growth.

### Exercise 18

Commit `analysis.qmd` with the message `"Crypto market composition analysis"` and push to GitHub.

```{r crypto-markets-18}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

###

The 22 coins in this dataset represent the major categories but exclude thousands of smaller coins ŌĆö category shares here are approximations, not exact measures of the full crypto market. By 2026, the top 10 coins by market cap still accounted for roughly 80% of total market value, so the broad patterns hold even though the tail is missing.

## Prediction markets
###

Picture a market where people bet on whether a future event will happen: a new
law passes, a company goes bankrupt, a model breaks a benchmark. Each trade
moves the price toward the bettor's belief, and the current price is the
crowd's best estimate of the probability of YES. As the resolution date
approaches and more information becomes available, does the crowd get smarter?

In this section you will join four parquet files representing 800 resolved
markets and 43,694 daily probability snapshots to answer that question. Along
the way you will encounter a **many-to-many join** ŌĆö a table relationship that
silently multiplies rows if you are not careful ŌĆö and use
**[lubridate](https://lubridate.tidyverse.org/)** and
**[forcats](https://forcats.tidyverse.org/)** to bucket time and build an
accuracy chart.

### Exercise 1

Download `markets.parquet` and `history.parquet` into `data/`. The stable URLs are:

`https://raw.githubusercontent.com/ptweeet/misc.tutorials/refs/heads/main/inst/extdata/r4ds-4/markets.parquet`
`https://raw.githubusercontent.com/ptweeet/misc.tutorials/refs/heads/main/inst/extdata/r4ds-4/history.parquet`

Open `history.parquet` without loading it into memory ŌĆö assign the result to `history_lazy` and print it. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r prediction-markets-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 4)
```

###

```{r prediction-markets-1-test}
#| echo: true
open_dataset("../../extdata/r4ds-4/history.parquet")
```

Expected output ŌĆö the schema printed without loading data into memory:

```
FileSystemDataset with 1 Parquet file
43,694 rows x 3 columns
$ market_id   <string>
$ date        <date32[day]>
$ probability <double>
```

###

**Knowledge drop:** Manifold Markets is a play-money prediction platform
launched in 2022. Users trade with Mana (ß╣Ć), a fictional currency ŌĆö not real
dollars. The crowd is still incentivized to be accurate because Mana has
social and competitive value, and research suggests play-money markets are
well-calibrated on short-horizon events. The events in this dataset are real
ŌĆö AI model releases, election outcomes, geopolitical events ŌĆö even if the
stakes are not.

### Exercise 2

Bring `history_lazy` into memory. Left-join it to `markets.parquet`, which you should also read into memory. Add a column `outcome_numeric` that is 1 when the outcome is YES and 0 otherwise. Assign the result to `joined`. Print the number of rows. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r prediction-markets-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r prediction-markets-2-test}
#| echo: true
nrow(pm_joined)
```

Expected output:

```
[1] 43694
```

The row count does not change. This is a one-to-many join: one market row
matched to many daily history rows. Each of the 43,694 history rows gets one
matching market row appended.

###

**Knowledge drop:** Prediction markets have a YES-bias problem in reverse:
63% of markets in this dataset resolved NO. Most questions take the form
"Will X happen?" ŌĆö and X usually doesn't. The outcome distribution (63% NO,
37% YES) is worth keeping in mind when interpreting summary statistics,
because averaging across outcomes is really averaging across two different
crowd behaviors.

### Exercise 3

Download `groups.parquet` and `market_groups.parquet` into `data/`. The stable URLs are:

`https://raw.githubusercontent.com/ptweeet/misc.tutorials/refs/heads/main/inst/extdata/r4ds-4/groups.parquet`
`https://raw.githubusercontent.com/ptweeet/misc.tutorials/refs/heads/main/inst/extdata/r4ds-4/market_groups.parquet`

Read both into memory. Print the structure of each ŌĆö types and a few example values per column. How many rows does each have? What is the relationship between these two files and `markets`? Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r prediction-markets-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

```{r prediction-markets-3-test}
#| echo: true
glimpse(pm_groups)
glimpse(pm_market_groups)
```

Expected output:

```
Rows: 115
Columns: 4
$ group_slug     <chr> "ai", "technology-default", ...
$ group_name     <chr> "Ai", "Technology Default", ...
$ broad_category <ord> AI / Technology, AI / Technology, ...
$ market_count   <int> 110, 160, ...

Rows: 2,820
Columns: 2
$ market_id  <chr> "...", ...
$ group_slug <chr> "ai", ...
```

###

**Knowledge drop:** The `groups` file is a reference table: one row per topic
group. The `market_groups` file is a **junction table**: it records which
groups each market belongs to. A single market can belong to several groups
(e.g., both "AI model releases" and "Technology"), and a single group contains
many markets. This many-to-many structure is about to matter.

###

### Exercise 4

Left-join `markets` to `market_groups` by `market_id`. Print the number of rows in the result. Compare to the original 800-row `markets` table and describe what happened. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r prediction-markets-4}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```{r prediction-markets-4-test}
#| echo: true
pm_markets |>
  left_join(pm_market_groups, by = "market_id") |>
  nrow()
```

Expected output:

```
[1] 2885
```

800 rows became 2,885 rows. The 735 markets that belong to at least one group
each appear multiple times ŌĆö once per group membership ŌĆö producing 2,820
group-matched rows. The 65 markets with no matching groups each produce one
row with `NA` for `group_slug`. A **many-to-many join** has silently multiplied
rows. Any per-market summary computed on this inflated table would be wrong.

###

**Knowledge drop:** A many-to-many join occurs when the join key appears
multiple times on *both* sides of the join: each market belongs to several
groups, and each group contains many markets. **dplyr** warns about this when
you pass `relationship = "many-to-many"` explicitly. The safest pattern: finish
all market-level calculations before joining to the junction table ŌĆö never join
first and summarize second.

### Exercise 5

Left-join `market_groups` to `groups` by `group_slug`. Count how many rows fall into each `broad_category`. Which category has the most? Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r prediction-markets-5}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```{r prediction-markets-5-test}
#| echo: true
pm_market_groups |>
  left_join(pm_groups, by = "group_slug") |>
  count(broad_category, sort = TRUE)
```

Expected output (approximate ŌĆö verify against actual data):

```
# A tibble: 9 x 2
  broad_category               n
  <ord>                    <int>
1 AI / Technology            ...
2 Politics                   ...
3 Economics / Finance        ...
4 World / Geopolitics        ...
...
```

The direction of this join matters: you are joining the junction table *up* to
the reference table, which is a many-to-one relationship with no row explosion.

###

**Knowledge drop:** The dataset skews heavily toward AI, politics, and
economics ŌĆö the three domains most actively traded on Manifold in late 2025.
AI topics dominate partly because the Manifold user base overlaps heavily with
the AI research and developer communities who have strong opinions (and
sometimes insider knowledge) about model releases and benchmark results.

### Exercise 6

In `markets`, parse the `created_at` and `resolved_at` columns into proper datetime objects. Compute `duration_days` ŌĆö the number of days between creation and resolution. What is the median duration? The maximum? Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r prediction-markets-6}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r prediction-markets-6-test}
#| echo: true
pm_markets |>
  summarise(
    median_duration = median(as.numeric(
      as.Date(resolved_at) - as.Date(created_at)
    )),
    max_duration = max(as.numeric(
      as.Date(resolved_at) - as.Date(created_at)
    ))
  )
```

Expected output:

```
# A tibble: 1 x 2
  median_duration max_duration
            <dbl>        <dbl>
1              37          241
```

###

**Knowledge drop:** Most Manifold markets are created in response to breaking
news and resolve within weeks. The median duration of 37 days reflects this.
The data covers Sep 2025 ŌĆō May 2026, a period that included Trump's second-term
policy rollout, ongoing AI model releases, the Iran nuclear negotiations, and
several crypto market events. Short median durations mean most of the
interesting probability movement happens in a short window ŌĆö which is exactly
why the bucket analysis will reveal a sharp accuracy improvement near the
deadline.

### Exercise 7

In your joined table, compute `prediction_error` ŌĆö the absolute difference between the crowd's probability and the true outcome. Summarize its distribution: what is the mean error? The median? Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r prediction-markets-7}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

```{r prediction-markets-7-test}
#| echo: true
pm_joined |>
  summarise(
    mean_error   = mean(prediction_error),
    median_error = median(prediction_error)
  )
```

Expected output (approximately):

```
# A tibble: 1 x 2
  mean_error median_error
       <dbl>        <dbl>
1      0.277        0.214
```

###

**Knowledge drop:** The unit of `prediction_error` is probability units (0 to
1). An error of 0.28 means the crowd's forecast was 28 percentage points away
from the true outcome, on average. An error of 0 is a perfect prediction; an
error of 1 is completely wrong. The wide spread (some predictions close to 0,
others close to 1) suggests that averaging across all time horizons masks a
lot of variation ŌĆö which the bucketing exercise will reveal.

### Exercise 8

Still in your joined table, compute `days_to_resolution` ŌĆö the number of calendar days remaining until each market's resolution date. Summarize: how many rows have zero days remaining (the final day)? How many have one to seven days remaining? Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r prediction-markets-8}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

```{r prediction-markets-8-test}
#| echo: true
pm_joined |>
  summarise(
    final_day    = sum(days_to_resolution == 0),
    one_to_seven = sum(days_to_resolution >= 1 & days_to_resolution <= 7)
  )
```

Expected output:

```
# A tibble: 1 x 2
  final_day one_to_seven
      <int>        <int>
1       800         5600
```

Exactly 800 final-day rows ŌĆö one per market, since `resolved_at` is the
calendar day of resolution. The 5,600 rows in the 1ŌĆō7 day window come from
800 markets each with up to 7 days in that range.

###

**Knowledge drop:** The 43,694 total rows are not evenly distributed across
time. Most rows fall in the 7ŌĆō90 day window because that is where most of a
typical market's lifetime is spent. The 800 final-day rows and 5,600
one-to-seven-day rows together represent the "late game" ŌĆö where the accuracy
signal will be strongest.

###

### Exercise 9

Your working chunk now builds the joined table with `outcome_numeric`,
`prediction_error`, and `days_to_resolution`. Clean it up: remove any
`glimpse()`, `count()`, `summarise()`, or `nrow()` calls. Keep only the
pipeline that defines the joined table and assigns it to a name. Label the
chunk `prediction-markets-data` by adding `#| label: prediction-markets-data`
to the chunk options. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r prediction-markets-9, echo=FALSE}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 10)
```

###

Expected: the chunk should contain only the pipeline that builds the joined
table (with `outcome_numeric`, `prediction_error`, `days_to_resolution`),
labeled `prediction-markets-data`, with no diagnostic calls.

###

**Knowledge drop:** The markets in this dataset span a wide range of topics.
The most accurately predicted markets tend to be AI model release dates ŌĆö
where the research community follows development timelines closely and has
strong priors. The least accurately predicted tend to be sudden geopolitical
events ŌĆö elections with unexpected results, negotiation collapses, rapid
policy reversals ŌĆö where new information arrives without warning and the
crowd has no privileged knowledge.

### Exercise 10

Add `#| cache: true` to the `prediction-markets-data` chunk. Render. In the
Terminal, run:

```
ls analysis_cache/
```

Paste the output below. CP/CR.

```{r prediction-markets-10, echo=FALSE}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

Expected output: a subdirectory named after the cached chunk, something like:

```
prediction-markets-data_<hash>/
```

###

**Knowledge drop:** `#| cache: true` stores the R objects a chunk produces in
`analysis_cache/`. On the next render, Quarto checks whether the chunk's source
code has changed; if not, it loads the cached objects instead of re-running the
code. For Arrow pipelines this saves time because `collect()` and joins over
large datasets can be slow. The `market_groups` junction table also illustrates
why row counts deserve attention: joining `markets` (800 rows) to
`market_groups` yields 2,820 rows ŌĆö each market appears once per group it
belongs to. Caching means you pay that cost only once.

### Exercise 11

Add `analysis_cache/` to your `.gitignore`. In the R Terminal, run:

```
show_file(".gitignore")
```

Paste the output below. CP/CR.

```{r prediction-markets-11, echo=FALSE}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

Expected: `.gitignore` contains `analysis_cache/` on its own line.

###

**Knowledge drop:** The final-day mean absolute error across all 800 markets
is 0.090. This means that even on the last day, the crowd is still meaningfully
uncertain about roughly 1 in 10 markets. These are the "black swan" resolutions
ŌĆö events that looked nearly certain until they weren't. Prediction markets are
not oracles. They are very good at pooling information, but they cannot predict
genuinely unpredictable events, no matter how close the deadline is.

###

### Exercise 12

Find the market with the most traders. Extract its `market_id`, filter the
joined table to that market's history rows, and plot `probability` over `date`,
coloring the line by `outcome`. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r prediction-markets-12, echo=FALSE}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r prediction-markets-12-test}
#| echo: true
pm_joined |>
  filter(market_id == pm_top_market) |>
  ggplot(aes(x = date, y = probability, color = outcome)) +
  geom_line(linewidth = 1.2) +
  labs(
    title = "Probability trajectory for most-traded market",
    x     = "Date",
    y     = "Crowd-implied probability",
    color = "Outcome"
  ) +
  theme_minimal()
```

Expected: a single probability trajectory from market creation to resolution.
The line wanders and then converges toward 0 or 1 as the deadline approaches.
Students see concretely what they are about to aggregate across all 800 markets.

###

**Knowledge drop:** No two markets look alike. Some start at 0.5 and barely
move for weeks; others spike wildly on a news event and then stabilize. What
the aggregate analysis will reveal is whether, **on average**, markets are
closer to the truth at the end than at the beginning ŌĆö regardless of how
erratic the individual path was. The aggregate story is in the mean; the
individual story is in this plot.

### Exercise 13

Add a `time_bucket` column to your joined table. Use `case_when()` to assign
each row to one of five ordered categories based on `days_to_resolution`:

- `days_to_resolution > 90`  ŌåÆ `"90+ days"`
- `days_to_resolution > 30`  ŌåÆ `"30ŌĆō90 days"`
- `days_to_resolution > 7`   ŌåÆ `"7ŌĆō30 days"`
- `days_to_resolution >= 1`  ŌåÆ `"1ŌĆō7 days"`
- otherwise                  ŌåÆ `"Final day"`

Then convert it to an ordered factor with `factor(..., ordered = TRUE)`, levels
running from `"90+ days"` (first) to `"Final day"` (last). Count rows per
bucket. Which bucket has the most rows? Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r prediction-markets-13, echo=FALSE}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 12)
```

###

```{r prediction-markets-13-test}
#| echo: true
pm_joined |>
  count(time_bucket)
```

Expected output:

```
# A tibble: 5 x 2
  time_bucket     n
  <ord>       <int>
1 90+ days     5239
2 30ŌĆō90 days  15588
3 7ŌĆō30 days   16467
4 1ŌĆō7 days     5600
5 Final day     800
```

###

**Knowledge drop:** The 7ŌĆō30 day window has the most rows because most markets
in this dataset last 30ŌĆō60 days: a market created today and resolving in 45
days spends roughly 15 of its days in the "7ŌĆō30 days" window and 30 days in
the "30ŌĆō90 days" window. The sparse "90+ days" bucket comes from the small
fraction of long-running markets (maximum duration: 241 days). Ordered factors
preserve this natural ordering in plots ŌĆö without them, **ggplot2** would sort
the buckets alphabetically and the downward trend would disappear.

### Exercise 14

Plot a box plot of `prediction_error` by `time_bucket`. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r prediction-markets-14, echo=FALSE}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r prediction-markets-14-test}
#| echo: true
pm_joined |>
  ggplot(aes(x = time_bucket, y = prediction_error)) +
  geom_boxplot() +
  labs(
    title = "Prediction error shrinks near the deadline",
    x     = "Days until resolution",
    y     = "Prediction error (|probability ŌłÆ outcome|)"
  ) +
  theme_minimal()
```

Expected: five boxes ordered from left (90+ days: tall, centered around 0.35)
to right (Final day: short, hugging 0). Both the median and the IQR shrink
steadily toward the deadline.

###

**Knowledge drop:** The shrinking **spread** matters as much as the shrinking
**center**. A market 90 days from resolution might have any probability ŌĆö the
crowd has not locked in. On the final day, almost all markets have converged to
a tight range near 0 or 1. The remaining outliers are the genuinely surprising
outcomes: events that looked decided until they weren't. These are visible as
the high-error points above the Final day box.

### Exercise 15

Aggregate the joined table to compute `mean_error` per `time_bucket`. Assign
the result to `accuracy`. Print it. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r prediction-markets-15, echo=FALSE}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```{r prediction-markets-15-test}
#| echo: true
pm_accuracy
```

Expected output:

```
# A tibble: 5 x 3
  time_bucket mean_error n_rows
  <ord>            <dbl>  <int>
1 90+ days         0.398   5239
2 30ŌĆō90 days       0.311  15588
3 7ŌĆō30 days        0.279  16467
4 1ŌĆō7 days         0.180   5600
5 Final day        0.0901   800
```

###

**Knowledge drop:** The biggest single drop is from "1ŌĆō7 days" (0.180) to
"Final day" (0.090). This acceleration in the final week likely reflects a
burst of concrete information: election results arriving overnight, company
earnings released, game scores finalized ŌĆö all of which resolve uncertainty
that was unresolvable even a day earlier. Markets that were genuinely
undecided at 7 days often have a clear answer by the final day. The crowd
is not getting smarter ŌĆö it is getting more information.

### Exercise 16

Build the final line chart using `accuracy`. Plot `time_bucket` on the x-axis
and `mean_error` on the y-axis. Add `group = 1` in the aesthetic so
`geom_line()` connects all five points. Add `geom_line()` and `geom_point()`.
Use a line-and-point chart. Title: `"Prediction markets get smarter near the
deadline"`. Subtitle: `"Mean absolute error in Manifold Markets forecasts, by
days remaining"`. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r prediction-markets-16, echo=FALSE}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 10)
```

###

```{r prediction-markets-16-test}
#| echo: true
ggplot(pm_accuracy, aes(x = time_bucket, y = mean_error, group = 1)) +
  geom_line(linewidth = 1.2, color = "#2166ac") +
  geom_point(size = 3, color = "#2166ac") +
  labs(
    title    = "Prediction markets get smarter near the deadline",
    subtitle = "Mean absolute error in Manifold Markets forecasts, by days remaining",
    x        = "Days until resolution",
    y        = "Mean absolute error"
  ) +
  theme_minimal()
```

Expected: a clear downward-sloping line from 0.398 (90+ days) to 0.090 (Final
day), with five visible points. The mystery resolves.

###

**Knowledge drop:** The pattern here matches what forecasting researchers call
**wisdom-of-crowds convergence**: as more information arrives and participants
update their beliefs, the aggregate probability reflects reality more accurately.
Philip Tetlock's work on superforecasters shows similar patterns ŌĆö the best
forecasters update aggressively in the final days as concrete evidence arrives.
But notice that even Manifold's crowd ŌĆö ordinary users, not professional
forecasters ŌĆö achieves a 4├Ś accuracy improvement near the deadline. The
improvement comes as much from the market mechanism as from the skill of
individual bettors.

###

### Exercise 17

Stage `analysis.qmd`, commit with the message
`"Add prediction market accuracy analysis"`, and push to GitHub. In the
Terminal, run:

```
git log --oneline -3
```

Paste the output below. CP/CR.

```{r prediction-markets-17, echo=FALSE}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

Expected output (hashes will differ):

```
abc1234 Add prediction market accuracy analysis
def5678 Add crypto market price analysis
...
```

###

## Summary
###

This tutorial covered key concepts from [Chapter 16: Factors](https://r4ds.hadley.nz/factors.html), [Chapter 17: Dates and Times](https://r4ds.hadley.nz/datetimes.html), [Chapter 18: Missing Values](https://r4ds.hadley.nz/missing-values.html), [Chapter 19: Joins](https://r4ds.hadley.nz/joins.html), and [Chapter 22: Arrow](https://r4ds.hadley.nz/arrow.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine ├ćetinkaya-Rundel, and Garrett Grolemund.

You learned about working with parquet files using **[arrow](https://arrow.apache.org/docs/r/)**, joining multiple datasets with shared keys, parsing mixed-format date strings with **[lubridate](https://lubridate.tidyverse.org/)**, cleaning messy string fields with **[stringr](https://stringr.tidyverse.org/)**, and reordering factor levels for visualization with **[forcats](https://forcats.tidyverse.org/)**.

### Exercise 1

Render to ensure that everything works. Check your Live Server tab ŌĆö the resulting HTML page should be attractive, showing a clean version of your final plot.

In the R Terminal, run:

```
show_file("analysis.qmd")
```

CP/CR.

```{r summary-1}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 30)
```

###

**arrow** only understands a subset of R expressions, so before you pull data into R you may not be able to write exactly the code you normally would ŌĆö the rest has to wait until the data is in R. The **arrow** documentation lists which operations it can push down to the on-disk query.

### Exercise 2

Publish your rendered QMD to GitHub Pages. In a bash Terminal, run:

```
quarto publish gh-pages analysis.qmd
```

If the bash Terminal is still running from rendering, stop it first using `Ctrl/Cmd + C`.

Copy/paste the resulting URL below.

```{r summary-2}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 1)
```

###

Publishing the QMD creates a public URL anyone can use to view the rendered project.

### Exercise 3

Commit and push any remaining changes. Copy/paste the URL to your GitHub repo.

```{r summary-3}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

###

GitHub repositories are where people can view the code you used to create your project.

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
