---
title: Fourth Tutorial for R4DS
author: David Kane
tutorial:
  id: r4ds-4
output:
  learnr::tutorial:
    df_print: default
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: 'This tutorial for "R for Data Science" covers Chapter 16: Factors, Chapter 17: Dates and Times, Chapter 18: Missing Values, Chapter 19: Joins, and Chapter 22: Arrow. Exercises use a three-file crypto markets dataset to practice arrow, forcats, lubridate, joins, string parsing, and categorical variables.'
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(tidyverse)
library(arrow)
library(viridis)

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(out.width = '90%')
options(tutorial.exercise.timelimit = 60, tutorial.storage = "local")
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

## Introduction
###

This tutorial covers key concepts from [Chapter 16: Factors](https://r4ds.hadley.nz/factors.html), [Chapter 17: Dates and Times](https://r4ds.hadley.nz/datetimes.html), [Chapter 18: Missing Values](https://r4ds.hadley.nz/missing-values.html), [Chapter 19: Joins](https://r4ds.hadley.nz/joins.html), and [Chapter 22: Arrow](https://r4ds.hadley.nz/arrow.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund.

You will learn about working with parquet files using **[arrow](https://arrow.apache.org/docs/r/)**, joining multiple datasets with **[dplyr](https://dplyr.tidyverse.org/)**, parsing dates with **[lubridate](https://lubridate.tidyverse.org/)**, cleaning messy strings with **[stringr](https://stringr.tidyverse.org/)**, and reordering categorical variables with **[forcats](https://forcats.tidyverse.org/)**. Key concepts include opening large datasets from disk without loading them fully into memory, joining tables through shared keys, parsing mixed-format date strings, and building an ordered factor for a final line chart.

We recommend using an agentic coding tool such as [Gemini CLI](https://github.com/google-gemini/gemini-cli) or [Claude Code](https://claude.ai/code). Our instructions are written with these tools in mind. You may also use a chat-based AI, but you will need to copy/paste code and data context manually.

### Exercise 1

You should be connected to a repo named `r4ds-4`. If you are not, create one and connect to it.

Create a new file, `analysis.qmd`, with the title `"Crypto Markets"` and your name as the author. In a bash Terminal, render it:

```
quarto render analysis.qmd
```

Open `analysis.html` with Live Server (right-click it in the Explorer → **Open with Live Server**) and keep the tab open. It refreshes on every render. Going forward, we will just tell you to "Render" when we want you to take these steps.

Create a `.gitignore` with `analysis_files` followed by a blank line. Commit and push.

In the R Terminal, run:

```
show_file(".gitignore")
```

If that fails, it is probably because you have not yet loaded `library(tutorial.helpers)` in the R Terminal.

CP/CR.

```{r introduction-1}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

###

```
analysis_files
```

###

Parquet files are column-oriented — organized column-by-column like R's data frame — which leads to better performance for data analysis than row-oriented CSV files. They also support efficient compression, making them much smaller than equivalent CSV files.

### Exercise 2

In your QMD, put the following in a new code chunk to load all required libraries:

```
suppressPackageStartupMessages({
  library(tidyverse)
  library(arrow)
  library(viridis)
})
```

Also add the following to the YAML header to remove all code echoes from the HTML:

```
execute:
  echo: false
```

Render. In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r introduction-2}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 6)
```

###

<pre><code>suppressPackageStartupMessages({
  library(tidyverse)
  library(arrow)
  library(viridis)
})
</code></pre>

###

Like CSV, parquet is used for rectangular data, but it is a custom binary format designed for big data — not a text format you can read in any editor. Both **arrow** and **dbplyr** provide **dplyr** backends, so you can use the same **dplyr** code whether your data is in a parquet file or a database.

### Exercise 3

Create a `data` directory at the top level of the `r4ds-4` repo. In the bash Terminal, run:

```
ls
```

CP/CR.

```{r introduction-3}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 5)
```

###

```
analysis.html  analysis.qmd  analysis_files  data
```

###

Parquet relies on [efficient encodings](https://parquet.apache.org/docs/file-format/data-pages/encodings/) to keep file size down and supports file compression. This helps make parquet files fast because there is less data to move from disk to memory.


## Crypto markets
###

Bitcoin launched the cryptocurrency market in 2009. By 2013, it represented over 90% of all crypto by market value. But by 2021, Bitcoin's share had fallen to around 40%. Where did the other 60% go?

In this section you will join three parquet files — daily prices, coin metadata, and category labels — to build a line chart showing how each category's share of the total crypto market evolved from 2018 to 2024. Along the way you will use **[arrow](https://arrow.apache.org/docs/r/)** to open files lazily, **[lubridate](https://lubridate.tidyverse.org/)** to parse dates, **[stringr](https://stringr.tidyverse.org/)** to clean messy fields, and **[forcats](https://forcats.tidyverse.org/)** to reorder categories for the final visualization.

### Exercise 1

Download all three parquet files into your `data/` directory. The stable URLs follow the pattern `https://github.com/PPBDS/misc.tutorials/raw/refs/heads/main/inst/extdata/r4ds-4/<file>` for each of `daily_prices.parquet`, `coin_metadata.parquet`, and `categories.parquet`.

In the bash Terminal, run:

```
ls data
```

CP/CR.

```{r crypto-markets-1}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

###

```
categories.parquet    coin_metadata.parquet    daily_prices.parquet
```

###

Parquet files are column-oriented and compressed. The three files for this section total just over 2 MB — small enough to keep in a CRAN package — but `daily_prices.parquet` holds 59,436 rows of daily price and market cap data for 22 cryptocurrencies spanning 2018 to 2026.

### Exercise 2

Open `data/daily_prices.parquet` with `open_dataset()`. Don't assign it or collect it — just print the result. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r crypto-markets-2}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

###

```{r crypto-markets-2-test}
#| echo: true
open_dataset("../../extdata/r4ds-4/daily_prices.parquet")
```

###

The result is not a tibble — it is an `ArrowDataset`. Arrow is showing you the schema (column names and types) without loading any rows. No data has moved from disk to memory yet. `date_raw` is a character column, not a Date — an intentional design choice so that you parse it explicitly in a later exercise.

### Exercise 3

Chain a `filter()` and a `select()` onto the `open_dataset()` call, then `collect()`. Filter to years 2020 and later, and select `date_raw`, `coin_id`, and `market_cap_usd`. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r crypto-markets-3}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 4)
```

###

```{r crypto-markets-3-test}
#| echo: true
open_dataset("../../extdata/r4ds-4/daily_prices.parquet") |>
  filter(as.integer(substr(date_raw, 1, 4)) >= 2020) |>
  select(date_raw, coin_id, market_cap_usd) |>
  collect()
```

###

`filter()` and `select()` on an Arrow dataset execute on disk before any rows transfer to R. Only `collect()` moves data into memory. Arrow's expression engine supports most base R string functions like `substr()` but does not support `lubridate::year()` in lazy evaluation — which is why the filter uses `as.integer(substr(date_raw, 1, 4))`. For this dataset, filtering to 2020 and later reduces the transferred rows by roughly 30%.

### Exercise 4

Collect all three parquet files into tibbles. Assign them to `prices`, `metadata`, and `categories`. Print `metadata`. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r crypto-markets-4}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 5)
```

###

```{r crypto-markets-4-test}
#| echo: true
prices     <- open_dataset("../../extdata/r4ds-4/daily_prices.parquet") |> collect()
metadata   <- open_dataset("../../extdata/r4ds-4/coin_metadata.parquet") |> collect()
categories <- open_dataset("../../extdata/r4ds-4/categories.parquet") |> collect()
metadata
```

###

"January 3, 2009" is when Satoshi Nakamoto mined the Bitcoin genesis block — the date that launched the entire cryptocurrency industry. The two date formats in `launch_date` ("January 3, 2009" and "2015-07-30") reflect a common real-world data quality issue: records pulled from different sources at different times use different date conventions. **lubridate** provides `mdy()` and `ymd()` to handle both.

### Exercise 5

Left-join `prices` to `metadata` by `coin_id`, then left-join the result to `categories` by `category_id`. Both tables have a `description` column — drop the one from `categories` before joining (it is not needed). Assign the result to `crypto`. Print the first 5 rows. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r crypto-markets-5}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 5)
```

###

```{r crypto-markets-5-test}
#| echo: true
crypto <- prices |>
  left_join(metadata, by = "coin_id") |>
  left_join(select(categories, -description), by = "category_id")
head(crypto, 5)
```

###

A left join keeps every row from the left table and adds matching columns from the right. Here, each of the 59,436 daily observations in `prices` picks up its coin's name, category, launch date, and description from `metadata`, and its category display name from `categories`. Notice that `category_id` was not in `daily_prices.parquet` — it came from the first join. That is why two joins were needed: prices → metadata adds `category_id`, then metadata → categories adds the human-readable category name.

### Exercise 6

`market_cap_str` contains values like `"$1.24T"`, `"$450.2B"`, `"$37.5M"`, and some `NA`. Add a new column `market_cap_parsed` to `crypto` that extracts the numeric value: detect the suffix letter, extract the number with `parse_number()`, and multiply by the appropriate scale (1e12 for T, 1e9 for B, etc.). Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r crypto-markets-6}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 8)
```

###

```{r crypto-markets-6-test}
#| echo: true
crypto <- crypto |>
  mutate(
    market_cap_parsed = case_when(
      str_detect(market_cap_str, "T") ~ parse_number(market_cap_str) * 1e12,
      str_detect(market_cap_str, "B") ~ parse_number(market_cap_str) * 1e9,
      str_detect(market_cap_str, "M") ~ parse_number(market_cap_str) * 1e6,
      str_detect(market_cap_str, "K") ~ parse_number(market_cap_str) * 1e3,
      !is.na(market_cap_str) ~ parse_number(market_cap_str),
      TRUE ~ NA_real_
    )
  )
```

###

`parse_number()` strips non-numeric characters like `$` but cannot handle a scale suffix (`T`, `B`, `M`). The `case_when()` + `parse_number()` pattern shown here is the standard approach for parsing formatted financial strings in R. About 5% of `market_cap_str` values are `NA` — simulating the kind of gap that appears when a dashboard exports data with occasional rendering failures. The clean numeric column `market_cap_usd` is already in the dataset; `market_cap_parsed` exists solely to practice this parsing pattern.

### Exercise 7

Parse dates. In `metadata`, detect which `launch_date` values use the long-form "Month Day, Year" format and parse them accordingly — the remaining values use "YYYY-MM-DD". Store the parsed result in a new column `launch_date_parsed`. Also parse `date_raw` in `crypto` into a proper date column called `date`. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r crypto-markets-7}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 8)
```

###

```{r crypto-markets-7-test}
#| echo: true
metadata <- metadata |>
  mutate(
    launch_date_parsed = case_when(
      str_detect(launch_date, ",") ~ mdy(launch_date),
      TRUE ~ ymd(launch_date)
    )
  )

crypto <- crypto |>
  mutate(date = ymd(date_raw))
```

###

Lubridate's `mdy()` and `ymd()` are forgiving about separators but strict about field order. When a dataset has two different date formats, conditional parsing with `case_when()` is safer than hoping one parser handles both. Bitcoin's `launch_date` parses to 2009-01-03; Ethereum's parses to 2015-07-30 — both correct despite looking completely different as strings.

### Exercise 8

Clean the `description` column in `crypto` by replacing embedded newlines (`\n`) with a space. Confirm the fix by filtering to `btc` and `eth` and showing their distinct `coin_id` and `description` values. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r crypto-markets-8}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 6)
```

###

```{r crypto-markets-8-test}
#| echo: true
crypto <- crypto |>
  mutate(description = str_replace_all(description, "\\n", " "))

crypto |>
  filter(coin_id %in% c("btc", "eth")) |>
  distinct(coin_id, description)
```

###

Embedded newlines appear when data is scraped from a web page or copied from a document that used visual line breaks. They silently break `str_detect()`, `str_length()`, and CSV export — the text looks fine in a print statement but causes unexpected behavior downstream. `str_replace_all()` with `"\\n"` (the regex for a literal newline) replaces every occurrence; the double backslash is required because R parses the first `\` as an escape character before the regex engine sees the string.

### Exercise 9

Add an `era` column to `crypto` using `case_when()` on `year(date)`: `"Bitcoin Era"` (2016 and earlier), `"ICO/Ethereum Era"` (2017–2019), `"DeFi Summer"` (2020–2021), `"Maturation Era"` (2022 onward). Convert it to an ordered factor with levels in chronological order. Count rows per era. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r crypto-markets-9}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 10)
```

###

```{r crypto-markets-9-test}
#| echo: true
crypto <- crypto |>
  mutate(
    era = factor(
      case_when(
        year(date) <= 2016 ~ "Bitcoin Era",
        year(date) <= 2019 ~ "ICO/Ethereum Era",
        year(date) <= 2021 ~ "DeFi Summer",
        TRUE ~ "Maturation Era"
      ),
      levels = c("Bitcoin Era", "ICO/Ethereum Era", "DeFi Summer", "Maturation Era")
    )
  )

count(crypto, era)
```

###

The ICO era (2017–2018) saw hundreds of new coins raise billions through initial coin offerings — mostly on the Ethereum network. Most of those projects failed. DeFi Summer (2020) was different: the protocols that launched — Uniswap, Aave, Compound, MakerDAO — created real financial primitives that still process billions of dollars today. The era labels here are interpretive, not standardized; different analysts draw the boundaries differently. What matters for the tutorial is that `factor()` with explicit `levels` gives you control over the order ggplot2 uses for color legends and facet panels.

### Exercise 10

The `crypto` chunk is getting long. Clean it up: keep only the code that defines `crypto` (the data loading, joins, and all the mutations), and remove the `count(crypto, era)` line at the bottom.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r crypto-markets-10}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 20)
```

###

<pre><code>prices     <- open_dataset("data/daily_prices.parquet") |> collect()
metadata   <- open_dataset("data/coin_metadata.parquet") |> collect()
categories <- open_dataset("data/categories.parquet") |> collect()

metadata <- metadata |>
  mutate(
    launch_date_parsed = case_when(
      str_detect(launch_date, ",") ~ mdy(launch_date),
      TRUE ~ ymd(launch_date)
    )
  )

crypto <- prices |>
  left_join(metadata, by = "coin_id") |>
  left_join(select(categories, -description), by = "category_id") |>
  mutate(
    market_cap_parsed = case_when(
      str_detect(market_cap_str, "T") ~ parse_number(market_cap_str) * 1e12,
      str_detect(market_cap_str, "B") ~ parse_number(market_cap_str) * 1e9,
      str_detect(market_cap_str, "M") ~ parse_number(market_cap_str) * 1e6,
      str_detect(market_cap_str, "K") ~ parse_number(market_cap_str) * 1e3,
      !is.na(market_cap_str) ~ parse_number(market_cap_str),
      TRUE ~ NA_real_
    ),
    date = ymd(date_raw),
    description = str_replace_all(description, "\\n", " "),
    era = factor(
      case_when(
        year(date) <= 2016 ~ "Bitcoin Era",
        year(date) <= 2019 ~ "ICO/Ethereum Era",
        year(date) <= 2021 ~ "DeFi Summer",
        TRUE ~ "Maturation Era"
      ),
      levels = c("Bitcoin Era", "ICO/Ethereum Era", "DeFi Summer", "Maturation Era")
    )
  )
</code></pre>

###

The fully joined and cleaned `crypto` tibble has one row per coin per day — 59,436 rows covering 22 coins from 2018 through 2026. Bitcoin's dominance, Ethereum's rise, stablecoin growth, and the meme-coin spike of 2021 are all latent in this table, waiting to be drawn out by the grouping and plotting exercises ahead.

### Exercise 11

Add `#| cache: true` to the chunk. Render. This first render with caching will take a moment; subsequent renders will load from disk.

In the bash Terminal, run:

```
ls
```

CP/CR.

```{r crypto-markets-11}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 5)
```

###

```
analysis.html  analysis.qmd  analysis_cache  analysis_files  data
```

###

Stablecoins like USDT grew from roughly $4B market cap in early 2019 to over $80B by 2022 — not because the token's price rose, but because more dollars flowed into the crypto ecosystem seeking a stable store of value. USDT, USDC, and DAI all maintained their $1 peg throughout; their rising market cap reflects real-dollar demand for crypto-denominated stability, not speculation.

### Exercise 12

Add `analysis_cache` to your `.gitignore`. The Source Control change count in VS Code jumped when the cache directory appeared; adding it to `.gitignore` drops it back. Render.

In the R Terminal, run:

```
show_file(".gitignore")
```

CP/CR.

```{r crypto-markets-12}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 4)
```

###

```
analysis_files
analysis_cache
```

###

The FTX exchange token FTT had a market cap of roughly $7.25B on November 8, 2022. By November 9 it had fallen to $1.82B — a 75% single-day drop. FTX, one of the largest crypto exchanges, filed for bankruptcy three days later. The collapse is visible as a near-vertical line in the data; it is one of the clearest single-event signals in the entire dataset.

### Exercise 13

In a new code chunk, compute the total market cap per day across all coins in `crypto` (sum `market_cap_usd` by `date`). Plot it as a line chart with date on the x-axis and total market cap in trillions on the y-axis. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r crypto-markets-13}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 8)
```

###

```{r crypto-markets-13-test}
#| echo: true
crypto |>
  group_by(date) |>
  summarise(total_mcap = sum(market_cap_usd, na.rm = TRUE)) |>
  ggplot(aes(x = date, y = total_mcap / 1e12)) +
  geom_line() +
  labs(x = NULL, y = "Total market cap (trillions USD)",
       title = "Total crypto market capitalization, 2020–2026")
```

###

The total crypto market cap grew from roughly $150B in early 2019 to over $2.5T at its November 2021 peak — a 16× increase in under three years. It then fell sharply through 2022 (the "crypto winter"), driven by the Luna/UST collapse in May and FTX in November. By 2024 it had recovered above $2T. This context is essential for the next chart: Bitcoin's *share* fell during the bull market while its *absolute value* rose.

### Exercise 14

Compute daily market cap per category by summing `market_cap_usd` across all coins in each `category_name`, per `date`. Plot a multi-line chart with date on the x-axis and category market cap on the y-axis, colored by `category_name`. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r crypto-markets-14}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 8)
```

###

```{r crypto-markets-14-test}
#| echo: true
crypto |>
  group_by(date, category_name) |>
  summarise(cat_mcap = sum(market_cap_usd, na.rm = TRUE), .groups = "drop") |>
  ggplot(aes(x = date, y = cat_mcap / 1e12, color = category_name)) +
  geom_line() +
  labs(x = NULL, y = "Market cap (trillions USD)", color = NULL,
       title = "Crypto market cap by category, 2020–2026")
```

###

This chart is hard to read because Bitcoin dominates the y-axis. Its absolute market cap dwarfs all other categories, compressing the other lines near zero. When one category dominates the absolute scale, relative share — each category as a percentage of the total — is the standard way to make the other trajectories visible.

### Exercise 15

Compute each category's share as a percentage of the total daily market cap. Plot the same multi-line chart with share (%) on the y-axis. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r crypto-markets-15}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 10)
```

###

```{r crypto-markets-15-test}
#| echo: true
crypto |>
  group_by(date, category_name) |>
  summarise(cat_mcap = sum(market_cap_usd, na.rm = TRUE), .groups = "drop") |>
  group_by(date) |>
  mutate(total_mcap = sum(cat_mcap),
         share_pct = cat_mcap / total_mcap * 100) |>
  ungroup() |>
  ggplot(aes(x = date, y = share_pct, color = category_name)) +
  geom_line() +
  labs(x = NULL, y = "Market cap share (%)", color = NULL,
       title = "Crypto market cap share by category, 2020–2026")
```

###

Now the story is visible. Bitcoin's line starts above 50% and shows a declining arc through 2021 before recovering. Stablecoins grow steadily — especially through 2022 when investors fled volatile assets. DeFi protocols spike in 2020–2021 then partially retreat. Meme coins (DOGE) cause a sharp spike in May 2021 — Elon Musk tweets drove DOGE's market cap briefly above $80B — then almost entirely collapse within weeks. The diversity of trajectories across categories is exactly the answer to the guiding mystery.

### Exercise 16

Reorder the `category_name` factor so that categories with the highest median share appear first in the legend. Apply a colorblind-safe color palette. Add a proper title, subtitle, axis labels, and a caption crediting the source. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r crypto-markets-16}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 14)
```

###

```{r crypto-markets-16-test}
#| echo: true
crypto |>
  group_by(date, category_name) |>
  summarise(cat_mcap = sum(market_cap_usd, na.rm = TRUE), .groups = "drop") |>
  group_by(date) |>
  mutate(total_mcap = sum(cat_mcap),
         share_pct = cat_mcap / total_mcap * 100) |>
  ungroup() |>
  mutate(category_name = fct_reorder(category_name, share_pct, .fun = median, .desc = TRUE)) |>
  ggplot(aes(x = date, y = share_pct, color = category_name)) +
  geom_line(linewidth = 0.8) +
  scale_color_viridis_d(option = "plasma", end = 0.9) +
  labs(x = NULL, y = "Share of total market cap (%)", color = NULL,
       title = "Who ate Bitcoin's lunch?",
       subtitle = "Crypto market cap share by category, 2020–2026",
       caption = "Source: Coin Metrics Community Data (CC BY 4.0)")
```

###

`fct_reorder()` sorts factor levels by a summary statistic — here, the median share across all dates. This puts Bitcoin at the top of the legend (it has the highest median), which aligns with its line being highest in the chart. When legend order matches visual order, readers scan less. `viridis` and `viridis_d` are colorblind-safe palettes that also print clearly in grayscale — a good default for any chart that may be viewed in multiple contexts.

### Exercise 17

Add annotations to highlight two historical events: the Luna/UST crash (approximately May 7–12, 2022) and the FTX collapse (approximately November 8–11, 2022). Use `annotate("rect", ...)` or `geom_vline()` — your choice. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r crypto-markets-17}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 12)
```

###

```{r crypto-markets-17-test}
#| echo: true
crypto |>
  group_by(date, category_name) |>
  summarise(cat_mcap = sum(market_cap_usd, na.rm = TRUE), .groups = "drop") |>
  group_by(date) |>
  mutate(total_mcap = sum(cat_mcap),
         share_pct = cat_mcap / total_mcap * 100) |>
  ungroup() |>
  mutate(category_name = fct_reorder(category_name, share_pct, .fun = median, .desc = TRUE)) |>
  ggplot(aes(x = date, y = share_pct, color = category_name)) +
  geom_line(linewidth = 0.8) +
  scale_color_viridis_d(option = "plasma", end = 0.9) +
  annotate("rect", xmin = as.Date("2022-05-07"), xmax = as.Date("2022-05-12"),
           ymin = -Inf, ymax = Inf, alpha = 0.15, fill = "red") +
  annotate("rect", xmin = as.Date("2022-11-08"), xmax = as.Date("2022-11-11"),
           ymin = -Inf, ymax = Inf, alpha = 0.15, fill = "red") +
  labs(x = NULL, y = "Share of total market cap (%)", color = NULL,
       title = "Who ate Bitcoin's lunch?",
       subtitle = "Crypto market cap share by category, 2020–2026",
       caption = "Source: Coin Metrics Community Data (CC BY 4.0)")
```

###

Terra's algorithmic stablecoin UST maintained its $1 peg through a supply-demand mechanism rather than dollar reserves. In May 2022, a large sell-off broke the mechanism; both UST and its sister token Luna lost almost all their value within 72 hours — the largest single wealth-destruction event in crypto history at that time. Fiat-backed stablecoins like USDT and USDC held their pegs throughout; their divergence from UST is visible in the stablecoin category line, which dips slightly in May 2022 then resumes growth.

### Exercise 18

Commit `analysis.qmd` with the message `"Crypto market composition analysis"` and push to GitHub.

```{r crypto-markets-18}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

###

The 22 coins in this dataset represent the major categories but exclude thousands of smaller coins — category shares here are approximations, not exact measures of the full crypto market. By 2026, the top 10 coins by market cap still accounted for roughly 80% of total market value, so the broad patterns hold even though the tail is missing.

## Summary
###

This tutorial covered key concepts from [Chapter 16: Factors](https://r4ds.hadley.nz/factors.html), [Chapter 17: Dates and Times](https://r4ds.hadley.nz/datetimes.html), [Chapter 18: Missing Values](https://r4ds.hadley.nz/missing-values.html), [Chapter 19: Joins](https://r4ds.hadley.nz/joins.html), and [Chapter 22: Arrow](https://r4ds.hadley.nz/arrow.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund.

You learned about working with parquet files using **[arrow](https://arrow.apache.org/docs/r/)**, joining multiple datasets with shared keys, parsing mixed-format date strings with **[lubridate](https://lubridate.tidyverse.org/)**, cleaning messy string fields with **[stringr](https://stringr.tidyverse.org/)**, and reordering factor levels for visualization with **[forcats](https://forcats.tidyverse.org/)**.

### Exercise 1

Render to ensure that everything works. Check your Live Server tab — the resulting HTML page should be attractive, showing a clean version of your final plot.

In the R Terminal, run:

```
show_file("analysis.qmd")
```

CP/CR.

```{r summary-1}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 30)
```

###

**arrow** only understands a subset of R expressions, so before you pull data into R you may not be able to write exactly the code you normally would — the rest has to wait until the data is in R. The **arrow** documentation lists which operations it can push down to the on-disk query.

### Exercise 2

Publish your rendered QMD to GitHub Pages. In a bash Terminal, run:

```
quarto publish gh-pages analysis.qmd
```

If the bash Terminal is still running from rendering, stop it first using `Ctrl/Cmd + C`.

Copy/paste the resulting URL below.

```{r summary-2}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 1)
```

###

Publishing the QMD creates a public URL anyone can use to view the rendered project.

### Exercise 3

Commit and push any remaining changes. Copy/paste the URL to your GitHub repo.

```{r summary-3}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

###

GitHub repositories are where people can view the code you used to create your project.

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
