---
title: Tidymodels
author: Harshil Kurapati
tutorial:
  id: tidymodels
output:
  learnr::tutorial:
    df_print: default
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: "Learn the tidymodels framework by fitting linear, logistic, multinomial, and ordinal regression models to Titanic passenger data."
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(tidymodels)
library(tidyverse)
library(broom)
library(ordered)

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(out.width = '90%')
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
options(tutorial.exercise.timelimit = 60,
        tutorial.storage = "local")

titanic <- read_csv("data/titanic.csv", show_col_types = FALSE) |>
  mutate(class_sex = paste(PClass, Sex, sep = "-"))

build_equation <- function(coefs, outcome) {
  has_intercept <- "(Intercept)" %in% coefs$term
  other <- coefs[coefs$term != "(Intercept)", ]
  parts <- purrr::map2_chr(other$term, round(other$estimate, 2), function(term, est) {
    sign <- if (est >= 0) "+" else "-"
    paste0(sign, " ", abs(est), " \\times \\text{", term, "}")
  })
  lead <- if (has_intercept) as.character(round(coefs$estimate[coefs$term == "(Intercept)"], 2)) else NULL
  rhs <- paste(c(lead, parts), collapse = " ")
  paste0("$$\\text{", outcome, "} = ", rhs, "$$")
}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

## Introduction
###

This tutorial introduces the **[tidymodels](https://www.tidymodels.org/)** framework for statistical modeling, following [*Tidy Modeling with R*](https://www.tmwr.org/) and the [tidymodels Get Started guide](https://www.tidymodels.org/start/). You will fit four different kinds of regression models to Titanic passenger data, learning the same specify-engine-fit pattern that tidymodels uses for every model type.

Every dataset is a shadow of some real-world process — the **data generating mechanism**, or DGM. This tutorial's four sections all propose a **structure** for that mechanism (what kind of outcome are you predicting?), pick a **method** for estimating it (`set_engine()`), then **fit** it to real data. Structure → Method → Fitting.

### Exercise 1

You should be connected to a repo named `tidymodels`. If you are not, create one and connect to it.

Create a new file, `analysis.qmd`, with the title `"Tidymodels"` and your name as the author. In a bash Terminal, render it:

```
quarto render analysis.qmd
```

Open `analysis.html` with Live Server (right-click it in the Explorer → **Open with Live Server**) and keep the tab open. It refreshes on every render. Going forward, we will just tell you to "Render" when we want you to take these steps.

Create a `.gitignore` with `analysis_files` followed by a blank line. Commit and push.

In the R Terminal, run:

```
show_file(".gitignore")
```

If that fails, it is probably because you have not yet loaded `library(tutorial.helpers)` in the R Terminal.

CP/CR.

```{r introduction-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

```
analysis_files
```

###

**[tidymodels](https://www.tidymodels.org/)** is a meta-package, much like the tidyverse — it attaches a whole family of modeling packages (**parsnip**, **recipes**, **rsample**, and others) with one `library()` call.

### Exercise 2

Create a `data` directory at the top level of the `tidymodels` repo, then download `titanic.csv` into it from this URL:

```
https://github.com/PPBDS/misc.tutorials/raw/refs/heads/main/inst/tutorials/tidymodels/data/titanic.csv
```

In the bash Terminal, run `ls data`. CP/CR.

```{r introduction-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

```
titanic.csv
```

###

This is the classic Titanic passenger dataset, compiled by Thomas Cason — 756 passengers with a known age, ticket class, sex, and survival outcome. You will use `Age`, `PClass`, `Sex`, and `Survived` across all four sections ahead, plus a derived `class_sex` column for the multinomial section.

### Exercise 3

In your QMD, add a setup chunk that loads **tidymodels**, along with **tidyverse**, **broom**, and the **ordered** package (an extension you will need later for ordinal regression), reads `data/titanic.csv` into a tibble named `titanic`, and adds a `class_sex` column that pastes `PClass` and `Sex` together with a `"-"` separator. Add `#| message: false` so package startup messages stay out of the rendered page, and add `execute: echo: false` to the YAML header to hide code.

Render. In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r introduction-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

<pre><code>#| message: false
library(tidymodels)
library(tidyverse)
library(broom)
library(ordered)

titanic <- read_csv("data/titanic.csv", show_col_types = FALSE) |>
  mutate(class_sex = paste(PClass, Sex, sep = "-"))
</code></pre>

###

**[parsnip](https://parsnip.tidymodels.org/)** turns a proposed structure and method into an actual model. Every section ahead names the structure (which `_reg()` function), names the method (`set_engine()`), then `fit()`s it — the same three moves, four times.

### Exercise 4

Commit and push your work with a message like "Initial project setup with Titanic data". Copy-paste the output of `git log --oneline -3` into the answer box.

```{r introduction-4}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```
Initial project setup with Titanic data
Add tidymodels tutorial data
Initial commit
```

###

Committing before any model fitting creates a clean baseline. Every model you build next starts from this same setup chunk, so a stable starting point makes it easy to recover if a `fit()` call throws an error.

## Linear regression
###

`linear_reg()` proposes the **structure**: our outcome, `Age`, is a number, so we're guessing at a mechanism that produces numbers. `set_engine("lm")` picks the **method** for estimating it. Only `fit()` — the third step — actually touches data.

### Exercise 1

In a new code chunk, specify a linear regression with `linear_reg()`, and set its engine to `"lm"` with `set_engine()`. Render, then run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r linear-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

```{r linear-1-test}
#| echo: true
linear_reg() |>
  set_engine("lm")
```

###

Structure and method are chosen before any data is touched — `fit()`, the step still ahead, is what actually happens to the data.

### Exercise 2

`fit()` it to predict `Age` from `PClass + Sex` using `titanic`, storing the fitted model as `mod_linear`. Print `mod_linear`. Render, then run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r linear-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r linear-2-test}
#| echo: true
mod_linear <- linear_reg() |>
  set_engine("lm") |>
  fit(Age ~ PClass + Sex, data = titanic)

mod_linear
```

###

`fit()` is where the mechanism meets real data — this is what a fitted DGM actually is. 2nd class passengers are predicted to be about 11.5 years younger than 1st class, and 3rd class about 14.8 years younger, holding sex fixed.

### Exercise 3

Replace the printed `mod_linear` with a call to `tidy(mod_linear)` to see the coefficients as a tidy tibble. Render, then run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r linear-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r linear-3-test}
#| echo: true
tidy(mod_linear)
```

###

`tidy()` reads results back out of the fitted DGM — it doesn't change what the machine learned, just how you view it.

### Exercise 4

Add a second fitted model, `mod_linear_glm`, that specifies the same `linear_reg()` but sets the engine to `"glm"` instead of `"lm"`, fit to the same formula and data. Add `tidy(mod_linear_glm)` below it so you can compare coefficients. Render, then run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r linear-4}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r linear-4-test}
#| echo: true
mod_linear_glm <- linear_reg() |>
  set_engine("glm") |>
  fit(Age ~ PClass + Sex, data = titanic)

print(tidy(mod_linear))
tidy(mod_linear_glm)
```

###

Only a chunk's last line displays automatically, so `print()` shows `mod_linear`'s coefficients too. `"lm"` and `"glm"` are two different **methods** for estimating the same structure — here they agree exactly, but that won't always be true.

### Exercise 5

For every model in this tutorial, you will now show its equation with the actual estimated numbers plugged in. At the top of your last chunk, above `mod_linear_glm`, add this function:

```
build_equation <- function(coefs, outcome) {
  has_intercept <- "(Intercept)" %in% coefs$term
  other <- coefs[coefs$term != "(Intercept)", ]
  parts <- purrr::map2_chr(other$term, round(other$estimate, 2), function(term, est) {
    sign <- if (est >= 0) "+" else "-"
    paste0(sign, " ", abs(est), " \\times \\text{", term, "}")
  })
  lead <- if (has_intercept) as.character(round(coefs$estimate[coefs$term == "(Intercept)"], 2)) else NULL
  rhs <- paste(c(lead, parts), collapse = " ")
  paste0("$$\\text{", outcome, "} = ", rhs, "$$")
}
```

Then replace the chunk's two `tidy()` lines with these two lines, and add `#| results: asis` to the top of the chunk:

```
cat(build_equation(tidy(mod_linear), "Age"))
cat(build_equation(tidy(mod_linear_glm), "Age (glm)"))
```

Render, then run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r linear-5}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 20)
```

###

```{r linear-5-test}
#| echo: true
#| results: asis
build_equation <- function(coefs, outcome) {
  has_intercept <- "(Intercept)" %in% coefs$term
  other <- coefs[coefs$term != "(Intercept)", ]
  parts <- purrr::map2_chr(other$term, round(other$estimate, 2), function(term, est) {
    sign <- if (est >= 0) "+" else "-"
    paste0(sign, " ", abs(est), " \\times \\text{", term, "}")
  })
  lead <- if (has_intercept) as.character(round(coefs$estimate[coefs$term == "(Intercept)"], 2)) else NULL
  rhs <- paste(c(lead, parts), collapse = " ")
  paste0("$$\\text{", outcome, "} = ", rhs, "$$")
}

mod_linear_glm <- linear_reg() |>
  set_engine("glm") |>
  fit(Age ~ PClass + Sex, data = titanic)

cat(build_equation(tidy(mod_linear), "Age"))
cat(build_equation(tidy(mod_linear_glm), "Age (glm)"))
```

###

Both equations come out identical here — "lm" and "glm" are two different methods for estimating the same structure, but that won't always be true.


### Exercise 6

Commit and push your work with a message like "Add linear regression model". Copy-paste the output of `git log --oneline -3` into the answer box.

```{r linear-6}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```
Add linear regression model
Initial project setup with Titanic data
Add tidymodels tutorial data
```

###

Two predictors barely uses this dataset's richness, but the specify-engine-fit pattern stays the same no matter how many predictors a formula has.

## Logistic regression
###

Not every outcome is a number. `Survived` is a category — 1 or 0, a genuine yes-or-no about whether a passenger lived.

### Exercise 1

Add a new chunk. Specify a logistic regression with `logistic_reg()`, set its engine to `"glm"`, and `fit()` it to predict `factor(Survived, levels = c(0, 1))` from `Age + PClass + Sex` using `titanic`, storing the result as `mod_logistic`. Show `tidy(mod_logistic)`. Render, then run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r logistic-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r logistic-1-test}
#| echo: true
mod_logistic <- logistic_reg() |>
  set_engine("glm") |>
  fit(factor(Survived, levels = c(0, 1)) ~ Age + PClass + Sex, data = titanic)

tidy(mod_logistic)
```

###

`logistic_reg()` proposes a different **structure** than `linear_reg()` did — a category instead of a number, so its coefficient is measured in log-odds, not the outcome's original units.

### Exercise 2

Replace `tidy(mod_logistic)` with the equation display, as you did for the linear model. Add `#| results: asis` to the top of the chunk:

```
cat(build_equation(tidy(mod_logistic), "logit(P(Survived))"))
```

Render, then run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r logistic-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r logistic-2-test}
#| echo: true
#| results: asis
cat(build_equation(tidy(mod_logistic), "logit(P(Survived))"))
```

###

Every coefficient here is negative — older age, lower class, and being male each independently pushed the log-odds of survival down. `build_equation()` doesn't know or care what structure produced these numbers; it just formats whatever `tidy()` gives it.

### Exercise 3

Clean up your chunk so it holds only what is needed to rebuild `mod_logistic`: remove the `cat(build_equation(...))` line, keeping just the `mod_logistic` definition. Render, then run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r logistic-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

<pre><code>mod_logistic <- logistic_reg() |>
  set_engine("glm") |>
  fit(factor(Survived, levels = c(0, 1)) ~ Age + PClass + Sex, data = titanic)
</code></pre>

###

This chunk is now nothing but the one definition needed to reproduce `mod_logistic` — no printed output, no exploration. That is exactly the shape a chunk should have right before you cache it.

### Exercise 4

Add `#| cache: true` to this chunk. Render. In the bash Terminal, run `ls`. CP/CR.

```{r logistic-4}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

```
analysis.html
analysis.qmd
analysis_cache
analysis_files
data
```

###

Rendering with `#| cache: true` created an `analysis_cache` directory alongside `analysis.qmd` — the file-based evidence that this chunk's result is now saved to disk. The next render will skip re-fitting `mod_logistic` entirely and load it from the cache, as long as the chunk's code has not changed.

### Exercise 5

Your Explorer's Source Control panel is now showing one more pending change: the new `analysis_cache` directory. Add `*_cache/` and `/.quarto/` to your `.gitignore`. In the R Terminal, run:

```
show_file(".gitignore")
```

CP/CR.

```{r logistic-5}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```
analysis_files
*_cache/
/.quarto/
```

###

`analysis_cache` and `.quarto` are regenerated by every render, so ignoring them keeps your repo's history to real changes only.

### Exercise 6

Commit and push your work with a message like "Add logistic regression model with caching". Copy-paste the output of `git log --oneline -3` into the answer box.

```{r logistic-6}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```
Add logistic regression model with caching
Add linear regression model
Initial project setup with Titanic data
```

###

A model that fits this exact dataset well may not generalize — that risk is called **overfitting**. Guarding against it properly means splitting data into training and test sets, which is beyond this introduction.

## Multinomial regression
###

`Survived` had two categories. `class_sex` has six — `1st-female`, `1st-male`, and so on. `multinom_reg()` proposes a mechanism for an outcome with three or more unordered categories.

### Exercise 1

Add a new chunk. Count how many passengers fall into each `class_sex` category, sorted from most to least common. Render, then run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r multinomial-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

```{r multinomial-1-test}
#| echo: true
count(titanic, class_sex, sort = TRUE)
```

###

Men outnumber women in every class, and 3rd class alone (318 passengers) holds more people than 1st and 2nd combined.

### Exercise 2

In place of the count, specify a multinomial regression with `multinom_reg()`, set its engine to `"nnet"`, and `fit()` it to predict `factor(class_sex)` from `Age + Survived` using `titanic`, storing the result as `mod_multinomial`. Print `mod_multinomial`. Render, then run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r multinomial-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r multinomial-2-test}
#| echo: true
mod_multinomial <- multinom_reg() |>
  set_engine("nnet") |>
  fit(factor(class_sex) ~ Age + Survived, data = titanic)

mod_multinomial
```

###

`class_sex` has six unordered categories, not two — a more complex structure, but the same three steps. Notice there is no single correct way to define this outcome, either: we chose to combine class and sex, but other groupings could define different categories just as validly. Choosing a structure is itself a judgment call.

### Exercise 3

Replace the printed `mod_multinomial` with `tidy(mod_multinomial)`. Render, then run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r multinomial-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r multinomial-3-test}
#| echo: true
tidy(mod_multinomial)
```

###

The `y.level` column names which category each row belongs to, relative to the reference level `1st-female` (alphabetically first) — still the same `tidy()`, just three coefficients repeated for each of five comparisons.

### Exercise 4

The model actually estimated five separate comparisons — one for each non-reference category against `1st-female`. Show the equation for just the `3rd-male` comparison. Replace `tidy(mod_multinomial)` with these two lines, and add `#| results: asis` to the top of the chunk:

```
mod_multinomial_coefs <- tidy(mod_multinomial) |> filter(y.level == "3rd-male")
cat(build_equation(mod_multinomial_coefs, "log-odds(3rd-male vs. 1st-female)"))
```

Render, then run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r multinomial-4}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r multinomial-4-test}
#| echo: true
#| results: asis
mod_multinomial_coefs <- tidy(mod_multinomial) |> filter(y.level == "3rd-male")
cat(build_equation(mod_multinomial_coefs, "log-odds(3rd-male vs. 1st-female)"))
```

###

This is just one of five equations the model actually fit — `1st-male`, `2nd-female`, `2nd-male`, and `3rd-female` each have their own, all against the same `1st-female` reference. `3rd-male` shows the starkest contrast: lower class and being male both pull its log-odds down relative to a 1st class woman.

### Exercise 5

Commit and push your work with a message like "Add multinomial regression model". Copy-paste the output of `git log --oneline -3` into the answer box.

```{r multinomial-5}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```
Add multinomial regression model
Add logistic regression model with caching
Add linear regression model
```

###

Six categories and two predictors is intentionally simple — more predictors without more data would only worsen the overfitting risk.

## Ordinal regression
###

`class_sex` had no natural order. `PClass` does: `"1st"` outranks `"2nd"`, which outranks `"3rd"`. `ordinal_reg()` proposes a structure that respects that order — categories, but not interchangeable ones.

### Exercise 1

Add a new chunk. Count how many passengers are in each `PClass`. Render, then run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r ordinal-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

```{r ordinal-1-test}
#| echo: true
count(titanic, PClass)
```

###

A plain `factor` would treat `"1st"`, `"2nd"`, and `"3rd"` as unordered labels, throwing away the fact that `"2nd"` sits between the other two.

### Exercise 2

In place of the count, specify an ordinal regression with `ordinal_reg()`, set its engine to `"polr"` — passing `Hess = TRUE` as an extra argument to `set_engine()`, which `tidy()` needs in order to work — and `fit()` it to predict `factor(PClass, levels = c("1st", "2nd", "3rd"), ordered = TRUE)` from `Age + Sex + Survived` using `titanic`, storing the result as `mod_ordered`. Show `tidy(mod_ordered)`. Render, then run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r ordinal-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r ordinal-2-test}
#| echo: true
mod_ordered <- ordinal_reg() |>
  set_engine("polr", Hess = TRUE) |>
  fit(factor(PClass, levels = c("1st", "2nd", "3rd"), ordered = TRUE) ~ Age + Sex + Survived, data = titanic)

tidy(mod_ordered)
```

###

`set_engine("polr", Hess = TRUE)` is still choosing a **method** — `Hess = TRUE` just passes an extra argument straight through to `MASS::polr()`. The two `scale` rows are cutpoints between classes, not predictor effects; only `Age`, `Sexmale`, and `Survived` describe what a predictor actually does.

### Exercise 3

Show the equation for the three real coefficients — not the cutpoints. Replace `tidy(mod_ordered)` with these two lines, and add `#| results: asis` to the top of the chunk:

```
mod_ordered_coefs <- tidy(mod_ordered) |> filter(coef.type == "coefficient")
cat(build_equation(mod_ordered_coefs, "Score"))
```

Render, then run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r ordinal-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r ordinal-3-test}
#| echo: true
#| results: asis
mod_ordered_coefs <- tidy(mod_ordered) |> filter(coef.type == "coefficient")
cat(build_equation(mod_ordered_coefs, "Score"))
```

###

`PClass` is ordered, so `ordinal_reg()` is the right structure — but notice we have now proposed four different mechanisms for four different outcomes from the same 756 passengers. This is data science's central difficulty: many possible DGMs exist for the same data, and we can never be certain which is true. We just use judgment — often guided by which coefficients' confidence intervals exclude zero.

### Exercise 4

Commit and push your work with a message like "Add ordinal regression model". Copy-paste the output of `git log --oneline -3` into the answer box.

```{r ordinal-4}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```
Add ordinal regression model
Add multinomial regression model
Add logistic regression model with caching
```

###

You have now used parsnip's specify → `set_engine()` → `fit()` → `tidy()` pattern for four different kinds of outcome — a number, two categories, six categories, and three ordered categories. That same four-step shape works for all of them, which is why tidymodels exists: swapping models never means learning a new interface.

## Summary
###

In this tutorial you used tidymodels to fit a linear, a logistic, a multinomial, and an ordinal regression model to Titanic passenger data — the same specify, set the engine, fit, and tidy pattern, four times over, each one shown with its own fitted equation.

### Exercise 1

In a bash Terminal, run `quarto render analysis.qmd` one more time to confirm everything still renders end to end. In the R Terminal, run:

```
show_file("analysis.qmd")
```

CP/CR.

```{r summary-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 20)
```

###

Your `analysis.qmd` should look something like this:

<pre><code>---
title: "Tidymodels"
author: "Your Name"
format: html
execute:
  echo: false
---

&#96;&#96;&#96;{r}
#| message: false
library(tidymodels)
library(tidyverse)
library(broom)
library(ordered)

titanic <- read_csv("data/titanic.csv", show_col_types = FALSE) |>
  mutate(class_sex = paste(PClass, Sex, sep = "-"))
&#96;&#96;&#96;

&#96;&#96;&#96;{r}
#| results: asis
build_equation <- function(coefs, outcome) {
  has_intercept <- "(Intercept)" %in% coefs$term
  other <- coefs[coefs$term != "(Intercept)", ]
  parts <- purrr::map2_chr(other$term, round(other$estimate, 2), function(term, est) {
    sign <- if (est >= 0) "+" else "-"
    paste0(sign, " ", abs(est), " \\times \\text{", term, "}")
  })
  lead <- if (has_intercept) as.character(round(coefs$estimate[coefs$term == "(Intercept)"], 2)) else NULL
  rhs <- paste(c(lead, parts), collapse = " ")
  paste0("$$\\text{", outcome, "} = ", rhs, "$$")
}

mod_linear <- linear_reg() |>
  set_engine("lm") |>
  fit(Age ~ PClass + Sex, data = titanic)

mod_linear_glm <- linear_reg() |>
  set_engine("glm") |>
  fit(Age ~ PClass + Sex, data = titanic)

cat(build_equation(tidy(mod_linear), "Age"))
cat(build_equation(tidy(mod_linear_glm), "Age (glm)"))
&#96;&#96;&#96;

&#96;&#96;&#96;{r}
#| cache: true
mod_logistic <- logistic_reg() |>
  set_engine("glm") |>
  fit(factor(Survived, levels = c(0, 1)) ~ Age + PClass + Sex, data = titanic)
&#96;&#96;&#96;

&#96;&#96;&#96;{r}
#| results: asis
mod_multinomial <- multinom_reg() |>
  set_engine("nnet") |>
  fit(factor(class_sex) ~ Age + Survived, data = titanic)

mod_multinomial_coefs <- tidy(mod_multinomial) |> filter(y.level == "3rd-male")
cat(build_equation(mod_multinomial_coefs, "log-odds(3rd-male vs. 1st-female)"))
&#96;&#96;&#96;

&#96;&#96;&#96;{r}
#| results: asis
mod_ordered <- ordinal_reg() |>
  set_engine("polr", Hess = TRUE) |>
  fit(factor(PClass, levels = c("1st", "2nd", "3rd"), ordered = TRUE) ~ Age + Sex + Survived, data = titanic)

mod_ordered_coefs <- tidy(mod_ordered) |> filter(coef.type == "coefficient")
cat(build_equation(mod_ordered_coefs, "Score"))
&#96;&#96;&#96;
</code></pre>

###

Five chunks — one for setup, one per model — is the entire analysis. Every model chunk follows the same specify → `set_engine()` → `fit()` shape, whether the outcome was a number, two categories, six categories, or three ordered categories, and every one ends by showing its own fitted equation. That brevity is the payoff of learning parsnip's pattern once instead of learning four separate model-fitting interfaces.

### Exercise 2

Commit and push any remaining changes. Copy-paste the URL to your GitHub repo.

```{r summary-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

Your repo holds the exact code that reproduced every model in this tutorial — anyone who clones it and runs `quarto render analysis.qmd` gets the same four fitted models back. That reproducibility, not any single coefficient, is what tidymodels' consistent interface is really buying you.

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
