---
title: Third Tutorial for R4DS
author: Luke Li, David Kane, and Sruthi Gandhi
tutorial:
  id: r4ds-3
output:
  learnr::tutorial:
    df_print: default
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: 'This tutorial for "R for Data Science" covers Chapter 12: Logical vectors, Chapter 13: Numbers, Chapter 14: Strings, Chapter 15: Regular expressions, and Chapter 21: Databases. Exercises practice stringr operations, parse_number, and querying data with DBI and dbplyr.'
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)

library(tidyverse)
library(babynames)
library(nycflights13)
library(DBI)
library(duckdb)
library(dbplyr)

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(out.width = '90%')
options(tutorial.exercise.timelimit = 600,
        tutorial.storage = "local")

con_names <- dbConnect(duckdb::duckdb(),
                       dbdir = "data/nameby_year.duckdb")
con_air <- dbConnect(duckdb::duckdb(),
                     dbdir = "data/nycflights13.duckdb")

dbplyr_names <- tbl(con_names, "babynames")
dbplyr_fly <- tbl(con_air, "flights")

x <- dbplyr_names |>
  filter(str_starts(name, "A"), year > 2000, n >= 10) |>
  mutate(length_class = if_else(str_length(name) <= 5, "Short", "Long")) |>
  summarize(total = sum(n), .by = c(year, length_class)) |>
  arrange(year) |>
  collect()

departure_data <- flights |>
  select(dep_time, sched_dep_time, dep_delay, tailnum) |>
  mutate(tail_number = parse_number(tailnum)) |>
  filter(!is.na(dep_time)) |>
  mutate(
    dep_hour = dep_time %/% 100,
    dep_minute = dep_time %% 100,
    sched_hour = sched_dep_time %/% 100,
    sched_minute = sched_dep_time %% 100,
    true_dep_delay = pmax(dep_delay, 0)
  ) |>
  count(sched_hour, dep_hour, wt = true_dep_delay, name = "total_delay")
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

## Introduction
###

This section covers key concepts from [Chapter 13: Numbers](https://r4ds.hadley.nz/numbers.html), [Chapter 14: Strings](https://r4ds.hadley.nz/strings.html), [Chapter 15: Regular Expressions](https://r4ds.hadley.nz/regexps.html), [Chapter 12: Logical Vectors](https://r4ds.hadley.nz/logicals.html), and [Chapter 21: Databases](https://r4ds.hadley.nz/databases.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund.

You will learn about working with numeric data and text patterns using core *Tidyverse* packages like **readr**, **dplyr**, and **stringr**, as well as database interaction using **dbplyr**. Key concepts include extracting numbers from strings, detecting text patterns, computing parallel maxima across vectors, and connecting to databases.

We recommend using an agentic coding tool such as [Gemini CLI](https://github.com/google-gemini/gemini-cli) or [Claude Code](https://claude.ai/code). Our instructions are written with these tools in mind. You may also use a chat-based AI, but you will need to copy/paste code and data context manually.

### Exercise 1

You should be connected to a repo named `r4ds-3`. If you are not, create one and connect to it.

Create a new file, `analysis.qmd`, with the title `"Analyzing Baby Names and Flights"` and your name as the author. In a bash Terminal, render it:

```
quarto render analysis.qmd
```

Open `analysis.html` with Live Server (right-click it in the Explorer → **Open with Live Server**) and keep the tab open. It refreshes on every render.

Create a `.gitignore` with `analysis_files` followed by a blank line. Commit and push.

In the R Terminal, run:

```
show_file(".gitignore")
```

If that fails, it is probably because you have not yet loaded `library(tutorial.helpers)` in the R Terminal.

CP/CR.

```{r introduction-1}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

###

```
analysis_files
```

###

We will be working with databases in this tutorial. Two key differences between data frames and database tables: database tables are stored on disk and can be arbitrarily large, while data frames are stored in memory; database tables almost always have indexes for fast row lookups, while data frames do not.

### Exercise 2

In your QMD, put `library(tidyverse)`, `library(DBI)`, `library(duckdb)`, and `library(dbplyr)` in a new code chunk. In a bash Terminal, run `quarto render analysis.qmd`. Check your Live Server tab — it should refresh automatically.

Notice that the file does not look good because the code is visible and there are annoying messages. To take care of this, add `#| message: false` to remove all the messages in this code chunk. Also add the following to the YAML header to remove all code echoes from the HTML:

```
execute:
  echo: false
```

In a bash Terminal, run `quarto render analysis.qmd`. Check your Live Server tab. In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r introduction-2}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 6)
```

###

<pre><code>```{r}
#| message: false
library(tidyverse)
library(DBI)
library(duckdb)
library(dbplyr)
```</code></pre>

###

Databases are run by database management systems (DBMS's). DuckDB is an in-process DBMS — it runs entirely within your R session, requires no separate server, and is ideal for analyzing large local datasets.

### Exercise 3

From the R Terminal, run these three commands:

```
getwd()
dir.create("data")
list.files()
```

CP/CR.

```{r introduction-3}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 5)
```

###

```
[1] "analysis.html"  "analysis.qmd"   "analysis_files" "data"
```

###

At the simplest level, a database is a collection of data frames — called tables. Like a data frame, a table is a set of named columns where every value in a column has the same type.


## Baby names
###

In this section we use a DuckDB database called `nameby_year.duckdb`. The data comes from the **babynames** package, which contains names used for American babies from 1880 to 2017.

Key concepts you'll use: opening a **connection** to the database, creating a **lazy table reference** (a query that hasn't run yet), and **collecting** the results — the moment the query actually executes and the data comes back into R.

```{r babynames-data}
dbplyr_names
```

### Exercise 1

Download the DuckDB database `nameby_year.duckdb` from this URL and save it in your `data/` directory:

```
https://github.com/PPBDS/misc.tutorials/raw/refs/heads/main/inst/tutorials/r4ds-3/data/nameby_year.duckdb
```

In a bash Terminal, run:

```
ls data
```

CP/CR.

```{r baby-names-1}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 5)
```

###

```
nameby_year.duckdb
```

###

Downloading a file means copying it from a URL to a local path in your project. A `.duckdb` database is a *binary* file, so it must be downloaded in binary mode — otherwise text-style line-ending conversion can corrupt it. The thing to carry away is the binary-versus-text distinction, not the mechanics.

### Exercise 2

Connect to the database file `data/nameby_year.duckdb`, then create a lazy reference to its `babynames` table. Assign the connection to `con_names` and the table reference to `dbplyr_names`. In a bash Terminal, run `quarto render analysis.qmd`. Check your Live Server tab.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r baby-names-2}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 5)
```

###

```{r baby-names-2-test}
dbplyr_names
```

###

Opening a **connection** is the first step in working with any database: the connection names the database system — here DuckDB — and where to find the file. **[DBI](https://dbi.r-dbi.org/)** handles the low-level SQL execution, while **[dbplyr](https://dbplyr.tidyverse.org/)** translates your **[dplyr](https://dplyr.tidyverse.org/)** code into the SQL that DBI runs.

### Exercise 3

Summarize the data: for each year since 2000, find the total number of babies given names that start with "A", split into "Short" names (5 letters or fewer) and "Long" names, and pull the result into R as `x`. In a bash Terminal, run `quarto render analysis.qmd`.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r baby-names-3}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 6)
```

###

```{r baby-names-3-test}
#| echo: true
dbplyr_names |>
  filter(str_starts(name, "A"), year > 2000, n >= 10) |>
  mutate(length_class = if_else(str_length(name) <= 5, "Short", "Long")) |>
  summarize(total = sum(n), .by = c(year, length_class)) |>
  arrange(year) |>
  collect()
```

###

If your result differs from ours, replace your code with the code above before moving on, so your `x` matches for the rest of the section.

A **lazy table** is the key idea: operations on it are translated to SQL and executed inside the database, not in R's memory, and the results are only retrieved into R when you explicitly **collect** them. Remarkably, **[dbplyr](https://dbplyr.tidyverse.org/)** even translates **[stringr](https://stringr.tidyverse.org/)** string operations into their SQL equivalents, so the same tidyverse-style code runs against the database.

### Exercise 4

Within the latest code chunk, add the option: `#| cache: true`. Assign the result of the pipe to `x`. If you have not done so already, add `analysis_cache` to your `.gitignore`.

In a bash Terminal, run `quarto render analysis.qmd`. This first render will take some time — by including `#| cache: true`, Quarto saves the result to disk so the next render loads from cache instead of re-running. Check your Live Server tab.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r baby-names-4}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 8)
```

###

Everything before the **collect** step in a **dbplyr** pipeline runs as SQL in the database; collecting transfers the results into R as a tibble. Grouping a summary on the fly — totals per year and length class — gets you the aggregated table without leaving a persistent grouping behind on the data.

### Exercise 5

Make a line graph from `x` showing the proportion of "Short" vs. "Long" names starting with "A" over the years, with `length_class` shown by color. Give it a proper title, subtitle, and caption. Don't assign the plot. In a bash Terminal, run `quarto render analysis.qmd`. Check your Live Server tab.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r baby-names-5}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 12)
```

###

```{r baby-names-5-test}
#| echo: true
ggplot(x, aes(x = year, y = total, color = length_class)) +
  geom_line(linewidth = 1.2) +
  labs(
    title = "Trends in Short vs Long A-Names Since 2000",
    subtitle = "Short baby names starting with A are becoming more popular",
    x = "Year",
    y = "Total Babies",
    color = "Name Length",
    caption = "Source: babynames"
  ) +
  scale_y_continuous(labels = scales::label_comma()) +
  theme_minimal()
```

###

Creating categorical variables like `length_class` transforms continuous data (name length in characters) into discrete groups that are easier to interpret and visualize, revealing broader patterns that individual values might obscure.

## Flights
###

This section covers key concepts from [Chapter 13: Numbers](https://r4ds.hadley.nz/numbers.html) and [Chapter 15: Regular Expressions](https://r4ds.hadley.nz/regexps.html). The `nycflights13` dataset contains information about all flights departing from New York City airports in 2013.

Key ideas you'll use: pulling a number out of a messy string, splitting four-digit clock times into hours and minutes with integer division and remainder, and flooring negative values at zero so early departures don't cancel out late ones.

```{r flights-data}
dbplyr_fly
```

### Exercise 1

Download the DuckDB database `nycflights13.duckdb` from this URL and save it in your `data/` directory:

```
https://github.com/PPBDS/misc.tutorials/raw/refs/heads/main/inst/tutorials/r4ds-3/data/nycflights13.duckdb
```

In a bash Terminal, run:

```
ls data
```

CP/CR.

```{r flights-1}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 5)
```

###

```
nameby_year.duckdb  nycflights13.duckdb
```

###

The `nycflights13.duckdb` database contains multiple tables: `airlines`, `airports`, `flights`, `planes`, and `weather`. A database connection can list the tables it holds, which is how you discover what's available after connecting.

### Exercise 2

Connect to the database file `data/nycflights13.duckdb`, then create a lazy reference to its `flights` table. Assign the connection to `con_air` and the table reference to `dbplyr_fly`. In a bash Terminal, run `quarto render analysis.qmd`. Check your Live Server tab.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r flights-2}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 5)
```

###

```{r flights-2-test}
dbplyr_fly
```

###

Lazy tables let you explore database structure without loading data into memory. When you print a lazy table, **dbplyr** shows column names, types, and a preview — all from a lightweight SQL query. This is crucial for understanding unfamiliar databases before writing complex queries.

### Exercise 3

Summarize departure delay patterns: for each combination of scheduled departure hour and actual departure hour, find the total number of delay minutes, treating early departures as zero delay. Pull the result into R as `departure_data`. In a bash Terminal, run `quarto render analysis.qmd`.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r flights-3}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 6)
```

###

```{r flights-3-test}
#| echo: true
dbplyr_fly |>
  select(dep_time, sched_dep_time, dep_delay, tailnum) |>
  filter(!is.na(dep_time)) |>
  collect() |>
  mutate(tail_number = parse_number(tailnum)) |>
  mutate(
    dep_hour = dep_time %/% 100,
    dep_minute = dep_time %% 100,
    sched_hour = sched_dep_time %/% 100,
    sched_minute = sched_dep_time %% 100,
    true_dep_delay = pmax(dep_delay, 0)
  ) |>
  count(sched_hour, dep_hour, wt = true_dep_delay, name = "total_delay")
```

###

If your result differs from ours, replace your code with the code above before moving on, so your `departure_data` matches for the rest of the section.

Four-digit military times like `517` pack the hour and minute together. Integer division by 100 gives the hour (5) and the remainder gives the minute (17) — a standard trick for unpacking a time stored as a number. Flooring each delay at zero (treating early departures as 0) keeps early flights from cancelling out late ones when the delays are summed.

### Exercise 4

Within the latest code chunk, add the option: `#| cache: true`. Assign the result of the pipe to `departure_data`. If you have not done so already, add `analysis_cache` to your `.gitignore`.

In a bash Terminal, run `quarto render analysis.qmd`. This first render will take some time. Check your Live Server tab.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r flights-4}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 8)
```

###

A **weighted count** turns a frequency table into a sum table: instead of counting how many flights fall into each scheduled/actual hour combination, we sum the total delay minutes — revealing which time combinations contribute most to overall delays.

### Exercise 5

Make a heatmap from `departure_data` with `sched_hour` on the x-axis, `dep_hour` on the y-axis, and `total_delay` as the fill color. Give it a proper title, subtitle, and caption. Don't assign the plot. In a bash Terminal, run `quarto render analysis.qmd`. Check your Live Server tab.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r flights-5}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 12)
```

###

```{r flights-5-test}
#| echo: true
ggplot(departure_data, aes(x = sched_hour, y = dep_hour, fill = total_delay)) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(option = "plasma", name = "Total Delay\n(minutes)") +
  labs(
    title = "Flight Delay Patterns: Scheduled vs Actual Departure Hours",
    subtitle = "Darker regions show higher total delay concentrations",
    x = "Scheduled Departure Hour",
    y = "Actual Departure Hour",
    caption = "Source: nycflights13"
  ) +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    plot.title = element_text(face = "bold", size = 16),
    plot.subtitle = element_text(size = 12)
  )
```

###

Once you map a variable to an aesthetic, **ggplot2** takes care of the rest, choosing a reasonable scale and building a legend that explains the mapping. For heatmaps, a **viridis** fill scale gives perceptually uniform colors that work well for continuous data.

## Summary

###

This tutorial covered [Chapter 12: Logical vectors](https://r4ds.hadley.nz/logicals.html), [Chapter 13: Numbers](https://r4ds.hadley.nz/numbers.html), [Chapter 14: Strings](https://r4ds.hadley.nz/strings.html), [Chapter 15: Regular expressions](https://r4ds.hadley.nz/regexps.html), and [Chapter 21: Databases](https://r4ds.hadley.nz/databases.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund.

You learned how to work with numbers, strings, and logical vectors using **readr**, **stringr**, and **dplyr**, and how to query databases using **DBI** and **dbplyr**.

### Exercise 1

In a bash Terminal, run `quarto render analysis.qmd` to ensure that everything works. Check your Live Server tab — the resulting HTML page should be attractive, showing clean versions of your plots.

In the R Terminal, run:

```
show_file("analysis.qmd")
```

CP/CR.

```{r summary-1}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 30)
```

###

Database operations with **[dbplyr](https://dbplyr.tidyverse.org/)** are translated to SQL and executed in the database, with the results only pulled into R when you collect them.

### Exercise 2

Publish your rendered QMD to GitHub Pages. In a bash Terminal, run:

````
quarto publish gh-pages analysis.qmd
````

If the bash Terminal is still running from rendering, stop it first using `Ctrl/Cmd + C`.

Copy/paste the resulting URL below.

```{r summary-2}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 1)
```

###

Publishing the QMD creates a public URL anyone can use to view the rendered project.

### Exercise 3

Commit and push all your files. Copy/paste the URL to your GitHub repo.

```{r summary-3}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

###

GitHub repositories are where people can view the code you used to create your project.

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
