---
title: Ducks
author: Var Kurapati
tutorial:
  id: ducks
output:
  learnr::tutorial:
    df_print: default
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: 'Students use GPS tracking data from 52 Mallard ducks fitted with GPS
  collars at Lake Constance, Germany to explore movement patterns, seasonal activity,
  and individual variation. Final artifact is a multi-page Quarto website with a
  map, seasonal analysis, and written interpretation.'
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(knitr)
library(tidyverse)
library(sf)

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(out.width = '90%')
options(tutorial.exercise.timelimit = 600,
        tutorial.storage = "local")

mallards <- read_csv("../../extdata/ducks/mallards.csv",
                     show_col_types = FALSE) |>
  mutate(
    timestamp = ymd_hms(timestamp),
    month     = month(timestamp, label = TRUE, abbr = TRUE),
    month_num = month(timestamp),
    year      = year(timestamp),
    season    = case_when(
      month_num %in% c(3, 4, 5)   ~ "Spring",
      month_num %in% c(6, 7, 8)   ~ "Summer",
      month_num %in% c(9, 10, 11) ~ "Autumn",
      TRUE                         ~ "Winter"
    ),
    season = factor(season, levels = c("Spring", "Summer", "Autumn", "Winter"))
  )

world <- map_data("world")
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

## Introduction
###

Mallards (*Anas platyrhynchos*) are the most familiar duck in the world — the green-headed bird you've seen in every park pond. But most people don't know how far they actually travel, or how differently individual ducks behave. This tutorial uses GPS tracking data from 52 Mallards fitted with solar-powered collars at Lake Constance, on the border of Germany and Switzerland, to explore where they go, when they move most, and how much individuals differ.

The data comes from a published study by Korner, Sauter, Fiedler, and Jenni (2016) — ["Variable allocation of activity to daylight and night in the mallard"](https://doi.org/10.1016/j.anbehav.2016.02.026) — and is freely available through [Movebank](https://movebank.org/), the world's largest archive of animal tracking data. The dataset is licensed under Creative Commons Attribution (CC BY).

We recommend using an agentic coding tool such as [Gemini CLI](https://github.com/google-gemini/gemini-cli) or [Claude Code](https://claude.ai/code), which can read and edit your files directly. Our instructions are written with these tools in mind. You may also use a chat-based AI, but you will need to copy/paste code and data context manually.

### Exercise 1

You should be connected to a repo named `ducks`. If you are not, create one and connect to it.

Create a new file, `index.qmd`, with the title `"Mallard Duck Movement"` and your name as the author. In a bash Terminal, render it:

```
quarto render index.qmd
```

Open `index.html` with Live Server (right-click it in the Explorer → **Open with Live Server**) and keep the tab open. It refreshes on every render. Going forward, we will just tell you to "Render" when we want you to take these steps.

Create a `.gitignore` with `*_files` and `*_cache` each on their own line, followed by a blank line. Commit and push.

In the R Terminal, run:

```
show_file(".gitignore")
```

If that fails, it is probably because you have not yet loaded `library(tutorial.helpers)` in the R Terminal.

CP/CR.

```{r introduction-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

<pre><code>*_files
*_cache
</code></pre>

###

This tutorial builds a website rather than a single page. The final product will have three pages: an introduction page (`index.qmd`), a map page (`map.qmd`), and a seasonal analysis page (`seasons.qmd`). A `_quarto.yml` file will tie them together into a navigable site. You will create these files one at a time as the tutorial progresses.

### Exercise 2


```
execute:
  echo: false
```

Render. In the R Terminal, run:

```
show_file("index.qmd", chunk = "Last")
```

CP/CR.

```{r introduction-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

<pre><code>#| message: false
library(tidyverse)
library(sf)
</code></pre>

###

**[lubridate](https://lubridate.tidyverse.org/)** handles dates and times — essential for GPS tracking data where every fix has a timestamp. **[maps](https://cran.r-project.org/package=maps)** provides world map outlines for geographic plots. Both are standard tools for any analysis that involves location and time.

### Exercise 3

Create a `data` directory at the top level of the `ducks` repo. In the bash Terminal, run `ls`.

CP/CR.

```{r introduction-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 4)
```

###

```
data  index.html  index.qmd  *_files
```

## Where do they go?
###

Lake Constance sits at the border of Germany, Switzerland, and Austria at roughly 47.5°N, 9.5°E. It is one of the largest freshwater lakes in Central Europe and an important staging area for waterfowl. But Mallards are not purely resident birds — some individuals make substantial seasonal movements while others barely leave the lake. The GPS data reveals exactly who goes where.

### Exercise 1

Download the mallard tracking data from this URL and save it in your `data/` directory:

```
https://github.com/PPBDS/misc.tutorials/raw/refs/heads/main/inst/extdata/ducks/mallards.csv
```

In the bash Terminal, run:

```
ls data
```

CP/CR.

```{r where-do-they-go-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 4)
```

###

```
mallards.csv
```

###

The original Movebank download contained 748,082 rows mixing GPS fixes and accelerometer records, plus dozens of technical columns. This teaching version keeps only the 83,279 GPS rows and the columns needed for movement analysis: timestamp, latitude, longitude, speed, heading, and duck identifier.

### Exercise 2

Create a new code chunk in `index.qmd` that reads `data/mallards.csv` and prints the first 10 rows. Render.

In the R Terminal, run `show_file("index.qmd", chunk = "Last")`. CP/CR.

```{r where-do-they-go-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```{r where-do-they-go-2-test}
#| echo: true
read_csv("../../extdata/ducks/mallards.csv", show_col_types = FALSE) |>
  head(10)
```

###

The dataset has 83,279 rows and 7 columns. Each row is one GPS fix — a single location recorded by one duck at one moment. The `duck_id` column combines a numeric tag ID with a location and serial number. Notice that GPS fixes are not evenly distributed across ducks: some individuals have thousands of records, others only dozens.

### Exercise 3

Count GPS fixes per duck, sorted from most to least. Render.

In the R Terminal, run `show_file("index.qmd", chunk = "Last")`. CP/CR.

```{r where-do-they-go-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```{r where-do-they-go-3-test}
#| echo: true
read_csv("../../extdata/ducks/mallards.csv", show_col_types = FALSE) |>
  count(duck_id, sort = TRUE)
```

###

Duck `2351` has 20,169 fixes — by far the most of any individual. Several ducks have fewer than 100 records, suggesting their collars failed early or the birds died. This kind of uneven data collection is typical of wildlife tracking studies. For any analysis that compares individuals, you should keep in mind that ducks with more fixes will have more complete pictures of their movement than ducks with few fixes.

### Exercise 4

Summarize the geographic range of the data: minimum and maximum latitude and longitude. Render.

In the R Terminal, run `show_file("index.qmd", chunk = "Last")`. CP/CR.

```{r where-do-they-go-4}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```{r where-do-they-go-4-test}
#| echo: true
read_csv("../../extdata/ducks/mallards.csv", show_col_types = FALSE) |>
  summarise(
    lat_min  = min(lat, na.rm = TRUE),
    lat_max  = max(lat, na.rm = TRUE),
    long_min = min(long, na.rm = TRUE),
    long_max = max(long, na.rm = TRUE)
  )
```

###

Latitudes range from 47.3°N to 53.9°N — roughly from Lake Constance in southern Germany up to the Netherlands or northern Germany. Longitudes span from 7.8°E to 23.6°E, reaching as far east as Poland or Hungary. These are large movements for a duck most people associate with park ponds. Lake Constance sits at the southwestern corner of this range, suggesting that at least some ducks leave the lake and travel hundreds of kilometers northeast.

### Exercise 5

Read `data/mallards.csv`, parse the timestamp with `ymd_hms()`, add `month`, `year`, and `season` columns, and assign the result to `mallards`. Print `mallards`. Render.

In the R Terminal, run `show_file("index.qmd", chunk = "Last")`. CP/CR.

```{r where-do-they-go-5}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 10)
```

###

```{r where-do-they-go-5-test}
#| echo: true
mallards <- read_csv("../../extdata/ducks/mallards.csv",
                     show_col_types = FALSE) |>
  mutate(
    timestamp = ymd_hms(timestamp),
    month     = month(timestamp, label = TRUE, abbr = TRUE),
    month_num = month(timestamp),
    year      = year(timestamp),
    season    = case_when(
      month_num %in% c(3, 4, 5)   ~ "Spring",
      month_num %in% c(6, 7, 8)   ~ "Summer",
      month_num %in% c(9, 10, 11) ~ "Autumn",
      TRUE                         ~ "Winter"
    ),
    season = factor(season, levels = c("Spring", "Summer", "Autumn", "Winter"))
  )
mallards
```

###

The `season` column is an ordered factor so that plots display seasons in natural calendar order rather than alphabetically. Without the explicit `levels` argument, R would sort them as Autumn, Spring, Summer, Winter — meaningless for a seasonal analysis.

### Exercise 6

Cache the `mallards` chunk. Add `#| cache: true` as the first line inside the chunk. Render. In the bash Terminal, run `ls`.

CP/CR.

```{r where-do-they-go-6}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 4)
```

###

```
data  index.html  index.qmd  index_cache  *_files
```

###

`index_cache/` appears after the first cached render. The cache stores the parsed `mallards` tibble so subsequent renders skip re-reading and processing the CSV. Each `.qmd` page gets its own cache directory — `map_cache` and `seasons_cache` will appear when you render those pages.

### Exercise 7

Add `*_cache` to your `.gitignore` on its own line. In the R Terminal, run `show_file(".gitignore")`.

CP/CR.

```{r where-do-they-go-7}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 4)
```

###

<pre><code>*_files
*_cache
</code></pre>

###

Cache files are machine-specific and regenerated on every render — they should never go to GitHub. Using `*_cache` covers all three pages at once — `index_cache`, `map_cache`, and `seasons_cache` — so you never need to add each one separately.

### Exercise 8

Create a new file `map.qmd` in your repo. Add the same YAML header and library chunk as `index.qmd`. Also add the same `mallards` data chunk with `#| cache: true`. Render `map.qmd`. In the R Terminal, run `show_file("map.qmd", chunk = "Last")`. CP/CR.

```{r where-do-they-go-8}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

<pre><code>#| cache: true
mallards <- read_csv("data/mallards.csv", show_col_types = FALSE) |>
  mutate(
    timestamp = ymd_hms(timestamp),
    month     = month(timestamp, label = TRUE, abbr = TRUE),
    month_num = month(timestamp),
    year      = year(timestamp),
    season    = case_when(
      month_num %in% c(3, 4, 5)   ~ "Spring",
      month_num %in% c(6, 7, 8)   ~ "Summer",
      month_num %in% c(9, 10, 11) ~ "Autumn",
      TRUE                         ~ "Winter"
    ),
    season = factor(season, levels = c("Spring", "Summer", "Autumn", "Winter"))
  )
</code></pre>

###

Each page of a Quarto website is an independent document with its own R session. That means `mallards` defined in `index.qmd` is not available in `map.qmd` — every page that needs the data must load it itself.

### Exercise 9

In `map.qmd`, add a new working chunk that makes a map of all GPS fixes, colored by individual duck, filtered to the region of interest (latitudes 46–55°N, longitudes 7–25°E). Render `map.qmd`.

In the R Terminal, run `show_file("map.qmd", chunk = "Last")`. CP/CR.

```{r where-do-they-go-9}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 14)
```

###

```{r where-do-they-go-9-test}
#| echo: true
world <- map_data("world")

mallards |>
  filter(lat >= 46, lat <= 55, long >= 7, long <= 25) |>
  ggplot() +
  geom_map(
    data = world, map = world,
    aes(map_id = region),
    fill = "gray90", color = "gray70", linewidth = 0.2
  ) +
  geom_point(
    aes(x = long, y = lat, color = duck_id),
    alpha = 0.2, size = 0.5
  ) +
  expand_limits(x = c(7, 25), y = c(46, 55)) +
  labs(
    title    = "Mallard duck GPS fixes, 2007–2017",
    subtitle = "Each color is one individual duck",
    x        = NULL,
    y        = NULL,
    caption  = "Source: Korner, Sauter, Fiedler & Jenni (2016) via Movebank (CC BY)"
  ) +
  theme_minimal() +
  theme(
    axis.text       = element_blank(),
    axis.ticks      = element_blank(),
    legend.position = "none"
  )
```

###

Most GPS fixes cluster tightly around Lake Constance in the southwest corner of the map — the ducks' home base. But a scatter of points extends northeast toward Poland and Hungary, showing that some individuals made substantial long-distance movements. The dense cluster and the scattered outliers tell two different duck stories: residents that stay near the lake and dispersers that travel hundreds of kilometers.

### Exercise 10

Write 1–2 sentences interpreting the map. Where do most ducks stay, and what do the outlying points tell you? Add your interpretation as prose in `map.qmd` below the map chunk. Render `map.qmd`.

Paste your interpretation below.

```{r where-do-they-go-10}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 4)
```

###

Most ducks stayed close to Lake Constance throughout the study, but a handful of individuals made long-distance movements of several hundred kilometers to the northeast. This individual variation is striking — the same species, tagged at the same location, shows a wide range of movement strategies from resident to long-distance disperser.

###

Mallards are considered a partially migratory species: some populations are fully resident year-round, others migrate long distances, and many individuals fall somewhere in between. Lake Constance is a mild wintering site that attracts ducks from colder regions further north, so some of the northeastern points may represent birds returning to their breeding grounds in spring rather than local residents exploring.

### Exercise 11

Commit `index.qmd` and `map.qmd` with the message `"Add duck map"` and push to GitHub.

CP/CR.

```{r where-do-they-go-11}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

The map shows where ducks go. The next section asks when — how does duck movement vary across seasons and months?

## When do they move?
###

GPS speed tells us not just where a duck was, but how actively it was moving at each fix. A speed near zero means the duck was resting or feeding; higher speeds mean active flight. Combined with timestamps, speed reveals the seasonal rhythm of duck activity across the 10-year study.

### Exercise 1

Create a new file `seasons.qmd` in your repo with the same YAML, libraries, and cached `mallards` chunk as the other pages. Then add a working chunk that counts GPS fixes per month. Render `seasons.qmd`.

In the R Terminal, run `show_file("seasons.qmd", chunk = "Last")`. CP/CR.

```{r when-do-they-move-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```{r when-do-they-move-1-test}
#| echo: true
mallards |>
  count(month)
```

###

September dominates with 20,977 fixes — more than double any other month. This is not because ducks move more in September; it reflects when the collars were most active. Most collars were deployed in August or September, so the early weeks of tracking accumulate many fixes before some collars fail or ducks leave the study area. Always check the raw record count before interpreting patterns as biological — an uneven sampling distribution can look like a behavioral signal.

### Exercise 2

Get summary statistics for `speed`. Render `seasons.qmd`.

In the R Terminal, run `show_file("seasons.qmd", chunk = "Last")`. CP/CR.

```{r when-do-they-move-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```{r when-do-they-move-2-test}
#| echo: true
mallards |>
  summarise(
    min_speed    = min(speed, na.rm = TRUE),
    median_speed = median(speed, na.rm = TRUE),
    mean_speed   = mean(speed, na.rm = TRUE),
    max_speed    = max(speed, na.rm = TRUE)
  )
```

###

The median speed is 0.22 m/s — most fixes catch a duck that is barely moving. The maximum of 127 m/s is physically impossible for a Mallard (real flight speeds top out around 25–30 m/s) and is almost certainly a GPS error. A handful of extreme outliers like this are common in collar data; they should be removed before any speed analysis to avoid distorting the results.

### Exercise 3

Filter `mallards` to `speed <= 30` to remove GPS errors, then compute mean speed by season and month. Render `seasons.qmd`.

In the R Terminal, run `show_file("seasons.qmd", chunk = "Last")`. CP/CR.

```{r when-do-they-move-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

```{r when-do-they-move-3-test}
#| echo: true
mallards |>
  filter(speed <= 30) |>
  group_by(season, month, month_num) |>
  summarise(mean_speed = mean(speed, na.rm = TRUE), .groups = "drop") |>
  arrange(month_num)
```

###

Mean speed is highest in August and September, corresponding to the period when ducks are most likely to make longer movements — either post-breeding dispersal or early autumn staging. Winter months (December–February) show the lowest speeds, consistent with ducks resting near the lake during cold weather. The difference is subtle because mean speed is always pulled toward zero by the many stationary fixes; the pattern is real but modest.

### Exercise 4

Make a bar chart of mean speed by month, colored by season, with GPS errors removed. Render `seasons.qmd`.

In the R Terminal, run `show_file("seasons.qmd", chunk = "Last")`. CP/CR.

```{r when-do-they-move-4}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 12)
```

###

```{r when-do-they-move-4-test}
#| echo: true
mallards |>
  filter(speed <= 30) |>
  group_by(season, month, month_num) |>
  summarise(mean_speed = mean(speed, na.rm = TRUE), .groups = "drop") |>
  ggplot(aes(x = fct_reorder(month, month_num), y = mean_speed, fill = season)) +
  geom_col() +
  scale_fill_manual(values = c(
    "Spring" = "#74c476",
    "Summer" = "#fd8d3c",
    "Autumn" = "#9ecae1",
    "Winter" = "#6baed6"
  )) +
  labs(
    title    = "Mean Mallard movement speed by month",
    subtitle = "Ducks move most in late summer and autumn; slowest in winter",
    x        = NULL,
    y        = "Mean ground speed (m/s)",
    fill     = "Season",
    caption  = "Source: Korner, Sauter, Fiedler & Jenni (2016) via Movebank (CC BY)"
  ) +
  theme_minimal()
```

###

The late summer and early autumn peak in movement speed aligns with the post-breeding season, when young ducks from this year's broods begin exploring beyond their natal area and adult ducks resume longer-range movements after the summer moult. Mallards are partially nocturnal movers — the published study found that ducks shifted activity between day and night flexibly across the year — but this GPS dataset only shows where they were, not what they were doing.

### Exercise 5

Write 1–2 sentences interpreting the seasonal chart. When do ducks move most and least, and what might explain the pattern? Add your interpretation as prose in `seasons.qmd` below the chart. Render `seasons.qmd`.

Paste your interpretation below.

```{r when-do-they-move-5}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 4)
```

###

Mallards move most actively in late summer and autumn, likely reflecting post-breeding dispersal and early staging movements before winter. Movement is slowest in winter when ducks concentrate near Lake Constance, a mild site that attracts wintering waterfowl from colder regions further north.

###

Mean speed per month understates how dramatic migration events actually are. Most GPS fixes still catch a stationary duck; only a small fraction capture active flight. A cleaner view of migration timing would filter to high-speed fixes only (e.g., `speed > 5 m/s`) and count how many occur each month — that would show the actual distribution of flight events rather than the mean diluted by thousands of resting fixes.

### Exercise 6

Commit `seasons.qmd` with the message `"Add seasonal movement analysis"` and push to GitHub.

CP/CR.

```{r when-do-they-move-6}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

The map and the seasonal chart answer different questions: the map shows the geographic range of duck movements, and the seasonal chart shows when ducks are most and least active. The final section ties these together into a published website.

## Summary
###

You have built three pages of analysis. Now you will wire them together into a navigable Quarto website, add introductory text to the home page, and publish the result.

### Exercise 1

Create a `_quarto.yml` file at the top level of your repo. It should define a website with a title, a navbar, and links to all three pages. Ask AI to create it for you. Render the site with:

```
quarto render
```

In the R Terminal, run `show_file("_quarto.yml")`. CP/CR.

```{r summary-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 15)
```

###

<pre><code>project:
  type: website

website:
  title: "Mallard Duck Movement"
  navbar:
    left:
      - href: index.qmd
        text: Home
      - href: map.qmd
        text: Map
      - href: seasons.qmd
        text: Seasons

format:
  html:
    theme: cosmo
</code></pre>

###

`_quarto.yml` is the configuration file that turns a folder of `.qmd` files into a website. Without it, each file renders independently; with it, all pages share a navbar, theme, and consistent layout. The `type: website` line is what triggers website mode.

### Exercise 2

Add introductory text to `index.qmd`. Ask AI to write 2–3 short paragraphs introducing the study: what Mallard ducks are, where Lake Constance is, and what the GPS tracking data reveals. Render.

In the R Terminal, run `show_file("index.qmd")`. CP/CR.

```{r summary-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 20)
```

###

<pre><code>---
title: "Mallard Duck Movement"
author: "Your Name"
execute:
  echo: false
---

#| message: false
library(tidyverse)
library(sf)

## About this project

Mallards (*Anas platyrhynchos*) are the most familiar duck in the world — the
green-headed bird found on park ponds across Europe and North America. Despite
their reputation as stay-at-home birds, Mallards are partially migratory: some
individuals travel hundreds of kilometers each year, while others barely leave
their home lake.

Lake Constance sits at the border of Germany, Switzerland, and Austria and is
one of the largest freshwater lakes in Central Europe. Between 2007 and 2017,
researchers fitted 52 Mallards with solar-powered GPS collars to track their
movements. Each collar recorded the duck's latitude, longitude, and speed every
few hours, building a detailed picture of where each individual went across
multiple years.

This website explores three questions from that dataset: where do the ducks go,
when do they move most, and how fast do they travel? The data comes from a
published study by Korner, Sauter, Fiedler, and Jenni (2016) and is freely
available through Movebank under a Creative Commons Attribution license.
</code></pre>

###

The home page should give a reader who knows nothing about the study enough context to understand the map and seasonal analysis pages. Good introductory text names the species, explains the study location, describes what GPS tracking data is, and previews the main findings — all in plain language without jargon.

### Exercise 3

Publish the website to GitHub Pages. In the bash Terminal, run:

```
quarto publish gh-pages
```

Copy/paste the resulting URL below.

```{r summary-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

`quarto publish gh-pages` without a filename publishes the whole website, not just a single page. The resulting URL gives access to all three pages via the navbar.

### Exercise 4

Commit and push any remaining changes. Copy/paste the URL to your GitHub repo.

```{r summary-4}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

Movebank hosts thousands of published animal tracking studies under open licenses. The same workflow used here — download a CSV, clean timestamps and coordinates, map locations, analyze movement patterns, build a website — applies to any species in the archive, from monarch butterflies to great white sharks. The `move2` R package connects directly to Movebank's API, so for a more advanced project you could pull any tracking dataset straight into R without a manual download.

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
