---
title: Analyzing US Census Data
author: David Kane, Ansh Pathak, and Paramanyu Gupta
tutorial:
  id: census
output:
  learnr::tutorial:
    df_print: default
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: "Companion to the tidycensus chapters of 'Analyzing US Census Data' by Kyle Walker. Pulls American Community Survey data with tidycensus and builds three published artifacts: choropleth maps of Texas household income and California educational attainment, and a bar chart of California median age."
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(tidyverse)
library(tidycensus)
library(sf)
library(knitr)

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(out.width = '90%')
options(tutorial.exercise.timelimit = 600,
        tutorial.storage = "local")

# The data below was pulled once with tidycensus and saved, so this tutorial
# renders without hitting the Census API. Students make the live calls in their
# own analysis.qmd.

# income_tx <- get_acs(geography = "county", variables = "B19013_001",
#                      state = "TX", year = 2020, geometry = TRUE)
# write_rds(income_tx, "data/income_tx.rds")
income_tx <- read_rds("data/income_tx.rds")

# edu_ca <- get_acs(geography = "county",
#                   variables = c("B15003_001", "B15003_022", "B15003_023", "B15003_024", "B15003_025"),
#                   state = "CA", year = 2020, geometry = TRUE, summary_var = "B15003_001")
# write_rds(edu_ca, "data/edu_ca.rds")
edu_ca <- read_rds("data/edu_ca.rds")

# age_ca <- get_acs(geography = "county",
#                   variables = c(median_age = "B01002_001", population = "B01003_001"),
#                   state = "CA", year = 2020, geometry = FALSE)
# write_rds(age_ca, "data/age_ca.rds")
age_ca <- read_rds("data/age_ca.rds")
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

## Introduction
###

This tutorial follows the **[tidycensus](https://walker-data.com/tidycensus/)** chapters of [Analyzing US Census Data](https://walker-data.com/census-r/) by Kyle Walker — [An introduction to tidycensus](https://walker-data.com/census-r/an-introduction-to-tidycensus.html) and [Mapping Census data with R](https://walker-data.com/census-r/mapping-census-data-with-r.html). You will pull data from the American Community Survey through the Census API and turn it into three published artifacts: choropleth maps of Texas household income and California educational attainment, and a bar chart of California median age.

Use AI however you like throughout. An agentic coding tool (such as Claude Code or Gemini CLI) works best, since it can read and edit your files directly; a chat-based assistant works too, with more copy/paste.

### Exercise 1

You should be in a repo named `census`. If you are not, create one and move to it. Create a new file and save it as `analysis.qmd`. Add a YAML header with a title (`"Analyzing US Census Data"`) and your name as author. In a bash Terminal, run:

```
quarto render analysis.qmd
```

Open `analysis.html` by right-clicking it in the Explorer and selecting **Open with Live Server**. Keep this tab open — it refreshes automatically whenever you re-render.

Create a `.gitignore` file with `analysis_files` on the first line and then a blank line. Commit and push your work. In the R Terminal, run:

```
show_file(".gitignore")
```

CP/CR.

```{r introduction-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

The American Community Survey (ACS) is the Census Bureau's rolling annual survey of about 3.5 million households. Unlike the Decennial Census, which counts everyone every ten years, the ACS estimates characteristics like income, education, and age from a sample — which is why every value arrives with a margin of error.

### Exercise 2

In `analysis.qmd`, add a code chunk that loads the three libraries you will need:

```
library(tidyverse)
library(tidycensus)
library(sf)
```

Rendering shows the code and the package start-up messages, which we don't want in a published document. Add `#| message: false` to the chunk, and add this to the YAML header to hide all code:

```
execute:
  echo: false
```

Render again. In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r introduction-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

**tidycensus** returns ACS data as a tidy data frame, one row per geography (or per geography-variable). If you also ask for the geographic boundaries, it hands you an **[sf](https://r-spatial.github.io/sf/)** object — a table with a spatial column attached — that **[ggplot2](https://ggplot2.tidyverse.org/)** can map directly. **sf** is the standard package for spatial data in R, and the boundaries themselves come from **[tigris](https://github.com/walkerke/tigris)**, which **tidycensus** uses under the hood.

### Exercise 3

In a bash Terminal, run `quarto render analysis.qmd` and confirm your Live Server tab refreshes on its own.

CP/CR.

```{r introduction-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

**tidycensus** works without an API key for light use — keyless users may make up to 500 queries per day, far more than this tutorial needs. For heavier work, request a free key at [the Census signup page](https://api.census.gov/data/key_signup.html) and set it once at the start of your session (your AI agent can do this for you). Most data-provider APIs work this way: anonymous access is rate-limited, and a free key raises the ceiling.

## Texas income
###

Median household income is one of the most-requested ACS variables. Its code is `B19013_001`. In this section you will map it across all 254 Texas counties.

### Exercise 1

Add a chunk to `analysis.qmd` that uses **tidycensus** to get 2020 median household income for every county in Texas, with county geometry attached for mapping, and saves it as `income_tx`. Add `#| message: false` to the chunk to hide the connection messages. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r texas-income-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

Our code:

```
income_tx <- get_acs(
  geography = "county",
  variables = "B19013_001",
  state = "TX",
  year = 2020,
  geometry = TRUE
)
```

###

Every ACS request comes down to three choices: the **geographic level** (county here, but census tract, place, or state are equally available), the **variables** — Census table codes like `B19013_001` for median household income — and the **year**, which for the ACS names a 5-year window (2020 means the 2016–2020 estimates, pooled across five years so even small counties have reliable numbers). The first render makes a real network call to the Census API, so it takes a moment.

### Exercise 2

Add a chunk that prints `income_tx` so you can see its structure. Render, then copy and paste the printed output from the HTML below.

```{r texas-income-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

Our output:

```{r texas-income-2-test}
income_tx
```

###

Every county arrives with a `GEOID` (its FIPS code), a `NAME`, the `variable` code, an `estimate`, and a margin of error (`moe`). The `estimate` is the survey's best guess; the `moe` is the half-width of its 90% confidence interval. Because the ACS is a *sample*, not a full count, every estimate carries a margin — and that margin is smaller in populous counties, where the survey reaches more households. Because you asked for boundaries, there is also a spatial column holding each county's polygon.

### Exercise 3

Change that chunk so that, instead of printing `income_tx`, it draws a choropleth map of median income across the counties using **ggplot2**. Render and check the map in your Live Server tab.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r texas-income-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

Our rough map:

```{r texas-income-3-test}
#| message: false
ggplot(income_tx) +
  geom_sf(aes(fill = estimate))
```

###

This is a **choropleth** — each county drawn from its boundary and shaded by its income. Two things make this first version hard to read: the default blue gradient runs dark-to-light in a direction many readers misjudge, and the gray panel with latitude/longitude gridlines adds clutter a map does not need. Fixing both is the next step.

### Exercise 4

Improve the map: give it a perceptually uniform color scale (the **viridis** family is the standard choice), a clear title and legend label, a caption naming the data source, and a clean map theme. Render and compare to ours.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r texas-income-4}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

Our improved map:

```{r texas-income-4-test}
#| message: false
ggplot(income_tx) +
  geom_sf(aes(fill = estimate), color = "white", size = 0.1) +
  scale_fill_viridis_c(option = "plasma", name = "Median income") +
  labs(
    title = "Median household income by county, Texas (2020)",
    caption = "Source: ACS 5-year estimates via tidycensus"
  ) +
  theme_void()
```

###

The **viridis** color scales are designed to be perceptually uniform and readable by colorblind viewers, which is why they have become a default for choropleths; **ggplot2** ships them, and the **[RColorBrewer](https://colorbrewer2.org/)** palettes are the other common choice. Clearing away the gray panel and the latitude/longitude gridlines a map does not need leaves the counties to carry the message.

### Exercise 5

In one or two sentences, describe what the map shows. Where is median household income highest and lowest across Texas, and what pattern do you notice?

```{r texas-income-5}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

Your answer might note that the highest incomes cluster in suburban counties around the big metros — Austin, Dallas–Fort Worth, and Houston — while many rural counties, especially along the Rio Grande border in the south, have the lowest median incomes.

###

A published map earns its place by making a pattern legible at a glance. The map tells the reader *where*; the caption and a sentence of interpretation tell them *what to take away* and *what the numbers can support*.

### Exercise 6

Commit your work. Stage `analysis.qmd`, commit with the message `"Add Texas income map"`, and push to GitHub. You can do this with your AI agent, the VS Code Source Control panel, or `git` on the command line.

CP/CR, or write a short note describing how you committed and pushed.

```{r texas-income-6}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

A choropleth shades each area by a value, so it is only as honest as the variable behind it. Median household income is a good fit because it is roughly comparable across counties; a raw count like "number of high earners" would mostly track county population and mislead.

## California bachelor's degrees
###

Educational attainment lives in ACS table B15003, which breaks the adult population into 25 schooling levels from "no schooling" up to "doctorate." In this section you will compute the share of adults in each California county with at least a bachelor's degree and map it.

### Exercise 1

Add a chunk that uses **tidycensus** to get 2020 educational attainment for every county in California, with county geometry attached, and save it as `edu_ca`. You need the table total plus the bachelor's-and-higher levels: variables `B15003_001` (total), `B15003_022`, `B15003_023`, `B15003_024`, and `B15003_025`. Pass `summary_var = "B15003_001"` so the total rides along on every row. Add `#| message: false`. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r california-bachelors-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

Our code:

```
edu_ca <- get_acs(
  geography = "county",
  variables = c("B15003_001", "B15003_022", "B15003_023", "B15003_024", "B15003_025"),
  state = "CA",
  year = 2020,
  geometry = TRUE,
  summary_var = "B15003_001"
)
```

###

These five variable codes report *completed* degrees, not enrollment. `B15003_022` through `B15003_025` are bachelor's, master's, professional, and doctorate degrees — four separate, non-overlapping levels — so adding them counts everyone with at least a bachelor's. Asking for the table total (`B15003_001`) to ride along on every row gives you the denominator — the county's whole adult population — that you need to turn those counts into a percentage.

### Exercise 2

Add a chunk that prints `edu_ca`. Render, then copy and paste the printed output from the HTML.

```{r california-bachelors-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

Our output:

```{r california-bachelors-2-test}
edu_ca
```

###

The data is in **long** format: one row per county-variable combination, so each of California's 58 counties appears five times (290 rows in all). The `estimate` column holds the count for that one variable, while `summary_est` repeats the county total on every row. To get a single percentage per county you will collapse those five rows back down.

### Exercise 3

You are done changing the chunk that downloads `edu_ca`, and re-downloading it on every render is slow. Add `#| cache: true` to that chunk. Render — this first render is slower because Quarto runs the code and writes the result to disk; later renders skip the Census download and load `edu_ca` from the cache.

Cached output should never go to GitHub. Add `analysis_cache` to your `.gitignore` on its own line. In the R Terminal, run:

```
show_file(".gitignore")
```

CP/CR.

```{r california-bachelors-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 4)
```

###

Your `.gitignore` should now contain both lines:

```
analysis_files
analysis_cache
```

###

Caching is for finished, expensive chunks — a network download like this one, not code you are still editing. The cache is generated output: useful locally, but it does not belong on GitHub. The same `#| cache: true` is worth adding to your other Census-download chunks (Texas, and the upcoming median-age pull) so every render after the first is fast.

### Exercise 4

Change your working chunk so it computes, for each county, the percentage of adults with at least a bachelor's degree — the summed bachelor's-and-higher counts divided by the county total — and draws the result as a choropleth map with **ggplot2**. Render and check the map.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r california-bachelors-4}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

Our code and map:

```
edu_ca |>
  group_by(GEOID) |>
  summarize(
    percent = 100 * sum(estimate[variable != "B15003_001"]) / unique(summary_est)
  ) |>
  ggplot() +
  geom_sf(aes(fill = percent)) +
  scale_fill_viridis_c(option = "C", name = "% with degree") +
  labs(
    title = "Adults with a bachelor's degree or higher by county, California (2020)",
    caption = "Source: ACS 5-year estimates via tidycensus"
  ) +
  theme_void()
```

```{r california-bachelors-4-test}
#| message: false
edu_ca |>
  group_by(GEOID) |>
  summarize(
    percent = 100 * sum(estimate[variable != "B15003_001"]) / unique(summary_est)
  ) |>
  ggplot() +
  geom_sf(aes(fill = percent)) +
  scale_fill_viridis_c(option = "C", name = "% with degree") +
  labs(
    title = "Adults with a bachelor's degree or higher by county, California (2020)",
    caption = "Source: ACS 5-year estimates via tidycensus"
  ) +
  theme_void()
```

###

The key step is **aggregating** the five rows per county down to one. Summing the four degree counts and dividing by the county total turns raw counts into a comparable percentage. Because the data is a spatial **sf** object, collapsing those rows also dissolves each county's five identical boundaries into a single polygon — so the per-county percentages map cleanly with no extra work.

### Exercise 5

In one or two sentences, describe the pattern. Which parts of California have the highest share of college graduates, and which have the lowest?

```{r california-bachelors-5}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

Your answer might note that the coastal and urban counties — the Bay Area, around Los Angeles, and a few mountain-resort counties — have the highest shares of adults with a bachelor's degree, while many inland agricultural counties in the Central Valley and the far north have the lowest.

###

Educational attainment is one of the strongest county-level correlates of income, so this map and the Texas income map are partly telling the same story through different variables. Noticing that overlap — and resisting the urge to claim one *causes* the other from a map alone — is the kind of judgment a data scientist brings to Census work.

### Exercise 6

Commit your work. Stage `analysis.qmd`, commit with the message `"Add California education map"`, and push to GitHub.

CP/CR, or write a short note describing how you committed and pushed.

```{r california-bachelors-6}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

Percentages make counties comparable regardless of size, which is why this map is fairer than mapping the raw count of degree-holders. A choropleth of counts would just light up Los Angeles and San Diego because that is where the people are.

## California median age
###

Not every Census question is a map. Here you will pull median age and total population for California counties and build a bar chart — and meet the reshaping step that long-format ACS data almost always needs.

### Exercise 1

Add a chunk that uses **tidycensus** to get 2020 median age and total population for every county in California, naming the variables `median_age` (`B01002_001`) and `population` (`B01003_001`). You are building a bar chart, not a map, so set `geometry = FALSE`. Save it as `age_ca` and add `#| message: false`. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r california-age-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

Our code:

```
age_ca <- get_acs(
  geography = "county",
  variables = c(median_age = "B01002_001", population = "B01003_001"),
  state = "CA",
  year = 2020,
  geometry = FALSE
)
```

###

You can give each variable a readable name — `median_age`, `population` — instead of carrying raw codes like `B01002_001` around, which makes everything downstream easier to read. And because this section builds a bar chart rather than a map, skipping the boundary download entirely is the right call — there is no reason to pull polygons you will not draw.

### Exercise 2

Add a chunk that prints `age_ca`. Render, then copy and paste the printed output from the HTML.

```{r california-age-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

Our output:

```{r california-age-2-test}
age_ca
```

###

Again the data is **long**: each county has two rows, one for `median_age` and one for `population`. To plot age against population — or even just to sort counties by population — you need the two measures side by side, each in its own column. **Reshaping** the data from long to wide does exactly that: it turns the repeated variable/value pairs into one column per variable. It is one of the most common moves in data preparation, precisely because so much analysis and plotting assumes the wide shape.

### Exercise 3

Change your working chunk so it reshapes `age_ca` to one row per county (a column for `median_age` and one for `population`) and draws a bar chart of median age by county. Render and look at it.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r california-age-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

Our rough chart:

```{r california-age-3-test}
age_ca |>
  select(NAME, variable, estimate) |>
  pivot_wider(names_from = variable, values_from = estimate) |>
  ggplot(aes(x = reorder(NAME, median_age), y = median_age)) +
  geom_col(fill = "#4daf4a") +
  coord_flip()
```

###

All 58 counties at once is too many bars to read — the labels collide and no county stands out. Turning the bars horizontal at least makes the county names legible, but the chart is still begging to be narrowed to a meaningful subset. That is the next move.

### Exercise 4

Improve the chart. Keep only the 15 most populous counties, sort the bars by median age, and add a title, clean axis labels, and a source caption. Render and compare to ours.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r california-age-4}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

Our improved chart:

```{r california-age-4-test}
age_ca |>
  select(NAME, variable, estimate) |>
  pivot_wider(names_from = variable, values_from = estimate) |>
  arrange(desc(population)) |>
  slice_head(n = 15) |>
  ggplot(aes(x = reorder(NAME, median_age), y = median_age)) +
  geom_col(fill = "#4daf4a") +
  coord_flip() +
  labs(
    title = "Median age, 15 most populous California counties (2020)",
    x = NULL,
    y = "Median age",
    caption = "Source: ACS 5-year estimates via tidycensus"
  ) +
  theme_minimal()
```

###

Filtering to the 15 most populous counties is an analytical choice, not just a cosmetic one: it focuses the chart on where most Californians actually live. Sorting the bars by median age makes the comparison immediate — an unordered bar chart forces the reader to hunt — and dropping the redundant axis label keeps the focus on the counties themselves.

### Exercise 5

In one or two sentences, describe what the chart shows. Among the most populous counties, which skew older and which skew younger?

```{r california-age-5}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

Your answer might note that even across the largest counties the spread in median age is only several years, with college-heavy and immigrant-heavy counties skewing somewhat younger and others a few years older — a reminder that median age varies far less between places than income or education does.

### Exercise 6

Commit your work. Stage `analysis.qmd`, commit with the message `"Add California age chart"`, and push to GitHub.

CP/CR, or write a short note describing how you committed and pushed.

```{r california-age-6}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

Median age is itself an ACS estimate with a margin of error, just like income. For large counties the margin is tiny, but for the smallest counties — which you filtered out here — a difference of a year or two between counties may be inside the margin and not real.

## Summary
###

This tutorial followed the **tidycensus** chapters of [Analyzing US Census Data](https://walker-data.com/census-r/) by Kyle Walker. You pulled American Community Survey data through the Census API with **tidycensus**, learned how ACS estimates carry margins of error, reshaped long-format results into the wide shape that plotting needs, and built three published artifacts: choropleth maps of Texas household income and California educational attainment, and a bar chart of California median age.

### Exercise 1

Run a final `quarto render analysis.qmd` and confirm all three figures appear in your Live Server tab. In the R Terminal, run:

```
show_file("analysis.qmd")
```

CP/CR.

```{r summary-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 20)
```

###

**tidycensus** accepts a state as a name (`"Texas"`), a postal code (`"TX"`), or a FIPS code (`"48"`), and will aggregate to whatever geographic level you ask for — county here, but also census tract, place, or state. The same request scales from a single county map to a national tract-level analysis without changing the approach.

### Exercise 2

Publish your work to GitHub Pages. In a bash Terminal, run:

```
quarto publish gh-pages analysis.qmd
```

If a render is still running, stop it first with `Ctrl/Cmd + C`. Copy and paste the resulting URL below.

```{r summary-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 1)
```

###

Publishing turns your `analysis.qmd` into a public page anyone can view — the same way most real Census analyses reach their audience.

### Exercise 3

Commit and push any remaining changes. Copy and paste the URL of your GitHub repo below.

```{r summary-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

The workflow you used here — **tidycensus** to pull ACS data (with boundaries when you want a map), **sf** to hold the geometry, and **ggplot2** to draw it — is the standard modern approach to US Census analysis. From here, Walker's book goes on to time-series comparisons, spatial statistics, and modeling, all built on the same foundation.

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
