---
title: Analyzing US Census Data
author: David Kane and Alec Rogers
tutorial:
  id: census
output:
  learnr::tutorial:
    df_print: default
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: "Companion to the tidycensus chapters of 'Analyzing US Census Data' by Kyle Walker. Pulls American Community Survey data with tidycensus and builds three published artifacts: a choropleth map of Texas household income, a ranked chart of state educational attainment, and a scatterplot of California county population and median age."
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(tidyverse)
library(tidycensus)
library(sf)
library(knitr)

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(out.width = '90%')
options(tutorial.exercise.timelimit = 600,
        tutorial.storage = "local")

# The data below was pulled once with tidycensus and saved, so this tutorial
# renders without hitting the Census API. Students make the live calls in their
# own analysis.qmd.

# income_tx <- get_acs(geography = "county", variables = "B19013_001",
# state = "TX", year = 2020, geometry = TRUE)
# write_rds(income_tx, "data/income_tx.rds")
income_tx <- read_rds("data/income_tx.rds")

# edu_state <- get_acs(geography = "state",
# variables = c("B15003_001", "B15003_022", "B15003_023", "B15003_024", "B15003_025"),
# year = 2020, geometry = FALSE, summary_var = "B15003_001")
# write_rds(edu_state, "data/edu_state.rds")
edu_state <- read_rds("data/edu_state.rds")

# age_ca <- get_acs(geography = "county",
# variables = c(median_age = "B01002_001", population = "B01003_001"),
# state = "CA", year = 2020, geometry = FALSE)
# write_rds(age_ca, "data/age_ca.rds")
age_ca <- read_rds("data/age_ca.rds")
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

## Introduction
###

This tutorial follows Kyle Walker's [An introduction to tidycensus](https://walker-data.com/census-r/an-introduction-to-tidycensus.html) and [Mapping Census data with R](https://walker-data.com/census-r/mapping-census-data-with-r.html) chapters from [Analyzing US Census Data](https://walker-data.com/census-r/). You will pull data from the American Community Survey through the Census API and turn it into three published artifacts: a choropleth map of Texas household income, a ranked chart of state educational attainment, and a scatterplot of California county population and median age.

Use AI however you like throughout. An agentic coding tool (such as Claude Code or Gemini CLI) works best, since it can read and edit your files directly; a chat-based assistant works too, with more copy/paste.

### Exercise 1

You should be doing this tutorial in a repo named `census`. If you are not, create one and then connect to it. You may need to restart this tutorial after you do so. Create a new file and save it as `analysis.qmd`. Add a YAML header with a title (`"Analyzing US Census Data"`) and your name as author. In a bash Terminal, run:

```
quarto render analysis.qmd
```

Open `analysis.html` by right-clicking it in the Explorer and selecting **Open with Live Server**. Keep this tab open — it refreshes automatically whenever you re-render. Going forward, we will just tell you to "Render" when we want you to take these steps.

Create a `.gitignore` with `analysis_files` followed by a blank line. Commit and push.

In the R Terminal, run:

```
show_file(".gitignore")
```

If that fails, it is probably because you have not yet loaded `library(tutorial.helpers)` in the R Terminal.

CP/CR.

```{r introduction-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

```
analysis_files
```

###

The American Community Survey (ACS) is the Census Bureau's rolling annual survey of about 3.5 million households. Unlike the Decennial Census, which counts everyone every ten years, the ACS estimates characteristics like income, education, and age from a sample — which is why every value arrives with a margin of error.

### Exercise 2

In `analysis.qmd`, add a code chunk that loads the three libraries you will need:

```
library(tidyverse)
library(tidycensus)
library(sf)
```

Rendering shows the code and the package start-up messages, which we don't want in a published document. Add `#| message: false` to the chunk, and add this to the YAML header to hide all code:

```
execute:
  echo: false
```

Render again. In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r introduction-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

<pre><code>#| message: false
library(tidyverse)
library(tidycensus)
library(sf)</code></pre>

###

**tidycensus** returns ACS data as a tidy data frame, one row per geography (or per geography-variable). If you also ask for the geographic boundaries, it hands you an **[sf](https://r-spatial.github.io/sf/)** object — a table with a spatial column attached — that **[ggplot2](https://ggplot2.tidyverse.org/)** can map directly. **sf** is the standard package for spatial data in R, and the boundaries themselves come from **[tigris](https://github.com/walkerke/tigris)**, which **tidycensus** uses under the hood.

### Exercise 3

Every call to `get_acs()` needs a Census API key. Sign up for a free one at the [Census API key signup page](https://api.census.gov/data/key_signup.html) — the key arrives by email within a few minutes.

In the R Terminal, run (with your own key in place of `"YOUR-KEY-HERE"`):

```
tidycensus::census_api_key("YOUR-KEY-HERE", install = TRUE)
```

This stores the key in your `.Renviron` file so every future R session finds it automatically. In the bash Terminal, run:

```
cat ~/.Renviron
```

You should see a line that starts with `CENSUS_API_KEY=`. Do not paste that output anywhere, because it contains your key.

Restart R (Session > Restart R, or close and reopen your R Terminal), then confirm the key loaded:

```
nzchar(Sys.getenv("CENSUS_API_KEY"))
```

Run only the second command for evidence. Never paste the key itself anywhere.

CP/CR.

```{r introduction-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

```
[1] TRUE
```

###

The key is tied to your email, not a Census account — there is no login to manage. Because `install = TRUE` writes it to `.Renviron` rather than `analysis.qmd`, it never appears in your code or your commits, and every `get_acs()` call in this tutorial picks it up automatically from here on.

## Texas income
###

Median household income is one of the most-requested ACS variables. Its code is `B19013_001`. In this section you will map it across all 254 Texas counties.

### Exercise 1

Add a chunk to `analysis.qmd` that uses **tidycensus** to get 2020 median household income for every county in Texas, with county geometry attached for mapping, and saves it as `income_tx`. Add `#| message: false` to the chunk to hide the connection messages. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r texas-income-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

<pre><code>#| message: false
income_tx <- get_acs(
  geography = "county",
  variables = "B19013_001",
  state = "TX",
  year = 2020,
  geometry = TRUE
)
</code></pre>

###

Every ACS request comes down to three choices: the **geographic level** (county here, but census tract, place, or state are equally available), the **variables** — Census table codes like `B19013_001` for median household income — and the **year**, which for the ACS names a 5-year window (2020 means the 2016–2020 estimates, pooled across five years so even small counties have reliable numbers). The first render makes a real network call to the Census API, so it takes a moment — and draws on the key you installed in the Introduction, which every `get_acs()` call in this tutorial uses automatically.

### Exercise 2

Add a chunk that prints `income_tx` so you can see its structure. Render. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r texas-income-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```{r texas-income-2-test}
#| echo: true
income_tx
```

###

Every county arrives with a `GEOID` (its FIPS code), a `NAME`, the `variable` code, an `estimate`, and a margin of error (`moe`). The `estimate` is the survey's best guess; the `moe` is the half-width of its 90% confidence interval. Because the ACS is a *sample*, not a full count, every estimate carries a margin — and that margin is smaller in populous counties, where the survey reaches more households. Because you asked for boundaries, there is also a spatial column holding each county's polygon.

### Exercise 3

Replace the working chunk with code that shows the structure of `income_tx`. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r texas-income-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```{r texas-income-3-test}
#| echo: true
glimpse(income_tx)
```

###

This is a good place for `glimpse()` because the table is narrow. The `geometry` column is the giveaway that `income_tx` is an **sf** object: it is still a table, but each row also carries the polygon needed to draw that county on a map.

### Exercise 4

Replace the working chunk with code that summarizes median household income and margin of error across the Texas counties. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r texas-income-4}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 7)
```

###

```{r texas-income-4-test}
#| echo: true
income_tx |>
  st_drop_geometry() |>
  summarize(
    min_income = min(estimate),
    median_income = median(estimate),
    max_income = max(estimate),
    median_moe = median(moe)
  )
```

###

The income spread is large enough that a map should be interesting: Texas counties are not all clustered around one value. Dropping the geometry before a numeric summary is a useful habit when the geometry is not part of the calculation.

### Exercise 5

Replace the working chunk with a histogram of county median household income. Render and look at the shape of the distribution.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r texas-income-5}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r texas-income-5-test}
#| echo: true
#| message: false
ggplot(income_tx, aes(x = estimate)) +
  geom_histogram(bins = 30, fill = "#2b8cbe", color = "white") +
  labs(
    title = "Distribution of median household income across Texas counties",
    x = "Median household income",
    y = "Number of counties"
  ) +
  theme_minimal()
```

###

The histogram shows the distribution but not the geography. A few high-income counties sit well above most of the state, and the next question is whether those counties are scattered randomly or clustered around particular metro areas.

### Exercise 6

Replace the histogram with a choropleth map of median income across the counties using **ggplot2**. Render and check the map in your Live Server tab.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r texas-income-6}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```{r texas-income-6-test}
#| echo: true
#| message: false
ggplot(income_tx) +
  geom_sf(aes(fill = estimate))
```

###

This is a **choropleth** — each county drawn from its boundary and shaded by its income. Two things make this first version hard to read: the default blue gradient runs dark-to-light in a direction many readers misjudge, and the gray panel with latitude/longitude gridlines adds clutter a map does not need. Fixing both is the next step.

### Exercise 7

Improve the map: give it a perceptually uniform color scale (the **[viridis](https://sjmgarnier.github.io/viridis/)** family is the standard choice), a clear title and legend label, a caption naming the data source, and a clean map theme. Render and compare to ours.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r texas-income-7}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r texas-income-7-test}
#| echo: true
#| message: false
ggplot(income_tx) +
  geom_sf(aes(fill = estimate), color = "white", size = 0.1) +
  scale_fill_viridis_c(option = "plasma", name = "Median income") +
  labs(
    title = "Median household income by county, Texas (2020)",
    caption = "Source: ACS 5-year estimates via tidycensus"
  ) +
  theme_void()
```

###

The **viridis** color scales are designed to be perceptually uniform and readable by colorblind viewers, which is why they have become a default for choropleths; **ggplot2** ships them, and the **[RColorBrewer](https://colorbrewer2.org/)** palettes are the other common choice. Clearing away the gray panel and the latitude/longitude gridlines a map does not need leaves the counties to carry the message. Reading the finished map, the highest incomes cluster in the suburban counties around the big metros — Austin, Dallas–Fort Worth, and Houston — while many rural counties, especially along the Rio Grande border in the south, carry the lowest median incomes. A published choropleth earns its place by making that pattern legible at a glance: the map tells the reader *where*, and the title and caption tell them *what to take away* and what the numbers can support.

### Exercise 8

Commit your work. Stage `analysis.qmd`, commit with the message `"Add Texas income map"`, and push to GitHub. You can do this with your AI agent, the VS Code Source Control panel, or `git` on the command line.

CP/CR, or write a short note describing how you committed and pushed.

```{r texas-income-8}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

A choropleth shades each area by a value, so it is only as honest as the variable behind it. Median household income is a good fit because it is roughly comparable across counties; a raw count like "number of high earners" would mostly track county population and mislead.

## Education by state
###

Educational attainment lives in ACS table B15003, which breaks the adult population into 25 completed schooling levels from "no schooling" up to "doctorate." In this section you will compare the share of adults with at least a bachelor's degree across all state-level geographies.

### Exercise 1

Add a chunk that uses **tidycensus** to get 2020 educational attainment for state-level geographies and save it as `edu_state`. You need the table total plus the bachelor's-and-higher levels: variables `B15003_001` (total), `B15003_022`, `B15003_023`, `B15003_024`, and `B15003_025`. Pass `summary_var = "B15003_001"` so the total rides along on every row. Add `#| message: false`. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r education-by-state-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 7)
```

###

<pre><code>#| message: false
edu_state <- get_acs(
  geography = "state",
  variables = c("B15003_001", "B15003_022", "B15003_023", "B15003_024", "B15003_025"),
  year = 2020,
  geometry = FALSE,
  summary_var = "B15003_001"
)
</code></pre>

###

These five variable codes report *completed* degrees, not enrollment. `B15003_022` through `B15003_025` are bachelor's, master's, professional, and doctorate degrees — four separate, non-overlapping levels — so adding them counts everyone with at least a bachelor's. Asking for the table total (`B15003_001`) to ride along on every row gives you the denominator you need to turn those counts into a percentage.

### Exercise 2

Add a working chunk that prints `edu_state`. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r education-by-state-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```{r education-by-state-2-test}
#| echo: true
edu_state
```

###

The result is **long**: each state-level geography appears once for each requested variable. With 52 geographies and five variables, `edu_state` has 260 rows. **tidycensus** includes the 50 states, the District of Columbia, and Puerto Rico when you request `geography = "state"`.

### Exercise 3

Replace the working chunk with code that shows the structure of `edu_state`. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r education-by-state-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```{r education-by-state-3-test}
#| echo: true
glimpse(edu_state)
```

###

This is a good place for `glimpse()` because the table is narrow. The important columns are `estimate`, the count for one education level, and `summary_est`, the repeated total adult population for that geography. The margin-of-error columns remind you that ACS values are estimates, even when the table looks like ordinary counts.

### Exercise 4

Replace the working chunk with a summary by `variable` that reports how many rows each variable has and the smallest and largest estimates. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r education-by-state-4}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r education-by-state-4-test}
#| echo: true
edu_state |>
  group_by(variable) |>
  summarize(
    rows = n(),
    min_estimate = min(estimate),
    max_estimate = max(estimate)
  )
```

###

Each variable has 52 rows, one for each state-level geography. The total adult population is much larger than any one degree category, which is the clue that raw counts are not the final measure: California will dominate raw counts because it has so many adults, not necessarily because it has the highest share with degrees.

### Exercise 5

You are done changing the chunk that downloads `edu_state`, and re-downloading it on every render is slow. `edu_state` earns its own cached create chunk. The next three exercises set that up: clean the chunk, turn caching on, add the cache directory to `.gitignore`.

First, clean up. Delete the working chunk that summarized `edu_state`, leaving the `edu_state <- get_acs(...)` chunk as the last chunk in the file. Render. In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r education-by-state-5}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 10)
```

###

<pre><code>#| message: false
edu_state <- get_acs(
  geography = "state",
  variables = c("B15003_001", "B15003_022", "B15003_023", "B15003_024", "B15003_025"),
  year = 2020,
  geometry = FALSE,
  summary_var = "B15003_001"
)
</code></pre>

###

A chunk ready to cache should contain only the artifact it builds. The exploratory chunk did its job: it helped you understand `edu_state`, but it does not need to remain in the finished QMD.

### Exercise 6

Add `#| cache: true` as the second line inside the chunk, just below `#| message: false`. Render the QMD. Rendering creates an `analysis_cache/` directory next to your `analysis.qmd`, which holds the saved tibble so Quarto can reload it on subsequent renders rather than re-downloading from the Census API. To confirm the directory is there, run `ls` in the **bash Terminal**. CP/CR.

```{r education-by-state-6}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

<pre><code>analysis.qmd  analysis_cache  analysis_files  analysis.html
</code></pre>

###

Caching is for finished, expensive chunks — a network request like this one, not code you are still editing. The cache is generated output: useful locally, but not something to publish to GitHub.

### Exercise 7

The Source Control change count just jumped, because `analysis_cache/` and everything inside it now look like untracked work to git. Cached files should never go to GitHub.

Add `analysis_cache` to your `.gitignore` on its own line. In the R Terminal, run `show_file(".gitignore")`. CP/CR.

```{r education-by-state-7}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

<pre><code>analysis_files
analysis_cache
</code></pre>

###

The Source Control change count should drop back to whatever it was before the render. This is the same discipline as ignoring `analysis_files`: generated output can help your local workflow without becoming part of the source project.

### Exercise 8

Add a new working chunk that collapses `edu_state` to one row per geography. Compute the percentage of adults with at least a bachelor's degree, then print the table sorted from highest to lowest percentage. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r education-by-state-8}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r education-by-state-8-test}
#| echo: true
edu_state |>
  group_by(GEOID, NAME) |>
  summarize(
    percent = 100 * sum(estimate[variable != "B15003_001"]) / unique(summary_est),
    adults = unique(summary_est)
  ) |>
  arrange(desc(percent))
```

###

The District of Columbia sits far above the states because it is a city-like jurisdiction with a highly educated labor market, not a large mixed urban-rural state. That is not a mistake, but it is a reminder to read geographies carefully before comparing them.

### Exercise 9

Replace the working chunk with code that summarizes the distribution of the state-level bachelor's-or-higher percentages. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r education-by-state-9}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r education-by-state-9-test}
#| echo: true
edu_state |>
  group_by(GEOID, NAME) |>
  summarize(
    percent = 100 * sum(estimate[variable != "B15003_001"]) / unique(summary_est),
    .groups = "drop"
  ) |>
  pull(percent) |>
  summary()
```

###

The middle state-level geography is around 31%, while the minimum is about 21% and the maximum is about 60%. A summary gives the range, but a plot makes the shape of the distribution easier to judge.

### Exercise 10

Replace the working chunk with a histogram of the state-level bachelor's-or-higher percentages. Render and look at the shape of the distribution.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r education-by-state-10}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r education-by-state-10-test}
#| echo: true

edu_state |>
  group_by(GEOID, NAME) |>
  summarize(
    percent = 100 * sum(estimate[variable != "B15003_001"]) / unique(summary_est),
    .groups = "drop"
  ) |>
  ggplot(aes(x = percent)) +
  geom_histogram(bins = 15, fill = "#756bb1", color = "white") +
  scale_x_continuous(labels = scales::label_percent(scale = 1)) +
  labs(
    title = "Distribution of bachelor's-or-higher rates across state-level geographies",
    x = "Share of adults 25 and over",
    y = "Number of geographies"
  ) +
  theme_minimal()
```

###

The histogram shows that most state-level geographies cluster in a fairly narrow middle range, with the District of Columbia far to the right. To identify every place, not just the distribution, the next plot needs to rank the geographies by name.

### Exercise 11

Replace the working chunk with a horizontal bar chart of the bachelor's-or-higher percentage for every state-level geography, sorted by percentage. Render and look at it.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r education-by-state-11}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 9)
```

###

```{r education-by-state-11-test}
#| echo: true
edu_state |>
  group_by(GEOID, NAME) |>
  summarize(
    percent = 100 * sum(estimate[variable != "B15003_001"]) / unique(summary_est),
    .groups = "drop"
  ) |>
  ggplot(aes(x = reorder(NAME, percent), y = percent)) +
  geom_col() +
  coord_flip()
```

###

Ordering the bars turns the distribution into a comparison. The rough chart shows the main ranking immediately, but it still needs labels, a readable percent scale, and a source note before it is worth publishing.

### Exercise 12

Improve the chart: add a clear title, axis label, source caption, percent-formatted axis labels, and a cleaner theme. Render and compare to ours.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r education-by-state-12}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 9)
```

###

```{r education-by-state-12-test}
#| echo: true
edu_state |>
  group_by(GEOID, NAME) |>
  summarize(
    percent = 100 * sum(estimate[variable != "B15003_001"]) / unique(summary_est),
    .groups = "drop"
  ) |>
  ggplot(aes(x = reorder(NAME, percent), y = percent, fill = percent)) +
  geom_col() +
  coord_flip() +
  scale_y_continuous(labels = scales::label_percent(scale = 1)) +
  scale_fill_viridis_c(option = "C", guide = "none") +
  labs(
    title = "Adults with a bachelor's degree or higher by state-level geography (2020)",
    x = NULL,
    y = "Share of adults 25 and over",
    caption = "Source: ACS 5-year estimates via tidycensus"
  ) +
  theme_minimal()
```

###

The finished chart shows a regional and institutional pattern that a table makes hard to scan: the highest values cluster in the District of Columbia and several Northeastern or knowledge-economy states, while the lowest values are concentrated in parts of Appalachia and the South. Because this is a percentage of adults 25 and over, the comparison is not just a population-size ranking.

### Exercise 13

Commit your work. Stage `analysis.qmd`, commit with the message `"Add state education chart"`, and push to GitHub.

CP/CR, or write a short note describing how you committed and pushed.

```{r education-by-state-13}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

This section used the same table as a county-level education map would use, but changed the geography to states and territories. That is one of **tidycensus**'s central ideas: the variables define what you measure, and the geography defines the units you compare.

## California median age
###

Not every Census question is a map. Here you will pull median age and total population for California counties, then ask whether larger counties tend to be younger or older. This section also revisits the reshaping step that long-format ACS data often needs.

### Exercise 1

Add a chunk that uses **tidycensus** to get 2020 median age and total population for every county in California, naming the variables `median_age` (`B01002_001`) and `population` (`B01003_001`). You are building a non-map chart, so set `geometry = FALSE`. Save it as `age_ca` and add `#| message: false`. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r california-median-age-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

<pre><code>#| message: false
age_ca <- get_acs(
  geography = "county",
  variables = c(median_age = "B01002_001", population = "B01003_001"),
  state = "CA",
  year = 2020,
  geometry = FALSE
)
</code></pre>

###

You can give each variable a readable name — `median_age`, `population` — instead of carrying raw codes like `B01002_001` around, which makes everything downstream easier to read. And because this section does not draw boundaries, skipping the geometry download is the right call — there is no reason to pull polygons you will not use.

### Exercise 2

Add a working chunk that prints `age_ca`. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r california-median-age-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```{r california-median-age-2-test}
#| echo: true
age_ca
```

###

Again the data is **long**: each county has two rows, one for `median_age` and one for `population`. That shape is natural for the Census API response, but not for a scatterplot comparing the two measures.

### Exercise 3

Replace the working chunk with code that shows the structure of `age_ca`. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r california-median-age-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

```{r california-median-age-3-test}
#| echo: true
glimpse(age_ca)
```

###

The table is narrow enough that `glimpse()` is useful. The `variable` column tells you which measure a row holds, while `estimate` holds the value. To compare median age with population, those two variable values need to become two columns.

### Exercise 4

Replace the working chunk with a summary by `variable` that reports how many rows each variable has and the smallest and largest estimates. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r california-median-age-4}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r california-median-age-4-test}
#| echo: true
age_ca |>
  group_by(variable) |>
  summarize(
    rows = n(),
    min_estimate = min(estimate),
    max_estimate = max(estimate)
  )
```

###

Each variable has 58 rows, one for each California county. The two variables are on very different scales: population ranges from tiny Alpine County to Los Angeles County, while median age varies by a few decades. A scatterplot can show both scales at once.

### Exercise 5

Replace the working chunk with code that reshapes `age_ca` to one row per county, with one column for `median_age` and one for `population`. Print the table sorted from largest to smallest population. Render.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r california-median-age-5}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 7)
```

###

```{r california-median-age-5-test}
#| echo: true
age_ca |>
  select(NAME, variable, estimate) |>
  pivot_wider(names_from = variable, values_from = estimate) |>
  arrange(desc(population))
```

###

**Reshaping** turns repeated variable/value rows into one column per variable. That wide shape is what most plots and models expect: one row per county, with `median_age` and `population` available side by side.

### Exercise 6

Replace the working chunk with a histogram of median age across California counties. Render and look at the shape of the distribution.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r california-median-age-6}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r california-median-age-6-test}
#| echo: true
#| message: false
age_ca |>
  select(NAME, variable, estimate) |>
  pivot_wider(names_from = variable, values_from = estimate) |>
  ggplot(aes(x = median_age)) +
  geom_histogram(bins = 15, fill = "#41ab5d", color = "white") +
  labs(
    title = "Distribution of median age across California counties",
    x = "Median age",
    y = "Number of counties"
  ) +
  theme_minimal()
```

###

The histogram shows that most counties sit in the mid-30s to mid-40s, with a few much older counties. It still does not tell whether those older counties are large population centers or smaller rural counties.

### Exercise 7

Replace the working chunk with a scatterplot of county population against median age. Render and look at the relationship.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r california-median-age-7}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r california-median-age-7-test}
#| echo: true
#| message: false
age_ca |>
  select(NAME, variable, estimate) |>
  pivot_wider(names_from = variable, values_from = estimate) |>
  ggplot(aes(x = population, y = median_age)) +
  geom_point()
```

###

The rough scatterplot has the right variables, but the population axis is hard to read because Los Angeles County is so much larger than most counties. A log scale is a better fit when one variable spans several orders of magnitude.

### Exercise 8

Improve the scatterplot: use a log-scaled population axis, label the largest counties, add a title, clean axis labels, and a source caption. Render and compare to ours.

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r california-median-age-8}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 10)
```

###

```{r california-median-age-8-test}
#| echo: true
#| message: false
age_ca |>
  select(NAME, variable, estimate) |>
  pivot_wider(names_from = variable, values_from = estimate) |>
  mutate(label = if_else(min_rank(desc(population)) <= 8, NAME, "")) |>
  ggplot(aes(x = population, y = median_age)) +
  geom_point(color = "#2c7fb8", alpha = 0.75, size = 2.5) +
  geom_text(aes(label = label), hjust = -0.05, check_overlap = TRUE, size = 3) +
  scale_x_log10(
    labels = scales::label_number(scale_cut = scales::cut_short_scale()),
    expand = expansion(mult = c(0.02, 0.25))
  ) +
  labs(
    title = "Smaller California counties tend to be older (2020)",
    x = "Population (log scale)",
    y = "Median age",
    caption = "Source: ACS 5-year estimates via tidycensus"
  ) +
  theme_minimal()
```

###

The final plot shows a negative relationship: California's largest counties are mostly in the 30s, while many of the oldest counties are small. The log scale keeps Los Angeles County visible without flattening every smaller county into the left edge of the chart.

### Exercise 9

Commit your work. Stage `analysis.qmd`, commit with the message `"Add California age scatterplot"`, and push to GitHub.

CP/CR, or write a short note describing how you committed and pushed.

```{r california-median-age-9}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

Median age is itself an ACS estimate with a margin of error, just like income. For large counties the margin is tiny, but for the smallest counties a difference of a year or two between counties may be inside the margin and not real.

## Summary
###

This tutorial followed the **tidycensus** chapters of [Analyzing US Census Data](https://walker-data.com/census-r/) by Kyle Walker. You pulled American Community Survey data through the Census API with **tidycensus**, learned how ACS estimates carry margins of error, reshaped long-format results into the wide shape that plotting needs, and built three published artifacts: a choropleth map of Texas household income, a ranked chart of state educational attainment, and a scatterplot of California county population and median age.

### Exercise 1

Render and confirm all three figures appear. In the R Terminal, run:

```
show_file("analysis.qmd")
```

CP/CR.

```{r summary-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 25)
```

###

<pre><code>---
title: "Analyzing US Census Data"
author: "&lt;your name&gt;"
execute:
  echo: false
---

&#96;&#96;&#96;{r}
#| message: false
library(tidyverse)
library(tidycensus)
library(sf)
&#96;&#96;&#96;

&#96;&#96;&#96;{r}
#| cache: true
income_tx <- get_acs(
  geography = "county",
  state = "TX",
  variables = "B19013_001",
  year = 2020,
  geometry = TRUE
)
&#96;&#96;&#96;

&#96;&#96;&#96;{r}
#| message: false
ggplot(income_tx) +
  geom_sf(aes(fill = estimate), color = "white", size = 0.1) +
  scale_fill_viridis_c(option = "plasma", name = "Median income") +
  labs(
    title = "Median household income by county, Texas (2020)",
    caption = "Source: ACS 5-year estimates via tidycensus"
  ) +
  theme_void()
&#96;&#96;&#96;

&#96;&#96;&#96;{r}
#| message: false
#| cache: true
edu_state <- get_acs(
  geography = "state",
  variables = c("B15003_001", "B15003_022", "B15003_023", "B15003_024", "B15003_025"),
  year = 2020,
  geometry = FALSE,
  summary_var = "B15003_001"
)
&#96;&#96;&#96;

&#96;&#96;&#96;{r}
edu_state |>
  group_by(GEOID, NAME) |>
  summarize(
    percent = 100 * sum(estimate[variable != "B15003_001"]) / unique(summary_est),
    .groups = "drop"
  ) |>
  ggplot(aes(x = reorder(NAME, percent), y = percent, fill = percent)) +
  geom_col() +
  coord_flip() +
  scale_y_continuous(labels = scales::label_percent(scale = 1)) +
  scale_fill_viridis_c(option = "C", guide = "none") +
  labs(
    title = "Adults with a bachelor's degree or higher by state-level geography (2020)",
    x = NULL,
    y = "Share of adults 25 and over",
    caption = "Source: ACS 5-year estimates via tidycensus"
  ) +
  theme_minimal()
&#96;&#96;&#96;

&#96;&#96;&#96;{r}
#| message: false
age_ca <- get_acs(
  geography = "county",
  state = "CA",
  variables = c(median_age = "B01002_001", population = "B01003_001"),
  year = 2020,
  geometry = FALSE
)
&#96;&#96;&#96;

&#96;&#96;&#96;{r}
age_ca |>
  select(NAME, variable, estimate) |>
  pivot_wider(names_from = variable, values_from = estimate) |>
  mutate(label = if_else(min_rank(desc(population)) <= 8, NAME, "")) |>
  ggplot(aes(x = population, y = median_age)) +
  geom_point(color = "#2c7fb8", alpha = 0.75, size = 2.5) +
  geom_text(aes(label = label), hjust = -0.05, check_overlap = TRUE, size = 3) +
  scale_x_log10(
    labels = scales::label_number(scale_cut = scales::cut_short_scale()),
    expand = expansion(mult = c(0.02, 0.25))
  ) +
  labs(
    title = "Smaller California counties tend to be older (2020)",
    x = "Population (log scale)",
    y = "Median age",
    caption = "Source: ACS 5-year estimates via tidycensus"
  ) +
  theme_minimal()
&#96;&#96;&#96;
</code></pre>

###

**tidycensus** accepts a state as a name (`"Texas"`), a postal code (`"TX"`), or a FIPS code (`"48"`), and will aggregate to whatever geographic level you ask for — county here, but also census tract, place, or state. The same request scales from a single county map to a national tract-level analysis without changing the approach.

### Exercise 2

Publish your work to GitHub Pages. In a bash Terminal, run:

```
quarto publish gh-pages analysis.qmd
```

If a render is still running, stop it first with `Ctrl/Cmd + C`. Copy and paste the resulting URL below.

```{r summary-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 1)
```

###

Publishing turns your `analysis.qmd` into a public page anyone can view — the same way most real Census analyses reach their audience.

### Exercise 3

Commit and push any remaining changes. Copy and paste the URL of your GitHub repo below.

```{r summary-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

The workflow you used here — **tidycensus** to pull ACS data (with boundaries when you want a map), **sf** to hold the geometry, and **ggplot2** to draw it — is the standard modern approach to US Census analysis. From here, Walker's book goes on to time-series comparisons, spatial statistics, and modeling, all built on the same foundation.

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
