---
title: DGM Logistic
author: Var Kurapati
tutorial:
  id: dgm-logistic
  version: 1.0.0
output:
  learnr::tutorial:
    df_print: default
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: 'Learn to fit and interpret logistic regression models using tidymodels and marginaleffects, applied to employee attrition.'
---

```{r setup, include=FALSE}
library(learnr)
library(tutorial.helpers)
library(knitr)
library(tidyverse)
library(tidymodels)
library(marginaleffects)

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(out.width = '90%')
options(tutorial.exercise.timelimit = 600,
        tutorial.storage = "local")

mod_logistic <- logistic_reg() |>
  set_engine("glm") |>
  fit(Attrition ~ MonthlyIncome + YearsAtCompany + OverTime, data = attrition)
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

## Introduction
###

Of 1,470 employees at a company, 237 left — an attrition rate of about 16%. The `attrition` dataset records, for each employee, whether they left and a set of features about their job: income, tenure, whether they work overtime, and more.

This tutorial asks a simple question: **can we predict whether an employee will leave from their income, tenure, and overtime status?**

We use the same two packages as before. **[tidymodels](https://www.tidymodels.org/)** fits the model. **[marginaleffects](https://marginaleffects.com/)** interprets it — turning coefficients into predicted probabilities.

### Exercise 1

Create a repo named `dgm-logistic` and connect to it. Create `analysis.qmd`, title `"Employee Attrition"`, your name as author, `execute: echo: false` in the YAML. Render, open with Live Server, create `.gitignore` with `analysis_files`, commit and push.

In the R Terminal, run:

```
show_file(".gitignore")
```

CP/CR.

```{r introduction-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

```
analysis_files
```

###

### Exercise 2

In a new code chunk, load the libraries you'll need. Add `#| message: false`. Render.

```
#| message: false
library(tidyverse)
library(tidymodels)
library(marginaleffects)
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r introduction-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 10)
```

###

<pre><code>#| message: false
library(tidyverse)
library(tidymodels)
library(marginaleffects)
</code></pre>

###

Same three libraries as last time, and the same reason: `tidymodels` brings in everything needed to fit and tidy a model, `marginaleffects` interprets it, `tidyverse` covers plotting and data wrangling. `attrition` is available directly once `tidymodels` is loaded, the same way `ames` was.

### Exercise 3

In a new code chunk, print the first 10 rows of `attrition`. Render.

```
attrition |>
  select(Attrition, MonthlyIncome, YearsAtCompany, OverTime) |>
  head(10)
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r introduction-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 4)
```

###

```{r introduction-3-test}
#| echo: true
attrition |>
  select(Attrition, MonthlyIncome, YearsAtCompany, OverTime) |>
  head(10)
```

###

`attrition` has 31 columns. This tutorial uses four: `Attrition` (Yes/No — the outcome), `MonthlyIncome` (dollars), `YearsAtCompany`, and `OverTime` (Yes/No).

### Exercise 4

Ask AI to create a boxplot of `MonthlyIncome` by `Attrition`. Paste its code into the chunk. Render.

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r introduction-4}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 10)
```

###

Here's our version, for comparison — yours doesn't need to match exactly, just the same idea:

```{r introduction-4-test}
#| echo: true
attrition |>
  ggplot(aes(x = Attrition, y = MonthlyIncome)) +
  geom_boxplot() +
  labs(
    title = "Employees who left tend to earn less",
    x = "Attrition",
    y = "Monthly income ($)"
  ) +
  scale_y_continuous(labels = scales::dollar)
```

###

Employees who left earn less, on average, than employees who stayed — a visible gap between the two boxes. `Attrition` is binary, which rules out a linear model: sale price could in principle be any number, but attrition can only be Yes or No. That's exactly the kind of outcome logistic regression is built for.

### Exercise 5

Commit and push, with a message like "Add attrition data and initial exploration." In the bash Terminal, run:

```
git log -1
```

CP/CR.

```{r introduction-5}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```
commit d4e5f6a7b8c9...
Author: Var Kurapati
Date:   Mon Aug 03 2026

    Add attrition data and initial exploration
```

###

The `attrition` dataset is a snapshot of one company's employees at one point in time — not a random sample of every workplace. Whatever this model learns describes *this* company's patterns, not employment in general. Keep that in mind as we build a model in the next section.

## Tidymodels
###

`Attrition` is binary — Yes or No — so the structure here is different from a linear model. Instead of predicting a number directly, logistic regression predicts the **log-odds** of the outcome, using this general form:

$$\log\left(\frac{p}{1-p}\right) = \beta_0 + \beta_1 x_1 + \beta_2 x_2 + \cdots$$

where $p$ is the probability of the outcome (here, `Attrition = Yes`). The right-hand side looks just like linear regression — a sum of coefficients times variables — but the left-hand side transforms a probability (which must stay between 0 and 1) into something that can range over all real numbers. That transformation is what makes this a different structure from last time, even though the fitting steps — Structure → Method → Fitting — are the same.

There is no single "correct" DGM for a given question. Which right-hand-side variables to include is a judgment call — one we will make, and remake, several times in this section.

### Exercise 1

In a new code chunk, specify a logistic regression model with `logistic_reg()`. Print the result. Render.

```
logistic_reg()
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r tidymodels-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 4)
```

###

```{r tidymodels-1-test}
#| echo: true
logistic_reg()
```

###

`logistic_reg()` proposes the structure — a binary outcome predicted by a linear combination of predictors, on the log-odds scale. It hasn't touched data yet.

### Exercise 2

Update your chunk to pipe into `set_engine("glm")`. Render.

```
logistic_reg() |>
set_engine("glm")
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r tidymodels-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 4)
```

###

```{r tidymodels-2-test}
#| echo: true
logistic_reg() |>
  set_engine("glm")
```

###

`set_engine("glm")` picks the method: R's classical maximum-likelihood fitting for generalized linear models. It's a separate decision from the structure, same as `"lm"` was for linear regression.

### Exercise 3

Update your chunk to pipe into `fit(Attrition ~ MonthlyIncome, data = attrition)`. Render.

```
logistic_reg() |>
set_engine("glm") |>
fit(Attrition ~ MonthlyIncome, data = attrition)
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r tidymodels-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

```{r tidymodels-3-test}
#| echo: true
logistic_reg() |>
  set_engine("glm") |>
  fit(Attrition ~ MonthlyIncome, data = attrition)
```

###

`fit()` is where the mechanism meets real data. This chunk is now your **working chunk** — you'll keep editing this same formula over the next few exercises.

### Exercise 4

Update your working chunk to pipe the fitted model into `tidy(conf.int = TRUE)`, then `select(term, estimate, conf.low, conf.high)`. Render.

```
logistic_reg() |>
set_engine("glm") |>
fit(Attrition ~ MonthlyIncome, data = attrition) |>
tidy(conf.int = TRUE) |>
select(term, estimate, conf.low, conf.high)
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r tidymodels-4}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r tidymodels-4-test}
#| echo: true
logistic_reg() |>
  set_engine("glm") |>
  fit(Attrition ~ MonthlyIncome, data = attrition) |>
  tidy(conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high)
```

###

Same heuristic as before: no p-value, no test statistic. We keep a variable if its confidence interval excludes zero. `MonthlyIncome`'s does — its estimate is negative, meaning higher income is associated with lower log-odds of leaving.

### Exercise 5

Edit the formula in your working chunk to add `YearsAtCompany`. Render.

```
logistic_reg() |>
set_engine("glm") |>
fit(Attrition ~ MonthlyIncome + YearsAtCompany, data = attrition) |>
tidy(conf.int = TRUE) |>
select(term, estimate, conf.low, conf.high)
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r tidymodels-5}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r tidymodels-5-test}
#| echo: true
logistic_reg() |>
  set_engine("glm") |>
  fit(Attrition ~ MonthlyIncome + YearsAtCompany, data = attrition) |>
  tidy(conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high)
```

###

Both intervals exclude zero — `YearsAtCompany` earns its place. Longer-tenured employees are less likely to leave, holding income fixed.

### Exercise 6

Edit the formula again to add `Department`. Render.

```
logistic_reg() |>
set_engine("glm") |>
fit(Attrition ~ MonthlyIncome + YearsAtCompany + Department, data = attrition) |>
tidy(conf.int = TRUE) |>
select(term, estimate, conf.low, conf.high)
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r tidymodels-6}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 12)
```

###

```{r tidymodels-6-test}
#| echo: true
logistic_reg() |>
  set_engine("glm") |>
  fit(Attrition ~ MonthlyIncome + YearsAtCompany + Department, data = attrition) |>
  tidy(conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high)
```

###

`Department` has three levels, so it produces two rows (each compared to a reference department) — and both intervals include zero. Once income and tenure are accounted for, which department someone works in doesn't add anything. We'll drop it.

### Exercise 7

Edit the formula once more: remove `Department`, add `OverTime` instead. Render.

```
logistic_reg() |>
set_engine("glm") |>
fit(Attrition ~ MonthlyIncome + YearsAtCompany + OverTime, data = attrition) |>
tidy(conf.int = TRUE) |>
select(term, estimate, conf.low, conf.high)
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r tidymodels-7}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 12)
```

###

```{r tidymodels-7-test}
#| echo: true
logistic_reg() |>
  set_engine("glm") |>
  fit(Attrition ~ MonthlyIncome + YearsAtCompany + OverTime, data = attrition) |>
  tidy(conf.int = TRUE) |>
  select(term, estimate, conf.low, conf.high)
```

###

`OverTime`'s interval is nowhere near zero, and the estimate is large and positive — working overtime is strongly associated with higher odds of leaving. Unlike `Department`, this one clearly belongs in the final model.

### Exercise 8

Edit your working chunk one last time: remove the `tidy()`/`select()` steps, and assign the result to `mod_logistic`. Print it. Render.

```
mod_logistic <- logistic_reg() |>
set_engine("glm") |>
fit(Attrition ~ MonthlyIncome + YearsAtCompany + OverTime, data = attrition)

mod_logistic
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r tidymodels-8}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r tidymodels-8-test}
#| echo: true
mod_logistic <- logistic_reg() |>
  set_engine("glm") |>
  fit(Attrition ~ MonthlyIncome + YearsAtCompany + OverTime, data = attrition)

mod_logistic
```

###

From here on, you never touch this formula again. `mod_logistic` is the fixed, saved model every later question in this tutorial addresses.

### Exercise 9

Add `#| cache: true` to the top of your `mod_logistic` chunk, and add `*_cache/` to your `.gitignore`. In the R Terminal, run:

```
show_file(".gitignore")
```

CP/CR.

```{r tidymodels-9}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 4)
```

###

```
analysis_files
*_cache/
```

###

Same reasoning as last time: fitting `mod_logistic` runs maximum-likelihood estimation over all 1,470 rows, and caching means that only reruns when the code producing it actually changes — not every time you edit unrelated text. The cache directory itself is regenerated every render, so it's excluded from git rather than committed.

### Exercise 10

This exercise assumes `mod_logistic` is exactly the model from Exercise 8. In a new code chunk, run the following to see its coefficients as odds ratios. Render.

```
mod_logistic |>
tidy(conf.int = TRUE, exponentiate = TRUE) |>
select(term, estimate, conf.low, conf.high) |>
mutate(across(where(is.numeric), \(x) round(x, 3)))
```

As you can see, the odds ratio for `OverTimeYes` is about 3.98. What does that mean?

```{r tidymodels-10}
question_text(NULL,
  answer("Employees who work overtime have odds of leaving that are about 3.98 times the odds for employees who don't, holding income and tenure fixed — an increase of roughly 298%, with a 95% confidence interval of about 196% to 434%.", correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

```{r tidymodels-10-test}
#| echo: true
mod_logistic |>
  tidy(conf.int = TRUE, exponentiate = TRUE) |>
  select(term, estimate, conf.low, conf.high) |>
  mutate(across(where(is.numeric), \(x) round(x, 3)))
```

###

Here's the answer, using the numbers above:

> Employees who work overtime have odds of leaving that are about 3.98 times the odds for employees who don't, holding income and tenure fixed — an increase of roughly 298%, with a 95% confidence interval of about 196% to 434%.

(These numbers come directly from your own `mod_logistic` — they should match exactly, since the same formula fit on the same data always produces the same coefficients. If yours look different, double check the formula in your working chunk from Exercise 8.)

###

`exponentiate = TRUE` converts a log-odds coefficient into an odds ratio, which is far easier to talk about than raw log-odds. An odds ratio above 1 means higher odds; below 1 means lower odds. Notice the shape of this answer: it's a ratio of odds, holding other variables fixed — not a claim that overtime *causes* anyone in particular to quit, only that the two groups' odds differ by about that much.

### Exercise 11

Using the same table from Exercise 10: the odds ratio for `MonthlyIncome` is about 1.000 — barely different from 1, because it reflects a $1 increase. What does an odds ratio very close to 1 for a $1 change in income actually tell you, and why is it not a useful way to describe this effect?

```{r tidymodels-11}
question_text(NULL,
  answer("An odds ratio of 1.000 for a $1 increase doesn't mean income has no effect — it means the effect of a single dollar is tiny. The coefficient describes a per-unit change, and $1 is too small a unit to see. Rescaling to a more meaningful unit, like $1,000, shows the real size of the effect.", correct = TRUE),
  allow_retry = FALSE,
  incorrect = NULL,
  rows = 6)
```

###

Here's the answer:

> An odds ratio of 1.000 for a $1 increase doesn't mean income has no effect — it means the effect of a single dollar is tiny. The coefficient describes a per-unit change, and $1 is too small a unit to see. Rescaling to a more meaningful unit, like $1,000, shows the real size of the effect.

###

At $1,000 increments, the odds ratio for `MonthlyIncome` is about 0.897 — each additional $1,000 in monthly income is associated with roughly a 10% decrease in the odds of leaving (95% CI: about 6% to 14% decrease), holding tenure and overtime fixed. This is the same underlying coefficient as the table above; only the unit of comparison changed. Always check what unit a coefficient is expressed in before deciding whether it looks "small."

### Exercise 12

Words, math, and code are the three languages of data science. Now write the math. Add the equation for `mod_logistic` as a LaTeX math block above the model chunk:

$$\log\left(\frac{P(\text{Attrition} = \text{Yes})}{1 - P(\text{Attrition} = \text{Yes})}\right) = -1.2767 - 0.0001 \times \text{MonthlyIncome} - 0.0454 \times \text{YearsAtCompany} + 1.3803 \times \text{OverTimeYes}$$

Render.

In the R Terminal, run `show_file("analysis.qmd")`. This prints the whole file, since the equation isn't inside a code chunk. From that output, CP/CR just the equation line and the `mod_logistic` chunk right below it — not the whole file.

```{r tidymodels-12}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 20)
```

###

The relevant part of your output should look like this:

<pre><code>$$\log\left(\frac{P(\text{Attrition} = \text{Yes})}{1 - P(\text{Attrition} = \text{Yes})}\right) = -1.2767 - 0.0001 \times \text{MonthlyIncome} - 0.0454 \times \text{YearsAtCompany} + 1.3803 \times \text{OverTimeYes}$$

#| cache: true
mod_logistic <- logistic_reg() |>
  set_engine("glm") |>
  fit(Attrition ~ MonthlyIncome + YearsAtCompany + OverTime, data = attrition)

mod_logistic
</code></pre>

Once rendered, the equation appears as a typeset formula, not raw text with dollar signs and backslashes. If it doesn't, check for a blank line above and below the block in your `.qmd`.

###

This is the same general form shown at the start of this section — $\log(p / (1-p)) = \beta_0 + \beta_1 x_1 + \cdots$ — now with real, fitted numbers in place of the Greek letters. Unlike a linear equation, you can't plug numbers in and read off a dollar amount directly: the left-hand side is a log-odds, which has to be converted back into a probability. That conversion is exactly what the next section does.

### Exercise 13

Commit and push, with a message like "Fit final logistic regression DGM for employee attrition." In the bash Terminal, run:

```
git log -1
```

CP/CR.

```{r tidymodels-13}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```
commit e5f6a7b8c9d0...
Author: Var Kurapati
Date:   Mon Aug 03 2026

    Fit final logistic regression DGM for employee attrition
```

###

Across this whole section, you never touched the formula for `mod_logistic` again after Exercise 8. The next section uses this exact fitted model, never refitting it, only asking it new questions — the same discipline as before, applied to a different kind of DGM.

## Marginal Effects
###

Interpreting a model with **marginaleffects** still comes down to one distinction: **marginal** vs. **conditional**. Every time you see `by`, you are averaging over real employees; every time you see `condition`, you are looking at a hypothetical.

The one thing that changes for a logistic model: `predictions()` no longer hands back a dollar amount. It hands back a **probability**, between 0 and 1.

###

In this section we cover the same three functions as before: `predictions()`, `avg_predictions()`, and `plot_predictions()`. Keep asking yourself, for every function call, which one you're looking at.

### Exercise 1

In a new code chunk, run `predictions()` for one hypothetical employee: $4,000/month, 3 years at the company, no overtime. Add `type = "prob"`. Render.

```
predictions(mod_logistic,
  newdata = data.frame(MonthlyIncome = 4000, YearsAtCompany = 3, OverTime = "No"),
  type = "prob")
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r predictions-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

```{r predictions-1-test}
#| echo: true
predictions(mod_logistic,
  newdata = data.frame(MonthlyIncome = 4000, YearsAtCompany = 3, OverTime = "No"),
  type = "prob")
```

###

Check the `estimate` column: it should be close to 0.136 — a 13.6% predicted probability of leaving. This is a **conditional** prediction; no real employee has to match this exact combination.

### Exercise 2

Now run `class()` on that same result, with `type = "prob"` added. Render.

```
class(predictions(mod_logistic,
newdata = data.frame(MonthlyIncome = 4000, YearsAtCompany = 3, OverTime = "No"),
type = "prob"))
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r predictions-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 4)
```

###

```{r predictions-2-test}
#| echo: true
class(predictions(mod_logistic,
  newdata = data.frame(MonthlyIncome = 4000, YearsAtCompany = 3, OverTime = "No"),
  type = "prob"))
```

###

```
[1] "predictions" "data.frame"
```

###

Same class as before, regardless of what kind of model produced it or how many rows it has — you can pipe it to `as_tibble()`, `filter()`, or `ggplot()` the same way.

### Exercise 3

Update your chunk to compare two employees: the same $4,000/month, 3-year employee, once with `OverTime = "No"` and once with `OverTime = "Yes"`. Add `type = "prob"`, then pipe to `filter(group == "Yes")` to keep just the probability of leaving for each employee. Render.

```
predictions(mod_logistic,
newdata = data.frame(
MonthlyIncome = c(4000, 4000),
YearsAtCompany = c(3, 3),
OverTime = c("No", "Yes")),
type = "prob") |>
filter(group == "Yes")
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r predictions-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r predictions-3-test}
#| echo: true
predictions(mod_logistic,
  newdata = data.frame(
    MonthlyIncome  = c(4000, 4000),
    YearsAtCompany = c(3, 3),
    OverTime       = c("No", "Yes")),
  type = "prob") |>
  filter(group == "Yes")
```

###

`type = "prob"` returns a row for every outcome category — as Exercise 1 showed, that's two rows per employee here. `filter(group == "Yes")` is the pattern this tutorial will use from here on: keep only the row for the outcome you actually care about, collapsing back to one row per unit. Check the `estimate` column above — same income and tenure, only `OverTime` differs, so the entire gap between the two rows is the `OverTimeYes` coefficient, converted from log-odds into something you can read directly.

### Exercise 4

Update your chunk to pipe the result to `as_tibble()`, after the `filter()` step. Render.

```
predictions(mod_logistic,
newdata = data.frame(
MonthlyIncome = c(4000, 4000),
YearsAtCompany = c(3, 3),
OverTime = c("No", "Yes")),
type = "prob") |>
filter(group == "Yes") |>
as_tibble()
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r predictions-4}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r predictions-4-test}
#| echo: true
predictions(mod_logistic,
  newdata = data.frame(
    MonthlyIncome  = c(4000, 4000),
    YearsAtCompany = c(3, 3),
    OverTime       = c("No", "Yes")),
  type = "prob") |>
  filter(group == "Yes") |>
  as_tibble()
```

###

`as_tibble()` converts the **marginaleffects** output into a standard tibble, same as before — the `filter()` step ahead of it is what keeps this readable.

### Exercise 5

Update your chunk to predict the probability for the first actual employee in the dataset. Instead of passing the whole row, select just the three columns the model actually uses — `attrition` has 31 columns, and passing all of them to a classification model's `predictions()` can blow up memory trying to account for every unused category. Add `type = "prob"` and the same `filter(group == "Yes")` pattern. Render.

```
predictions(mod_logistic,
newdata = attrition[1, c("MonthlyIncome", "YearsAtCompany", "OverTime")],
type = "prob") |>
filter(group == "Yes")
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r predictions-5}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```{r predictions-5-test}
#| echo: true
predictions(mod_logistic,
  newdata = attrition[1, c("MonthlyIncome", "YearsAtCompany", "OverTime")],
  type = "prob") |>
  filter(group == "Yes")
```

###

This employee has about a 30.5% predicted probability of leaving. Check that against whether they actually left (`attrition$Attrition[1]`) — the model doesn't predict certainty, only a probability. Unlike the linear model's dollar-valued residual, there's no simple "gap" to subtract here: the outcome is binary, while the prediction is a number between 0 and 1. A well-built model gives higher probabilities to employees who actually left than to those who didn't, on average — but for any one person, being "right" or "wrong" isn't really the right way to think about it.

### Exercise 6

`newdata` is optional. Update your chunk to call `predictions(mod_logistic, type = "prob")` with no `newdata` argument, filter to `group == "Yes"`, pipe to `as_tibble()`, and print `nrow()`. Render.

```
predictions(mod_logistic, type = "prob") |>
  filter(group == "Yes") |>
  as_tibble() |>
  nrow()
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r predictions-6}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

```{r predictions-6-test}
#| echo: true
predictions(mod_logistic, type = "prob") |>
  filter(group == "Yes") |>
  as_tibble() |>
  nrow()
```

###

Without `newdata`, `predictions()` runs on every row of the data the model was fit on. Without the `filter()`, you'd get 2,940 rows — two per employee, one for each outcome category, exactly like every hypothetical unit you've predicted on so far. With the filter, it drops back to 1,470 — one probability-of-leaving per real employee, matching the count from the Introduction.

###

`avg_predictions(mod_logistic, type = "prob")` with no `by` argument averages the model's predicted probabilities across every real employee, for each outcome category — filtering to `group == "Yes"` and comparing that number against the actual attrition rate in the data (`mean(attrition$Attrition == "Yes")`) should show them matching almost exactly, the same guarantee you saw with the linear model in the last tutorial. Try it yourself in a scratch chunk if you want to confirm it before moving on.

### Exercise 7

Update your chunk to use `avg_predictions()` with `by = "OverTime"` and `type = "prob"`, filtered to `group == "Yes"`, to get the average predicted probability of leaving for each overtime group. Render.

```
avg_predictions(mod_logistic, by = "OverTime", type = "prob") |>
filter(group == "Yes")
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r predictions-7}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```{r predictions-7-test}
#| echo: true
avg_predictions(mod_logistic, by = "OverTime", type = "prob") |>
  filter(group == "Yes")
```

###

Compare the two `estimate` values above — employees who work overtime should show a noticeably higher average predicted probability of leaving than those who don't. `by` doesn't care whether the grouping variable is a predictor in the model or not; here it happens to be one, which is perfectly normal. It's still the same operation as always: predict for every real employee, then average within groups.

### Exercise 8

Now build a **conditional** prediction on purpose, using `datagrid()`: three hypothetical employees at $3,000, $6,000, and $9,000 monthly income, with tenure and overtime status held at their typical values automatically. Add `type = "prob"` and filter to `group == "Yes"`. Render.

```
predictions(mod_logistic,
newdata = datagrid(MonthlyIncome = c(3000, 6000, 9000)),
type = "prob") |>
filter(group == "Yes")
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r predictions-8}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r predictions-8-test}
#| echo: true
predictions(mod_logistic,
  newdata = datagrid(MonthlyIncome = c(3000, 6000, 9000)),
  type = "prob") |>
  filter(group == "Yes")
```

###

`datagrid()` fixes every variable you don't mention at a representative value — the mean for numeric variables (`YearsAtCompany`), the most common category for factor variables (`OverTime`) — and varies only the ones you specify. This is the **conditional** counterpart to Exercise 7's **marginal** `avg_predictions()` — no real employee needs to exist at any of these three incomes.

### Exercise 9

Ask your AI assistant to write code using `plot_predictions(mod_logistic, condition = "MonthlyIncome", type = "prob", draw = FALSE)` — with `type = "prob"` added — piped into a `filter(group == "Yes")` step and then `ggplot()`, to visualize how predicted probability of leaving varies with `MonthlyIncome`, with a shaded confidence band, a title, axis labels, and a percent-formatted y-axis. Paste the code it gives you into a new chunk. Render, and adjust the AI's code if it does not match the pattern below.

```
plot_predictions(mod_logistic, condition = list(MonthlyIncome = seq(1000, 20000, by = 500)), type = "prob", draw = FALSE) |>
filter(group == "Yes") |>
ggplot(aes(x = MonthlyIncome, y = estimate)) +
geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) +
geom_line() +
labs(
title = "Predicted attrition probability decreases with income",
x = "Monthly income ($)",
y = "Predicted probability of leaving"
) +
scale_x_continuous(labels = scales::dollar) +
scale_y_continuous(labels = scales::percent)
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r predictions-9}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 10)
```

###

```{r predictions-9-test}
#| echo: true
grid9 <- data.frame(
  MonthlyIncome  = seq(1000, 20000, by = 500),
  YearsAtCompany = mean(attrition$YearsAtCompany),
  OverTime       = "No")

predictions(mod_logistic, newdata = grid9, type = "prob") |>
  filter(group == "Yes") |>
  ggplot(aes(x = MonthlyIncome, y = estimate)) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) +
  geom_line() +
  labs(
    title = "Predicted attrition probability decreases with income",
    x = "Monthly income ($)",
    y = "Predicted probability of leaving"
  ) +
  scale_x_continuous(labels = scales::dollar) +
  scale_y_continuous(labels = scales::percent)
```

###

`draw = FALSE` hands back the raw numbers behind the plot rather than a finished plot to edit — but now with two rows per income value (one per `Group`) rather than one, which is exactly why the `filter()` step comes before `ggplot()` this time, not after. Without it, both the "stayed" and "left" curves would plot on top of each other with no way to tell them apart. The curve bends rather than running in a straight line, unlike the linear model's plot: probabilities are squeezed between 0 and 1, so the same-sized change in log-odds produces a smaller change in probability out at the extremes than it does in the middle.

### Exercise 10

Ask your AI assistant to extend that plot to show predicted probability across both `MonthlyIncome` and `OverTime`, using `condition = c("MonthlyIncome", "OverTime")` with `type = "prob"` and `draw = FALSE`, filtered to `group == "Yes"`, then piped into `ggplot()` — color the lines by `OverTime`. Paste its code into a new chunk. Render, and adjust if needed.

```
grid10 <- tidyr::crossing(MonthlyIncome = seq(1000, 20000, by = 500), OverTime = c("No", "Yes")) |> dplyr::mutate(YearsAtCompany = mean(attrition$YearsAtCompany))

predictions(mod_logistic, newdata = grid10, type = "prob") |>
  filter(group == "Yes") |>
ggplot(aes(x = MonthlyIncome, y = estimate, color = OverTime)) +
geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = OverTime),
alpha = 0.15, color = NA) +
geom_line() +
labs(
title = "Predicted attrition probability by income and overtime",
x = "Monthly income ($)",
y = "Predicted probability of leaving"
) +
scale_x_continuous(labels = scales::dollar) +
scale_y_continuous(labels = scales::percent)
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r predictions-10}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 10)
```

###

```{r predictions-10-test}
#| echo: true
grid10 <- tidyr::crossing(MonthlyIncome = seq(1000, 20000, by = 500), OverTime = c("No", "Yes")) |> dplyr::mutate(YearsAtCompany = mean(attrition$YearsAtCompany))

predictions(mod_logistic, newdata = grid10, type = "prob") |>
  filter(group == "Yes") |>
  ggplot(aes(x = MonthlyIncome, y = estimate, color = OverTime)) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = OverTime),
    alpha = 0.15, color = NA) +
  geom_line() +
  labs(
    title = "Predicted attrition probability by income and overtime",
    x = "Monthly income ($)",
    y = "Predicted probability of leaving"
  ) +
  scale_x_continuous(labels = scales::dollar) +
  scale_y_continuous(labels = scales::percent)
```

###

When two variables are passed to `condition`, `plot_predictions()` varies the first continuously and holds the second at its distinct values — that's what `color = OverTime` splits into two separate curves. Compare where the overtime curve sits relative to the no-overtime curve at every income level in your own plot above.

### Exercise 11

Now plot a **marginal** quantity instead of a conditional one: use `plot_predictions()` with `by = "OverTime"`, `type = "prob"`, and `draw = FALSE`, filtered to `group == "Yes"`, piped into `ggplot()`, to visualize average predicted probability by overtime group. Render.

```
plot_predictions(mod_logistic, by = "OverTime", type = "prob", draw = FALSE) |>
filter(group == "Yes") |>
ggplot(aes(x = OverTime, y = estimate)) +
geom_pointrange(aes(ymin = conf.low, ymax = conf.high)) +
labs(
title = "Average predicted attrition probability by overtime status",
x = NULL,
y = "Average predicted probability of leaving"
) +
scale_y_continuous(labels = scales::percent)
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r predictions-11}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 10)
```

###

```{r predictions-11-test}
#| echo: true
plot_predictions(mod_logistic, by = "OverTime", type = "prob", draw = FALSE) |>
  filter(group == "Yes") |>
  ggplot(aes(x = OverTime, y = estimate)) +
  geom_pointrange(aes(ymin = conf.low, ymax = conf.high)) +
  labs(
    title = "Average predicted attrition probability by overtime status",
    x = NULL,
    y = "Average predicted probability of leaving"
  ) +
  scale_y_continuous(labels = scales::percent)
```

###

Same function, same package — but passing `by` instead of `condition` switches from a hypothetical relationship to an average over real employees, grouped by overtime status. `condition` (the last two exercises) is a conditional, counterfactual view; `by` (this one) is a marginal, descriptive one — the same two numbers you already saw in Exercise 7, now as a picture.

### Exercise 12

Commit and push, with a message like "Add marginaleffects predictions and plots." In the bash Terminal, run:

```
git log -1
```

CP/CR.

```{r predictions-12}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 5)
```

###

```
commit f6a7b8c9d0e1...
Author: Var Kurapati
Date: Mon Aug 03 2026

Add marginaleffects predictions and plots
```

###

You have now used `predictions()`, `avg_predictions()`, and `plot_predictions()` on a logistic regression model, and seen the one real structural difference from linear regression: every prediction comes back as one row per outcome category, filtered down to the one you care about. The same three functions work identically on multinomial and ordinal models — except there, `filter(group == "Yes")` won't be enough, since there are more than two categories to choose from.

## Summary
###

You have built an analysis page that fits a logistic regression model and visualizes its predictions. Now wire it into a navigable Quarto website and publish it.

### Exercise 1

Create a `_quarto.yml` file at the top level of your repo:

```
project:
  type: website

website:
  title: "Employee Attrition"
  navbar:
    left:
      - href: analysis.qmd
        text: Analysis
      - href: about.qmd
        text: About

format:
  html:
    theme: cosmo
```

In the bash Terminal, run:

```
quarto render
ls
```

In the R Terminal, run `show_file("_quarto.yml")`. CP/CR.

```{r summary-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 15)
```

###

<pre><code>project:
  type: website

website:
  title: "Employee Attrition"
  navbar:
    left:
      - href: analysis.qmd
        text: Analysis
      - href: about.qmd
        text: About

format:
  html:
    theme: cosmo
</code></pre>

###

Same pattern as before: `_quarto.yml` turns a folder of `.qmd` files into a navigable website.

### Exercise 2

Create `about.qmd` with a YAML title of `"About"`, your name, and today's date. Clean up `analysis.qmd` so it contains only: a setup chunk, the final `mod_logistic` model chunk, its equation, and the final three visualizations from the Marginal Effects plotting exercises. Remove all intermediate output. Render.

In the R Terminal, run `show_file("about.qmd")`, then `show_file("analysis.qmd")`. CP/CR both.

```{r summary-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 12)
```

###

```
---
title: "About"
---

Var Kurapati, August 2026.
```

###

Your `analysis.qmd` should now have exactly five code chunks — setup, the `mod_logistic` model, and the three `plot_predictions()` calls from the plotting exercises — plus one math block.

### Exercise 3

Preview the site locally. In the bash Terminal, run:

```
quarto preview
```

Confirm both pages load with no errors. Stop the preview and CP/CR the first few lines of terminal output.

```{r summary-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

Preparing to preview
Watching files for changes
GET: /
GET: /analysis.html
GET: /about.html
Browse at http://localhost:3417/

(your port number and exact lines may differ — that's fine)

###

Catching broken links or missing pages before they go live is much cheaper than catching them after publishing.

### Exercise 4

Publish the website to GitHub Pages. In the bash Terminal, run:

```
quarto publish gh-pages
```

Copy/paste the resulting URL below.

```{r summary-4}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

Both pages are now live, linked by the navigation bar.

### Exercise 5

Commit and push any remaining changes. Copy/paste the URL to your GitHub repo.

```{r summary-5}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

The same pattern you used here — `logistic_reg() |> set_engine() |> fit()` followed by `predictions()`, `avg_predictions()`, and `plot_predictions()`, always keeping marginal and conditional straight — applies to multinomial and ordinal regression too. The interface never changes; only the structure, and what the predictions mean, does.

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
