<!--VK: Realized I went by-model sections (Linear/Logistic/Multi/Ordinal, wiht each having predictions() and plot_predictions() rather than by-function. -->

---
title: Marginal Effects
author: Var Kurapati
tutorial:
  id: marginaleffects
  version: 1.0.2
output:
  learnr::tutorial:
    df_print: default
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: 'Learn to interpret statistical models using the marginaleffects package. Covers predictions() and plot_predictions() across four model types.'
---

```{r setup, include=FALSE}
library(learnr)
library(tutorial.helpers)
library(knitr)
library(tidyverse)
library(tidymodels)
library(marginaleffects)
library(ggdist)
library(nnet)
library(ordinal)
library(ordered)

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(out.width = '90%')
options(tutorial.exercise.timelimit = 600,
        tutorial.storage = "local")

titanic <- read_csv("data/titanic.csv", show_col_types = FALSE) |>
  mutate(Survived = as.factor(Survived),
         PClass = as.factor(PClass),
         PClass_ord = factor(PClass, levels = c("3rd", "2nd", "1st"), ordered = TRUE))

mod_linear <- linear_reg() |> set_engine("lm") |>
  fit(Age ~ PClass + SexCode, data = titanic)

mod_logistic <- logistic_reg() |> set_engine("glm") |>
  fit(Survived ~ Age * SexCode * PClass, data = titanic)
 
# DK: I don't think we need the below.
# predictions(mod, type = "prob")

mod_multinomial <- multinom_reg() |> set_engine("nnet") |>
  fit(PClass ~ Age + SexCode, data = titanic)

mod_ordered <- ordinal_reg() |> set_engine("polr") |>
  fit(PClass_ord ~ Age + SexCode, data = titanic)
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

## Introduction
###

On April 15, 1912, the RMS Titanic sank after striking an iceberg in the North Atlantic. Of the 2,224 passengers and crew aboard, only 710 survived. But survival was not random — your age, sex, and ticket class had a significant effect on whether you made it onto a lifeboat.

This tutorial uses the [**marginaleffects**](https://marginaleffects.com/) package to ask: **what does our model actually predict, and how do those predictions change depending on how we ask the question?** We cover two functions: `predictions()` and `plot_predictions()`.

`by` takes every real passenger in the dataset, runs the model on each one, then averages within groups. `condition` builds hypothetical passengers by holding all other variables fixed and only varying the one you specify — showing the model's behavior "all else equal."

We assume you have already completed the **tidymodels** tutorial. The four fitted models from that tutorial — `mod_linear`, `mod_logistic`, `mod_multinomial`, and `mod_ordered` — are used throughout this tutorial.

### Exercise 1

You should be connected to a repo named `marginaleffects`. If you are not, create one and connect to it.

Create a new file, `analysis.qmd`, with the title `"Titanic Survival"` and your name as the author. Add `execute: echo: false` to the YAML header. In a bash Terminal, render it:

```
quarto render analysis.qmd
```

Open `analysis.html` with Live Server. Create a `.gitignore` with `*_files` on the first line. Commit and push.

In the R Terminal, run:

```
show_file(".gitignore")
```

CP/CR.

```{r introduction-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

```
*_files
```

###

The **marginaleffects** package allows you to answer questions using any fitted model in a uniform way — the same functions work across over 100 different model types.

### Exercise 2

Create a `data` directory at the top level of the repo. Download `titanic.csv` into it from:

```
https://github.com/PPBDS/misc.tutorials/raw/refs/heads/main/inst/tutorials/marginaleffects/data/titanic.csv
```

In the bash Terminal, run `ls data`. CP/CR.

```{r introduction-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

```
titanic.csv
```

###

The Titanic dataset has five key variables: `Survived` (1 = survived, 0 = died), `Age` (in years), `SexCode` (1 = female, 0 = male), `PClass` (ticket class: 1st, 2nd, 3rd), and `PClass_ord` (same as PClass but ordered: 3rd < 2nd < 1st).

### Exercise 3

In a new code chunk in `analysis.qmd`, load the libraries and read `data/titanic.csv`. Add `#| message: false`. Convert `Survived` and `PClass` to factors and create `PClass_ord` as an ordered factor. Add `#| cache: true`. Render.

```
#| message: false
#| cache: true
library(tidyverse)
library(tidymodels)
library(marginaleffects)
library(ggdist)
library(nnet)
library(ordinal)
library(ordered)

titanic <- read_csv("data/titanic.csv", show_col_types = FALSE) |>
  mutate(Survived = as.factor(Survived),
         PClass = as.factor(PClass),
         PClass_ord = factor(PClass, levels = c("3rd", "2nd", "1st"),
                             ordered = TRUE))
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r introduction-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 12)
```

###

<pre><code>#| message: false
#| cache: true
library(tidyverse)
library(tidymodels)
library(marginaleffects)
library(ggdist)
library(nnet)
library(ordinal)
library(ordered)

titanic <- read_csv("data/titanic.csv", show_col_types = FALSE) |>
  mutate(Survived = as.factor(Survived),
         PClass = as.factor(PClass),
         PClass_ord = factor(PClass, levels = c("3rd", "2nd", "1st"),
                             ordered = TRUE))
</code></pre>

###

Each row is one passenger. We use `Age`, `SexCode`, and `PClass` to predict different outcomes across four model types.

### Exercise 4

Add a new code chunk with the four fitted models. Copy and paste this exactly:

```
#| cache: true
mod_linear <- linear_reg() |> set_engine("lm") |>
  fit(Age ~ PClass + SexCode, data = titanic)

mod_logistic <- logistic_reg() |> set_engine("glm") |>
  fit(Survived ~ Age * SexCode * PClass, data = titanic)

mod_multinomial <- multinom_reg() |> set_engine("nnet") |>
  fit(PClass ~ Age + SexCode, data = titanic)

mod_ordered <- ordinal_reg() |> set_engine("polr") |>
  fit(PClass_ord ~ Age + SexCode, data = titanic)
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r introduction-4}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 15)
```

###

<pre><code>#| cache: true
mod_linear <- linear_reg() |> set_engine("lm") |>
  fit(Age ~ PClass + SexCode, data = titanic)

mod_logistic <- logistic_reg() |> set_engine("glm") |>
  fit(Survived ~ Age * SexCode * PClass, data = titanic)

mod_multinomial <- multinom_reg() |> set_engine("nnet") |>
  fit(PClass ~ Age + SexCode, data = titanic)

mod_ordered <- ordinal_reg() |> set_engine("polr") |>
  fit(PClass_ord ~ Age + SexCode, data = titanic)
</code></pre>

###

`mod_linear` predicts `Age` — a numeric outcome. `mod_logistic` predicts `Survived` — a binary outcome. `mod_multinomial` predicts `PClass` — a categorical outcome with 3 unordered levels. `mod_ordered` predicts `PClass_ord` — a categorical outcome with 3 ordered levels (3rd < 2nd < 1st).

### Exercise 5

Commit and push. In the bash Terminal, run:

```
git log -1
```

CP/CR.

```{r introduction-5}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

```
commit 3f8a2c1
Author: Var Kurapati <vkurapat@example.com>
Date:   Mon Jul 14 10:23:41 2025 -0700

    Add Titanic data and four models
```

###

You now have four fitted models — one for each type of regression. The next four sections apply `predictions()` and `plot_predictions()` to each one.

## Linear Regression
###

In this section we use two functions: `predictions()` and `plot_predictions()`. We apply them to `mod_linear`, which predicts `Age` from `PClass` and `SexCode`. The outcome is numeric, so `predictions()` returns one row per input with a single estimated value.

### Exercise 1

In a new code chunk, run `class()` on the result of `predictions()` for one hypothetical passenger: a female in 1st class. Render.

```
class(predictions(mod_linear,
  newdata = data.frame(PClass = "1st", SexCode = 1)))
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r linear-0}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 4)
```

###

```{r linear-0-test}
#| echo: true
class(predictions(mod_linear,
  newdata = data.frame(PClass = "1st", SexCode = 1)))
```

###

```
[1] "predictions" "data.frame"
```

###

`predictions()` returns an object of class `predictions` and `data.frame` — you can pipe it to `as_tibble()`, `filter()`, or `ggplot()`.

### Exercise 2

In a new code chunk, run `predictions()` on `mod_linear` for one hypothetical passenger: a female in 1st class. Render.

```
predictions(mod_linear,
  newdata = data.frame(PClass = "1st", SexCode = 1))
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r linear-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

```{r linear-1-test}
#| echo: true
predictions(mod_linear,
  newdata = data.frame(PClass = "1st", SexCode = 1))
```

###

```
 Estimate Std. Error  z Pr(>|z|)   S 2.5 % 97.5 % SexCode PClass
     41.9      0.959 44   <0.001 Inf  40.0   43.8       1    1st
Type: response
```

###

The `estimate` column gives the predicted age — 41.9 years for a female 1st class passenger. Linear regression predicts a single numeric value, so there is always one row per input.

### Exercise 3

Update your chunk to run `predictions()` on two passengers: a female in 1st class and a male in 3rd class. Render.

```
predictions(mod_linear,
  newdata = data.frame(PClass = c("1st", "3rd"),
                       SexCode = c(1, 0)))
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r linear-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r linear-2-test}
#| echo: true
predictions(mod_linear,
  newdata = data.frame(PClass = c("1st", "3rd"),
                       SexCode = c(1, 0)))
```

###

```
 Estimate Std. Error  z Pr(>|z|)   S 2.5 % 97.5 % SexCode PClass
     41.9      0.959 44   <0.001 Inf  40.0   43.8       1    1st
     35.5      0.833 43   <0.001 Inf  33.9   37.2       0    3rd
Type: response
```

###

First class female passengers were predicted to be older on average than male 3rd class passengers. This reflects selection bias in the data — wealthier, older passengers were more likely to be in 1st class.

### Exercise 4

Update your chunk to pipe `predictions()` to `as_tibble()`. Render.

```
predictions(mod_linear,
  newdata = data.frame(PClass = c("1st", "3rd"),
                       SexCode = c(1, 0))) |>
  as_tibble()
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r linear-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r linear-3-test}
#| echo: true
predictions(mod_linear,
  newdata = data.frame(PClass = c("1st", "3rd"),
                       SexCode = c(1, 0))) |>
  as_tibble()
```

###

`as_tibble()` converts the **marginaleffects** output into a standard tibble, which makes it easier to work with using **tidyverse** functions like `select()`, `filter()`, and `ggplot()`.

### Exercise 5

Update your chunk to run `avg_predictions()` on `mod_linear` with `by = "PClass"`. Render.

```
avg_predictions(mod_linear, by = "PClass")
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r linear-4}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

```{r linear-4-test}
#| echo: true
avg_predictions(mod_linear, by = "PClass")
```

###

```
 PClass Estimate Std. Error    z Pr(>|z|)     S 2.5 % 97.5 %
    1st     40.9     0.6960 58.8   <0.001 Inf    39.6   42.3
    2nd     30.0     0.7490 40.1   <0.001 Inf    28.5   31.5
    3rd     25.3     0.5080 49.9   <0.001 Inf    24.3   26.3
Type: response
```

###

This is a **marginal** prediction — the model's average predicted age for each ticket class, weighted by the actual passengers in each class.

### Exercise 6

Update your chunk to visualize how predicted age varies with `PClass` and `SexCode` using `plot_predictions()`. Render.

```
plot_predictions(mod_linear,
  condition = c("PClass", "SexCode"))
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r linear-5}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

```{r linear-5-test}
#| echo: true
plot_predictions(mod_linear,
  condition = c("PClass", "SexCode"))
```

###

`condition` fixes all other predictors at their means/modes and shows how the prediction changes across the specified variables. First class passengers are predicted to be older than 3rd class, and females slightly older than males within each class.

### Exercise 7

Commit and push. In the bash Terminal, run:

```
git log -1
```

CP/CR.

```{r linear-6}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

```
commit 5b3c1d2
Author: Var Kurapati <vkurapat@example.com>
Date:   Mon Jul 14 11:00:00 2025 -0700

    Add linear regression predictions
```

###

Linear regression produces one prediction per input row. Next, logistic regression produces a probability — a number between 0 and 1.

## Logistic Regression
###

In this section we use two functions: `predictions()` and `plot_predictions()`. We apply them to `mod_logistic`, which predicts `Survived` — a binary outcome (0 or 1). For logistic models, `predictions()` requires `type = "prob"` to return probabilities, or `type = "class"` to return the best guess.

### Exercise 1

In a new code chunk, run `class()` on the result of `predictions()` for one hypothetical passenger: a 30-year-old male in 1st class. Use `type = "prob"`. Render.

```
class(predictions(mod_logistic,
  newdata = data.frame(Age = 30, SexCode = 0, PClass = "1st"),
  type = "prob"))
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r logistic-0}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 4)
```

###

```{r logistic-0-test}
#| echo: true
class(predictions(mod_logistic,
  newdata = data.frame(Age = 30, SexCode = 0, PClass = "1st"),
  type = "prob"))
```

###

```
[1] "predictions" "data.frame"
```

###

`predictions()` returns an object of class `predictions` and `data.frame` — you can pipe it to `as_tibble()`, `filter()`, or `ggplot()`.

### Exercise 2

In a new code chunk, run `predictions()` on `mod_logistic` for one hypothetical passenger: a 30-year-old male in 1st class. Use `type = "prob"`. Render.

```
predictions(mod_logistic,
  newdata = data.frame(Age = 30, SexCode = 0, PClass = "1st"),
  type = "prob")
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r logistic-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

```{r logistic-1-test}
#| echo: true
predictions(mod_logistic,
  newdata = data.frame(Age = 30, SexCode = 0, PClass = "1st"),
  type = "prob")
```

###

```
 Group Estimate Std. Error    z Pr(>|z|)    S 2.5 % 97.5 % Age SexCode PClass
     0    0.531     0.0591 8.98   <0.001 61.7 0.415  0.646  30       0    1st
     1    0.469     0.0591 7.95   <0.001 48.9 0.354  0.585  30       0    1st
Type: prob
```

###

Unlike linear regression, logistic regression returns **two rows** — one for each possible outcome (died = 0, survived = 1). The `group` column identifies which outcome each row refers to.

### Exercise 3

Update your chunk to predict survival for the first actual passenger in the dataset. Use `type = "prob"`. Render.

```
predictions(mod_logistic, newdata = titanic[1, ], type = "prob")
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r logistic-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

```{r logistic-2-test}
#| echo: true
predictions(mod_logistic, newdata = titanic[1, ], type = "prob")
```

###

```
 Group Estimate Std. Error    z Pr(>|z|)    S 2.5 % 97.5 %
     0   0.0539     0.0111 4.84   <0.001 19.5 0.036 0.0798
     1   0.9461     0.0111 85.0   <0.001  Inf 0.920 0.9638
Type: prob
```

###

We already know this passenger survived — it is in the data. The model predicts a 94.6% survival probability. We predict this passenger not because we don't know the answer, but because we want to ask: what would the model predict if something about them changed?

### Exercise 4

Update your chunk to use `type = "class"` instead of `type = "prob"`. Render.

```
predictions(mod_logistic, newdata = titanic[1, ], type = "class")
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r logistic-3b}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 4)
```

###

```{r logistic-3b-test}
#| echo: true
predictions(mod_logistic, newdata = titanic[1, ], type = "class")
```

###

```
 Estimate
        1
Type: class
```

###

`type = "class"` gives the model's best guess — survived (1) or died (0). `type = "prob"` gives two rows with probabilities; `type = "class"` gives one row with the predicted category.

### Exercise 5

Update your chunk to predict survival for the same first passenger, but change their ticket class to 3rd. Use `type = "prob"`. Render.

```
predictions(mod_logistic,
  newdata = titanic[1, ] |> mutate(PClass = "3rd"),
  type = "prob")
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r logistic-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

```{r logistic-3-test}
#| echo: true
predictions(mod_logistic,
  newdata = titanic[1, ] |> mutate(PClass = "3rd"),
  type = "prob")
```

###

```
 Group Estimate Pr(>|z|)   S 2.5 % 97.5 %
     0     0.55     0.37 1.4 0.441  0.655
     1     0.45     0.37 1.4 0.345  0.559
Type: prob
```

###

The same passenger, but with a 3rd class ticket, has only a 45% predicted survival probability — down from 95%. This is the counterfactual question: same person, different ticket.

### Exercise 6

Update your chunk to use `avg_predictions()` with `by = "PClass"` and `type = "prob"`. Render.

```
avg_predictions(mod_logistic, by = "PClass", type = "prob")
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r logistic-4}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

```{r logistic-4-test}
#| echo: true
avg_predictions(mod_logistic, by = "PClass", type = "prob")
```

###

```
 PClass Estimate Std. Error    z Pr(>|z|)     S 2.5 % 97.5 %
    1st    0.615     0.0241 25.6   <0.001 476.1 0.568  0.662
    2nd    0.453     0.0214 21.2   <0.001 328.1 0.411  0.495
    3rd    0.245     0.0226 10.9   <0.001  88.8 0.201  0.290
Type: prob
```

###

First class passengers had a 62% average predicted survival probability, compared to 25% for 3rd class. This reflects the historical record — lifeboats were loaded from the upper decks first.

### Exercise 7

Update your chunk to visualize predicted survival probability across ages and ticket classes using `plot_predictions()`. Add `geom_dots()` from **ggdist** to show actual outcomes. Use `type = "prob"`. Render.

```
plot_predictions(mod_logistic,
  condition = c("Age", "PClass"),
  type = "prob") +
  geom_dots(
    alpha = .8, scale = .3, pch = 18,
    data = titanic,
    aes(x = Age,
        y = as.numeric(as.character(Survived)),
        side = ifelse(as.numeric(as.character(Survived)) == 1,
                      "bottom", "top"))
  ) +
  labs(
    title = "Predicted survival probability on the Titanic",
    x = "Age", y = "Predicted probability of survival",
    color = "Ticket class", fill = "Ticket class"
  ) +
  scale_y_continuous(labels = scales::percent)
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r logistic-5}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 15)
```

###

```{r logistic-5-test}
#| echo: true
plot_predictions(mod_logistic,
  condition = c("Age", "PClass"),
  type = "prob") +
  geom_dots(
    alpha = .8, scale = .3, pch = 18,
    data = titanic,
    aes(x = Age,
        y = as.numeric(as.character(Survived)),
        side = ifelse(as.numeric(as.character(Survived)) == 1,
                      "bottom", "top"))
  ) +
  labs(
    title = "Predicted survival probability on the Titanic",
    x = "Age", y = "Predicted probability of survival",
    color = "Ticket class", fill = "Ticket class"
  ) +
  scale_y_continuous(labels = scales::percent)
```

###

Every point on these curves is a **conditional** prediction — the model's predicted survival probability at a given age, holding other variables at their means. The dots show actual outcomes: survivors below the curve, deaths above.

###

Notice that 1st and 2nd class have similar survival odds for children, but diverge sharply for adults. By middle age, 2nd class passengers had much lower survival odds than 1st class.

### Exercise 8

Commit and push. In the bash Terminal, run:

```
git log -1
```

CP/CR.

```{r logistic-6}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

```
commit 6c4d2e3
Author: Var Kurapati <vkurapat@example.com>
Date:   Mon Jul 14 12:00:00 2025 -0700

    Add logistic regression predictions
```

###

Logistic regression returns two rows per input — one probability for each outcome. Next, multinomial regression returns three rows — one per outcome category.

## Multinomial Regression
###

In this section we use two functions: `predictions()` and `plot_predictions()`. We apply them to `mod_multinomial`, which predicts `PClass` — a categorical outcome with three unordered levels. Because there are three possible outcomes, `predictions()` returns three rows per input — one probability for each class.

### Exercise 1

In a new code chunk, run `class()` on the result of `predictions()` for one hypothetical passenger: a 30-year-old male. Use `type = "prob"`. Render.

```
class(predictions(mod_multinomial,
  newdata = data.frame(Age = 30, SexCode = 0),
  type = "prob"))
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r multinomial-0}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 4)
```

###

```{r multinomial-0-test}
#| echo: true
class(predictions(mod_multinomial,
  newdata = data.frame(Age = 30, SexCode = 0),
  type = "prob"))
```

###

```
[1] "predictions" "data.frame"
```

###

`predictions()` always returns the same class regardless of model type — this is the key benefit of **marginaleffects**: a uniform interface across all model types.

### Exercise 2

In a new code chunk, run `predictions()` on `mod_multinomial` for one hypothetical passenger: a 30-year-old male. Use `type = "prob"`. Render.

```
predictions(mod_multinomial,
  newdata = data.frame(Age = 30, SexCode = 0),
  type = "prob")
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r multinomial-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

```{r multinomial-1-test}
#| echo: true
predictions(mod_multinomial,
  newdata = data.frame(Age = 30, SexCode = 0),
  type = "prob")
```

###

```
 Group Estimate Std. Error    z Pr(>|z|)     S 2.5 % 97.5 % Age SexCode
   1st    0.216     0.0207 10.4   <0.001  82.2 0.176  0.257  30       0
   2nd    0.299     0.0223 13.4   <0.001 133.7 0.255  0.343  30       0
   3rd    0.485     0.0248 19.6   <0.001 281.2 0.436  0.533  30       0
Type: prob
```

###

Unlike linear and logistic models, multinomial regression returns **three rows per input** — one probability for each ticket class. The probabilities sum to 1. A 30-year-old male was most likely to be in 3rd class (49%), then 2nd (30%), then 1st (22%).

### Exercise 3

Update your chunk to use `type = "class"` to get the model's best guess for a 30-year-old male. Render.

```
predictions(mod_multinomial,
  newdata = data.frame(Age = 30, SexCode = 0),
  type = "class")
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r multinomial-2b}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 4)
```

###

```{r multinomial-2b-test}
#| echo: true
predictions(mod_multinomial,
  newdata = data.frame(Age = 30, SexCode = 0),
  type = "class")
```

###

```
 Estimate Age SexCode
      3rd  30       0
Type: class
```

###

`type = "class"` gives the most likely ticket class — 3rd for a 30-year-old male. Compare this to `type = "prob"` which gives three rows of probabilities.

### Exercise 4

Update your chunk to pipe `predictions()` to `as_tibble()`. Use `type = "prob"`. Render.

```
predictions(mod_multinomial,
  newdata = data.frame(Age = 30, SexCode = 0),
  type = "prob") |>
  as_tibble()
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r multinomial-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r multinomial-2-test}
#| echo: true
predictions(mod_multinomial,
  newdata = data.frame(Age = 30, SexCode = 0),
  type = "prob") |>
  as_tibble()
```

###

The `group` column identifies which outcome category each row corresponds to. Converting to a tibble makes it easy to `filter()` to just the class you care about or `ggplot()` the probabilities.

### Exercise 5

Update your chunk to use `avg_predictions()` with `by = "SexCode"` and `type = "prob"`. Render.

```
avg_predictions(mod_multinomial, by = "SexCode", type = "prob")
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r multinomial-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

```{r multinomial-3-test}
#| echo: true
avg_predictions(mod_multinomial, by = "SexCode", type = "prob")
```

###

`avg_predictions()` still returns multiple rows per group for multinomial models — one row per outcome category per group value. The `group` column tells you which class each row refers to.

### Exercise 6

Update your chunk to visualize predicted class probabilities across age using `plot_predictions()`. Use `type = "prob"`. Render.

```
plot_predictions(mod_multinomial,
  condition = c("Age", "group"),
  type = "prob")
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r multinomial-4}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

```{r multinomial-4-test}
#| echo: true
plot_predictions(mod_multinomial,
  condition = c("Age", "group"),
  type = "prob")
```

###

For multinomial models, `group` is a special variable that identifies the outcome category. Passing it to `condition` produces one line per ticket class, showing how the predicted probability of being in each class changes with age.

### Exercise 7

Commit and push. In the bash Terminal, run:

```
git log -1
```

CP/CR.

```{r multinomial-5}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

```
commit 7d5e3f4
Author: Var Kurapati <vkurapat@example.com>
Date:   Mon Jul 14 13:00:00 2025 -0700

    Add multinomial regression predictions
```

###

Multinomial regression returns one row per outcome category. Ordered regression is similar, but the categories have a natural ordering.

## Ordinal Regression
###

In this section we use two functions: `predictions()` and `plot_predictions()`. We apply them to `mod_ordered`, which predicts `PClass_ord` — an ordered version of ticket class where 3rd < 2nd < 1st. Like the multinomial model, `predictions()` returns three rows per input.

### Exercise 1

In a new code chunk, run `class()` on the result of `predictions()` for one hypothetical passenger: a 30-year-old male. Use `type = "prob"`. Render.

```
class(predictions(mod_ordered,
  newdata = data.frame(Age = 30, SexCode = 0),
  type = "prob"))
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r ordinal-0}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 4)
```

###

```{r ordinal-0-test}
#| echo: true
class(predictions(mod_ordered,
  newdata = data.frame(Age = 30, SexCode = 0),
  type = "prob"))
```

###

```
[1] "predictions" "data.frame"
```

###

`predictions()` returns the same class for all four model types — linear, logistic, multinomial, and ordinal. The interface is uniform even though the underlying models are very different.

### Exercise 2

In a new code chunk, run `predictions()` on `mod_ordered` for one hypothetical passenger: a 30-year-old male. Use `type = "prob"`. Render.

```
predictions(mod_ordered,
  newdata = data.frame(Age = 30, SexCode = 0),
  type = "prob")
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r ordinal-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

```{r ordinal-1-test}
#| echo: true
predictions(mod_ordered,
  newdata = data.frame(Age = 30, SexCode = 0),
  type = "prob")
```

###

```
 Group Estimate Age SexCode
 value    0.456  30       0
 value    0.315  30       0
 value    0.230  30       0
Type: prob
```

###

Like the multinomial model, the ordinal model returns three rows — one per ticket class. Notice the probabilities are slightly different from the multinomial model because the ordinal model uses the ordering constraint (3rd < 2nd < 1st) to estimate the probabilities.

### Exercise 3

Update your chunk to use `type = "class"` to get the model's best guess for a 30-year-old male. Render.

```
predictions(mod_ordered,
  newdata = data.frame(Age = 30, SexCode = 0),
  type = "class")
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r ordinal-1b}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 4)
```

###

```{r ordinal-1b-test}
#| echo: true
predictions(mod_ordered,
  newdata = data.frame(Age = 30, SexCode = 0),
  type = "class")
```

###

```
 Estimate Age SexCode
      3rd  30       0
Type: class
```

###

`type = "class"` gives the most likely ticket class. For the ordinal model, the ordering constraint (3rd < 2nd < 1st) affects the probabilities but `type = "class"` still returns the single most likely category.

### Exercise 4

Update your chunk to run `predictions()` for a 30-year-old female and compare to the male predictions. Use `type = "prob"`. Render.

```
predictions(mod_ordered,
  newdata = data.frame(Age = c(30, 30),
                       SexCode = c(0, 1)),
  type = "prob")
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r ordinal-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r ordinal-2-test}
#| echo: true
predictions(mod_ordered,
  newdata = data.frame(Age = c(30, 30),
                       SexCode = c(0, 1)),
  type = "prob")
```

###

The model predicts that females were more likely to be in higher ticket classes than males of the same age — reflecting the historical tendency for women to travel with wealthier companions.

### Exercise 5

Update your chunk to use `plot_predictions()` with `condition = c("Age", "group")` and `type = "prob"`. Render.

```
plot_predictions(mod_ordered,
  condition = c("Age", "group"),
  type = "prob")
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r ordinal-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

```{r ordinal-3-test}
#| echo: true
plot_predictions(mod_ordered,
  condition = c("Age", "group"),
  type = "prob")
```

###

The ordinal model produces smooth, monotonically ordered probability curves. Older passengers were more likely to be in 1st class and less likely to be in 3rd — consistent with the historical pattern of wealthier, older travelers buying premium tickets.

### Exercise 6

Commit and push. In the bash Terminal, run:

```
git log -1
```

CP/CR.

```{r ordinal-4}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

```
commit 8e6f4g5
Author: Var Kurapati <vkurapat@example.com>
Date:   Mon Jul 14 14:00:00 2025 -0700

    Add ordinal regression predictions
```

###

You have now used `predictions()` and `plot_predictions()` with all four model types. The final section wires everything into a published website.

## Summary
###

You have built four sections of analysis. Now wire them together into a navigable Quarto website and publish the result.

### Exercise 1

Create a `_quarto.yml` file at the top level of your repo with the following contents:

```
project:
  type: website

website:
  title: "Titanic Analysis"
  navbar:
    left:
      - href: analysis.qmd
        text: Analysis
      - href: about.qmd
        text: About

format:
  html:
    theme: cosmo
```

In the bash Terminal, run:

```
quarto render
ls
```

In the R Terminal, run `show_file("_quarto.yml")`. CP/CR.

```{r summary-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 15)
```

###

<pre><code>project:
  type: website

website:
  title: "Titanic Analysis"
  navbar:
    left:
      - href: analysis.qmd
        text: Analysis
      - href: about.qmd
        text: About

format:
  html:
    theme: cosmo
</code></pre>

###

`_quarto.yml` turns a folder of `.qmd` files into a navigable website. Without it, each page renders independently.

### Exercise 2

Create `about.qmd` with a YAML title of `"About"`, your name, and today's date. Also clean up `analysis.qmd` so it contains only: a setup chunk, a data-loading chunk, a model chunk, and the four final plots. Remove all intermediate tables. Render.

In the R Terminal, run `show_file("about.qmd")`. CP/CR.

```{r summary-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 10)
```

###

```
---
title: "About"
---

Var Kurapati, July 2025.
```

###

An About page tells visitors who made the site and when.

### Exercise 3

Publish the website to GitHub Pages. In the bash Terminal, run:

```
quarto publish gh-pages
```

Copy/paste the resulting URL below.

```{r summary-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

`quarto publish gh-pages` without a filename publishes the whole website. Both pages are now live, linked by the navigation bar.

### Exercise 4

Commit and push any remaining changes. Copy/paste the URL to your GitHub repo.

```{r summary-4}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

The **marginaleffects** package works with over 100 model types. The same two functions — `predictions()` and `plot_predictions()` — apply to all of them in exactly the same way.

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
