---
title: Marginal Effects
author: Var Kurapati
tutorial:
  id: marginaleffects
  version: 1.0.1
output:
  learnr::tutorial:
    df_print: default
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: 'Learn to interpret statistical models using the marginaleffects package, with a focus on the difference between conditional and marginal quantities.'
---

```{r setup, include=FALSE}
library(learnr)
library(tutorial.helpers)
library(knitr)
library(tidyverse)
library(marginaleffects)
library(ggdist)

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(out.width = '90%')
options(tutorial.exercise.timelimit = 600,
        tutorial.storage = "local")

titanic <- read_csv("data/titanic.csv", show_col_types = FALSE)
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

## Introduction
###

On April 15, 1912, the RMS Titanic sank after striking an iceberg in the North Atlantic. Of the 2,224 passengers and crew aboard, only 710 survived. But survival was not random — your age, sex, and ticket class had a significant effect on whether you made it onto a lifeboat.

This tutorial uses logistic regression and the `marginaleffects` package to ask: **what does our model actually predict about who survived, and how do those predictions change depending on how we ask the question?**

We recommend using an agentic coding tool such as [Gemini CLI](https://github.com/google-gemini/gemini-cli) or [Claude Code](https://claude.ai/code), which can read and edit your files directly. Our instructions are written with these tools in mind.

### Exercise 1

You should be connected to a repo named `marginaleffects`. If you are not, create one and connect to it.

Create a new file, `analysis.qmd`, with the title `"Titanic Survival"` and your name as the author. In a bash Terminal, render it:

```
quarto render analysis.qmd
```

Open `analysis.html` with Live Server (right-click it in the Explorer → **Open with Live Server**) and keep the tab open. It refreshes on every render. Going forward, we will just tell you to "Render" when we want you to take these steps.

Create a `.gitignore` with `*_files` followed by a blank line. Commit and push.

In the R Terminal, run:

```
show_file(".gitignore")
```

If that fails, it is probably because you have not yet loaded `library(tutorial.helpers)` in the R Terminal.

CP/CR.

```{r introduction-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

```
*_files
```

###

The rendered HTML creates a support directory called `*_files/` that holds fonts and scripts. It does not belong in your repo because it is generated output.

###

The marginaleffects package transforms those raw numbers into quantities that are easy to understand.

### Exercise 2

Create a `data` directory at the top level of the `marginaleffects` repo. Download `titanic.csv` into it from this URL:

```
https://github.com/PPBDS/misc.tutorials/raw/refs/heads/main/inst/tutorials/marginaleffects/data/titanic.csv
```

In the bash Terminal, run `ls data`. CP/CR.

```{r introduction-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

```
titanic.csv
```

###

The Titanic dataset has four key variables: `Survived` (1 = survived, 0 = died), `Age` (in years), `SexCode` (1 = female, 0 = male), and `PClass` (ticket class: 1st, 2nd, 3rd).

### Exercise 3

In your `analysis.qmd`, add a new code chunk that loads the libraries and reads `data/titanic.csv` into a tibble named `titanic`. Add `#| message: false` to suppress messages. Add `execute: echo: false` to the YAML header to hide code globally. Print the first 10 rows. Render.

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r introduction-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

<pre><code>#| message: false
library(tidyverse)
library(marginaleffects)
library(ggdist)

titanic <- read_csv("data/titanic.csv", show_col_types = FALSE)
head(titanic, 10)
</code></pre>

###

Each row is one passenger. We will use `Age`, `SexCode`, and `PClass` to predict whether they survived.

### Exercise 4

Commit and push. Ask AI for a commit message. In the bash Terminal, run:

```
git log -1
```

CP/CR.

```{r introduction-4}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

```
commit 3f8a2c1
Author: Var Kurapati <vkurapat@example.com>
Date:   Mon Jul 14 10:23:41 2025 -0700

    Add Titanic data and initial setup
```

###

Committing before any analysis creates a clean baseline you can always return to.

## Predictions
###

A **prediction** is the outcome expected by a fitted model for a given combination of predictor values. Instead of reporting a log-odds coefficient, we report a predicted probability — something any reader can understand.

### Exercise 1

In a new code chunk, fit a logistic regression model predicting `Survived` from `Age`, `SexCode`, and `PClass` with all interactions. Assign it to `mod` and print a summary. Render.

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r predictions-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

```{r predictions-1-test}
#| echo: true
mod <- glm(Survived ~ Age * SexCode * PClass, data = titanic, family = binomial)
summary(mod)
```

###

The fitted model is not the answer to anything — it is the machine we built so we can answer questions. `predictions()` runs that machine and returns the expected outcome for each unit you specify.

###

The raw coefficients are log-odds ratios — essentially impossible to interpret intuitively. `marginaleffects` transforms them into predicted probabilities on the same scale as the outcome.

### Exercise 2

Update your working chunk to use `predictions()` to compute fitted values for every passenger. Print the result. Render.

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r predictions-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

```{r predictions-2-test}
#| echo: true
predictions(mod)
```

###

The `estimate` column gives the predicted probability of survival for each passenger, expressed as a number between 0 and 1.

### Exercise 3

Update your working chunk to build a one-row-per-class grid using `datagrid()`, fixing `Age = 30` and `SexCode = 0` (a 30-year-old man) and letting `PClass` vary. Feed the grid to `predictions()`. This is a **conditional** prediction: one hypothetical passenger evaluated in each class. Render.

```
predictions(mod,
  newdata = datagrid(Age = 30, SexCode = 0, PClass = unique))
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r predictions-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r predictions-3-test}
#| echo: true
predictions(mod,
  newdata = datagrid(Age = 30, SexCode = 0, PClass = unique))
```

###

###

This is a **conditional** prediction — we ran the machine on imaginary passengers we constructed, holding other variables at chosen values.

### Exercise 4

Update your working chunk to use `avg_predictions()` with `by = "PClass"`. This is a **marginal** prediction: it averages the model-based prediction for every actual passenger within each class. Render.

```
avg_predictions(mod, by = "PClass")
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r predictions-4}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

```{r predictions-4-test}
#| echo: true
avg_predictions(mod, by = "PClass")
```

###

###

This is a **marginal** prediction — we ran the machine on every real passenger, then averaged the results within each class.

### Exercise 5

Add a new code chunk that prints both the conditional prediction for a 30-year-old man in 2nd class and the marginal prediction for 2nd class, side by side. Render.

```
predictions(mod,
  newdata = datagrid(Age = 30, SexCode = 0, PClass = "2nd"))
avg_predictions(mod, by = "PClass")
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r predictions-5}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

```{r predictions-5-test}
#| echo: true
predictions(mod,
  newdata = datagrid(Age = 30, SexCode = 0, PClass = "2nd"))
avg_predictions(mod, by = "PClass")
```

###

These two 2nd-class numbers are very different despite coming from the same model. The conditional number describes one hypothetical man, age 30. The marginal number averages over all actual 2nd class passengers — men and women included.

###

The same model, two different questions. The conditional number describes one hypothetical man, age 30. The marginal number averages over all actual 2nd class passengers — men and women included.

### Exercise 6

Neither answer is wrong. They answer different questions — and knowing which question you are asking is exactly the thinking `marginaleffects` demands.

```
plot_predictions(mod, condition = c("Age", "PClass")) +
  geom_dots(
    alpha = .8, scale = .3, pch = 18,
    data = titanic,
    aes(x = Age, y = Survived,
        side = ifelse(Survived == 1, "bottom", "top"))
  ) +
  labs(
    title = "Predicted survival probability on the Titanic",
    x = "Age", y = "Predicted probability of survival",
    color = "Ticket class", fill = "Ticket class"
  )
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r predictions-6}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 12)
```

###

```{r predictions-6-test}
#| echo: true
plot_predictions(mod, condition = c("Age", "PClass")) +
  geom_dots(
    alpha = .8, scale = .3, pch = 18,
    data = titanic,
    aes(x = Age, y = Survived,
        side = ifelse(Survived == 1, "bottom", "top"))
  ) +
  labs(
    title = "Predicted survival probability on the Titanic",
    x = "Age", y = "Predicted probability of survival",
    color = "Ticket class", fill = "Ticket class"
  )
```

###

Every point on these curves is a conditional prediction. The dots show what actually happened — where the curves align with the dots, the model fits well.

### Exercise 7

Commit and push. Ask AI for a commit message. In the bash Terminal, run:

```
git log -1
```

CP/CR.

```{r predictions-7}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

```
commit 7d4e9b2
Author: Var Kurapati <vkurapat@example.com>
Date:   Mon Jul 14 14:51:03 2025 -0700

    Add Titanic survival prediction plot
```

###

Your first plot shows predicted probabilities. Next, we explore counterfactual comparisons — how much does predicted survival change when we manipulate a specific variable?

## Comparisons
###

A **counterfactual comparison** is a function of two or more model-based predictions made with different predictor values. Instead of asking "what is the predicted survival probability?", we ask "how much does it change if we change sex from male to female?"

### Exercise 1

In a new code chunk, use `comparisons()` to estimate the **conditional** change in predicted survival associated with being female vs. male, for a specific passenger: 30 years old, in 2nd class. Render.

```
grid <- data.frame(Age = 30, PClass = "2nd", SexCode = 1)
comparisons(mod, variables = "SexCode", newdata = grid)
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r comparisons-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r comparisons-1-test}
#| echo: true
grid <- data.frame(Age = 30, PClass = "2nd", SexCode = 1)
comparisons(mod, variables = "SexCode", newdata = grid)
```

###

The estimate tells us how much the predicted survival probability changes when `SexCode` shifts from 0 to 1, for this specific 30-year-old 2nd class passenger.

### Exercise 2

Update your working chunk to compute the **marginal** version: the average change in predicted survival associated with being female, broken down by ticket class, using `avg_comparisons()` with `by`. Render.

```
avg_comparisons(mod, variables = "SexCode", by = "PClass")
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r comparisons-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

```{r comparisons-2-test}
#| echo: true
avg_comparisons(mod, variables = "SexCode", by = "PClass")
```

###

The `by` argument averages the effect of being female across all actual passengers in each class. "Women and children first" was enforced policy — and the data confirms it.

### Exercise 3

Add a new code chunk that prints both the conditional comparison for a 30-year-old 2nd class passenger and the marginal comparison for 2nd class, side by side. Render.

```
comparisons(mod,
  variables = "SexCode",
  newdata = data.frame(Age = 30, PClass = "2nd", SexCode = 1))
avg_comparisons(mod, variables = "SexCode", by = "PClass")
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r comparisons-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

```{r comparisons-3-test}
#| echo: true
comparisons(mod,
  variables = "SexCode",
  newdata = data.frame(Age = 30, PClass = "2nd", SexCode = 1))
avg_comparisons(mod, variables = "SexCode", by = "PClass")
```

###

The conditional number fixes age at exactly 30. The marginal number lets age vary across the real distribution of 2nd class passengers and averages the effects.

###

Whenever a conditional and a marginal number disagree, the fix is not to pick "the right one" — it is to ask which question you actually meant to answer.

### Exercise 4

Update your working chunk to visualize the conditional comparison across a continuous range of ages, separately by class. Use `plot_comparisons()` with `condition = c("Age", "PClass")`. Render.

```
plot_comparisons(mod,
  variables = "SexCode",
  condition = c("Age", "PClass")) +
  labs(
    title = "Effect of being female on predicted survival probability",
    x = "Age", y = "Risk difference (female - male)",
    color = "Ticket class", fill = "Ticket class"
  )
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`. CP/CR.

```{r comparisons-4}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8)
```

###

```{r comparisons-4-test}
#| echo: true
plot_comparisons(mod,
  variables = "SexCode",
  condition = c("Age", "PClass")) +
  labs(
    title = "Effect of being female on predicted survival probability",
    x = "Age", y = "Risk difference (female - male)",
    color = "Ticket class", fill = "Ticket class"
  )
```

###

`plot_comparisons()` runs the machine twice for each passenger — once female, once male — subtracts, and plots the gap. This is the counterfactual question made visible: not two groups of different people, but the same passengers imagined in two worlds.

###

The gap can only be interpreted causally if we know how passengers ended up in each group. The picture is only as trustworthy as the mechanism behind it.

### Exercise 5

Commit and push. Ask AI for a commit message. In the bash Terminal, run:

```
git log -1
```

CP/CR.

```{r comparisons-5}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6)
```

###

```
commit 9c2b5f3
Author: Var Kurapati <vkurapat@example.com>
Date:   Mon Jul 14 16:20:11 2025 -0700

    Add counterfactual comparisons section
```

###

You have now computed both predictions and counterfactual comparisons. The final section wires everything together into a published website.

## Summary
###

You have built two sections of analysis. Now wire them together into a navigable Quarto website and publish the result.

### Exercise 1

Create a `_quarto.yml` file at the top level of your repo with the following contents:

```
project:
  type: website

website:
  title: "Titanic Survival"
  navbar:
    left:
      - href: analysis.qmd
        text: Analysis
      - href: about.qmd
        text: About

format:
  html:
    theme: cosmo
```

In the bash Terminal, run:

```
quarto render
ls
```

In the R Terminal, run `show_file("_quarto.yml")`. CP/CR.

```{r summary-1}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 15)
```

###

<pre><code>project:
  type: website

website:
  title: "Titanic Survival"
  navbar:
    left:
      - href: analysis.qmd
        text: Analysis
      - href: about.qmd
        text: About

format:
  html:
    theme: cosmo
</code></pre>

###

`_quarto.yml` turns a folder of `.qmd` files into a navigable website. Without it, each page renders independently.

### Exercise 2

Create `about.qmd` with a YAML title of `"About"`. Add 2-3 sentences explaining who you are and why you investigated Titanic survival using `marginaleffects`. Also add an introductory paragraph to `analysis.qmd` before the plots explaining what the analysis investigates and naming the difference between conditional and marginal quantities as the central idea. Render.

In the R Terminal, run `show_file("about.qmd")`. CP/CR.

```{r summary-2}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 10)
```

###

Your `about.qmd` should have a YAML title and at least two sentences of prose.

###

An About page gives a human face to a data analysis — anyone who finds your published site can learn who made it and why.

### Exercise 3

Publish the website to GitHub Pages. In the bash Terminal, run:

```
quarto publish gh-pages
```

Copy/paste the resulting URL below.

```{r summary-3}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

`quarto publish gh-pages` without a filename publishes the whole website. Both pages are now live, linked by the navigation bar.

### Exercise 4

Commit and push any remaining changes. Copy/paste the URL to your GitHub repo.

```{r summary-4}
question_text(NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3)
```

###

The `marginaleffects` package works with over 100 model types. The same workflow — fit a model, decide whether you want conditional or marginal quantities, compute, visualize — applies to all of them.

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```


