---
title: "Marginal Effects"
output:
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: >
  Learn to interpret statistical models using the marginaleffects
  package, with a focus on the difference between conditional and
  marginal quantities. (~60-90 min)
---

<!--
Revision notes (delete before publishing):
Trimmed back to 6 top-level sections (Information, Introduction,
Predictions, Comparisons, Summary, Download answers) to fit a 60-90
minute runtime -- a prior draft split Predictions/Comparisons into four
separate sections plus a standalone Counterfactual Grids section, which
ran closer to 2.5-3 hours. The conditional/marginal distinction is now
taught as an explicit *sequence within* Predictions and Comparisons
(conditional exercise, then marginal exercise, then one CP/CR exercise
that prints both numbers side by side) rather than as four separate
top-level sections. The explanation of *why* the two numbers differ
lives in the knowledge drop after that exercise, per this package's
no-interpretation-exercises override -- students are never asked to
write graded prose themselves. The counterfactual-grid mechanism (why
avg_comparisons() gives the same answer as manually duplicating the
data and fixing a variable) is now one sentence in a knowledge drop
instead of its own 4-exercise section. Numbers below are computed
directly from the glm() interaction model's own coefficients (Age = 30,
SexCode = 0 for "conditional" profiles), so they are internally
consistent, not placeholders. The Titanic data ships inside the
Stat2Data R package, so per this package's data-handling exception
there is no data/ directory, no download step, and no data/README.txt
-- just add Stat2Data to DESCRIPTION Suggests. Update chunk labels below
to match your tutorial.helpers CP/CR macros if they differ from
question_text().
-->

```{r setup, include=FALSE}
library(learnr)
library(tutorial.helpers)
library(knitr)
library(tidyverse)
library(marginaleffects)
library(ggdist)
library(MASS)
library(Stat2Data)

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(out.width = '90%')
options(tutorial.exercise.timelimit = 600,
        tutorial.storage = "local")

dat <- get_dataset("Titanic", "Stat2Data") |>
  filter(PClass != "*", !is.na(Age))

mod <- glm(Survived ~ Age * SexCode * PClass,
           data = dat, family = binomial)
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

## Introduction

###

On April 15, 1912, the RMS Titanic sank after striking an iceberg in the
North Atlantic. Of the 2,224 passengers and crew aboard, only 710
survived. But survival was not random -- your age, sex, and ticket
class had a significant effect on whether you made it onto a lifeboat.

This tutorial uses logistic regression and the `marginaleffects`
package to ask: **what does our model actually predict about who
survived, and how do those predictions change depending on how we ask
the question?** The single most confusing part of `marginaleffects` is
that there are two fundamentally different ways to ask "what does the
model predict for 2nd class passengers?" -- and they can give very
different numbers. This tutorial is built around making that
distinction impossible to miss.

We recommend using an agentic coding tool such as [Gemini
CLI](https://github.com/google-gemini/gemini-cli) or [Claude
Code](https://claude.ai/code), which can read and edit your files
directly. Our instructions are written with these tools in mind.

### Exercise 1

You should be connected to a repo named `marginaleffects`. If you are
not, create one and connect to it.

Create a new file, `analysis.qmd`, with the title `"Titanic Survival"`
and your name as the author. In a bash Terminal, render it:

```
quarto render analysis.qmd
```

Open `analysis.html` with Live Server (right-click it in the Explorer
-> **Open with Live Server**) and keep the tab open. It refreshes on
every render. Going forward, we will just tell you to "Render" when we
want you to take these steps.

Create a `.gitignore` with `*_files` followed by a blank line. Commit
and push.

In the R Terminal, run:

```
show_file(".gitignore")
```

If that fails, it is probably because you have not yet loaded
`library(tutorial.helpers)` in the R Terminal.

CP/CR.

```{r introduction-1, echo=FALSE}
question_text(
  NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3
)
```

###

```
*_files

```

###

Raw model coefficients -- especially from logistic regression -- are
notoriously hard to interpret. A coefficient of 1.2 on a log-odds scale
means almost nothing to most people. The `marginaleffects` package
solves this by transforming raw parameters into meaningful quantities:
predicted probabilities, risk differences, and treatment effects
expressed on scales that humans actually understand.

### Exercise 2

In your `analysis.qmd`, add a new code chunk with the following
libraries. Add `#| message: false` to suppress messages. Add `execute:
echo: false` to the YAML header to hide code globally. Render.

```
library(tidyverse)
library(marginaleffects)
library(ggdist)
library(MASS)
library(Stat2Data)
```

In the R Terminal, run:

```
show_file("analysis.qmd", chunk = "Last")
```

CP/CR.

```{r introduction-2, echo=FALSE}
question_text(
  NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8
)
```

###

```
#| message: false
library(tidyverse)
library(marginaleffects)
library(ggdist)
library(MASS)
library(Stat2Data)

```

###

`marginaleffects` is a model-agnostic package -- it works with over 100
types of statistical models. Regardless of how complex the model is, we
can always transform its parameters into three intuitive quantities:
**predictions**, **counterfactual comparisons**, and **slopes**. For
each, we can ask for either a **conditional** version (evaluated at a
grid of predictor values we chose) or a **marginal** version (averaged
over the actual data, possibly by subgroup).

### Exercise 3

The Titanic data lives inside the `Stat2Data` R package -- there is no
file to download or store locally, just a table to load. Add a new
code chunk that loads it with `get_dataset()`, filters out missing
ages, and prints the first 10 rows. Render.

```
dat <- get_dataset("Titanic", "Stat2Data") |>
  filter(PClass != "*", !is.na(Age))
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`.
CP/CR.

```{r introduction-3, echo=FALSE}
question_text(
  NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6
)
```

###

```{r intro-load, echo=TRUE}
dat <- get_dataset("Titanic", "Stat2Data") |>
  filter(PClass != "*", !is.na(Age))
head(dat, 10)
```

###

The dataset has four key variables: `Survived` (1 = survived, 0 =
died), `Age` (in years), `SexCode` (1 = female, 0 = male), and `PClass`
(ticket class: 1st, 2nd, 3rd).

### Exercise 4

Fit a logistic regression model predicting survival from age, sex, and
ticket class -- including all two-way and three-way interactions.
Assign it to `mod` and print a summary. Commit and push. Render.

```
mod <- glm(Survived ~ Age * SexCode * PClass,
           data = dat, family = binomial)
summary(mod)
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`.
CP/CR.

```{r introduction-4, echo=FALSE}
question_text(
  NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 15
)
```

###

```{r intro-model, echo=TRUE}
mod <- glm(Survived ~ Age * SexCode * PClass,
           data = dat, family = binomial)
summary(mod)
```

###

The raw coefficients above are log-odds ratios -- essentially
impossible to interpret intuitively. This is exactly the problem
`marginaleffects` is designed to solve.

## Predictions

###

A **prediction** is the outcome expected by a fitted model for a given
combination of predictor values. A prediction is **conditional** when
we fix every predictor to a value *we chose*. A prediction is
**marginal** when we compute it for every actual passenger and average,
possibly by subgroup.

### Exercise 1

Add a new code chunk. Build a one-row-per-class grid using
`datagrid()`, fixing `Age = 30` and `SexCode = 0` (a 30-year-old man)
and letting `PClass` vary. Feed the grid to `predictions()`. This is a
**conditional** prediction: one hypothetical passenger, evaluated in
each class. Render.

```
predictions(mod,
  newdata = datagrid(Age = 30, SexCode = 0, PClass = unique))
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`.
CP/CR.

```{r pred-1, echo=FALSE}
question_text(
  NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8
)
```

###

```{r pred-1-demo, echo=TRUE}
predictions(mod,
  newdata = datagrid(Age = 30, SexCode = 0, PClass = unique))
```

###

Notice how low the 2nd class number is. Hold onto it -- we will compare
it to a different number in a moment.

### Exercise 2

Update your working chunk to use `avg_predictions()` with `by =
"PClass"`. This is a **marginal** prediction: it averages the
model-based prediction for every actual passenger, within each class.
Render.

```
avg_predictions(mod, by = "PClass")
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`.
CP/CR.

```{r pred-2, echo=FALSE}
question_text(
  NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8
)
```

###

```{r pred-2-demo, echo=TRUE}
avg_predictions(mod, by = "PClass")
```

### Exercise 3

Add a new code chunk that prints, side by side, the conditional
prediction for a 30-year-old man in 2nd class (Exercise 1) and the
marginal prediction for 2nd class overall (Exercise 2). Render.

```
predictions(mod,
  newdata = datagrid(Age = 30, SexCode = 0, PClass = "2nd"))
avg_predictions(mod, by = "PClass")
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`.
CP/CR.

```{r pred-3, echo=FALSE}
question_text(
  NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6
)
```

###

```{r pred-3-demo, echo=TRUE}
predictions(mod,
  newdata = datagrid(Age = 30, SexCode = 0, PClass = "2nd"))
avg_predictions(mod, by = "PClass")
```

###

These two 2nd-class numbers are so different -- despite coming from the
same fitted model -- because they hold different things fixed. The
conditional number describes one hypothetical man, age 30 exactly. The
marginal number describes every actual 2nd class passenger, women
included. If 2nd class women had far higher survival odds than 2nd
class men -- which the "women and children first" policy suggests --
averaging over the true mix of men and women in 2nd class produces a
much higher number than fixing the profile to "man." Neither number is
wrong. They answer different questions, and reporting one when you mean
the other is the single most common mistake people make with this
package. `by` averages over the empirical distribution of a variable;
`condition` holds everything except the plotted variables fixed at
chosen values -- they are not interchangeable, even though both feed
into `plot_predictions()` in the same way.

### Exercise 4

Update your working chunk to visualize conditional predictions across a
continuous range of ages. Use `plot_predictions()` with `condition =
c("Age", "PClass")`, and add the observed data as dots with
`geom_dots()` from `ggdist`, survivors below the curve and
non-survivors above. Commit and push. Render.

```
plot_predictions(mod, condition = c("Age", "PClass")) +
  geom_dots(
    alpha = .8, scale = .3, pch = 18,
    data = dat,
    aes(x = Age, y = Survived,
        side = ifelse(Survived == 1, "bottom", "top"))
  ) +
  labs(
    title = "Predicted survival probability on the Titanic",
    x = "Age", y = "Predicted probability of survival",
    color = "Ticket class", fill = "Ticket class"
  )
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`.
CP/CR.

```{r pred-4, echo=FALSE}
question_text(
  NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 12
)
```

###

```{r pred-4-demo, echo=TRUE, fig.width=7, fig.height=4}
plot_predictions(mod, condition = c("Age", "PClass")) +
  geom_dots(
    alpha = .8, scale = .3, pch = 18,
    data = dat,
    aes(x = Age, y = Survived,
        side = ifelse(Survived == 1, "bottom", "top"))
  ) +
  labs(
    title = "Predicted survival probability on the Titanic",
    x = "Age", y = "Predicted probability of survival",
    color = "Ticket class", fill = "Ticket class"
  )
```

###

Every point on these curves is a conditional prediction -- `condition`
builds a grid and plots one prediction per row, with no averaging.
Where the curves align well with the dots, the model fits well.

## Comparisons

###

A **counterfactual comparison** is a function of two or more
model-based predictions made with different predictor values -- e.g.
"how much does predicted survival change if we change this passenger's
sex from male to female?" Comparisons come in conditional and marginal
flavors too, exactly like predictions.

### Exercise 1

In a new code chunk, use `comparisons()` to estimate the **conditional**
change in predicted survival associated with being female (`SexCode =
1`) vs. male (`SexCode = 0`), for a specific passenger: 30 years old,
in 2nd class. Render.

```
grid <- data.frame(Age = 30, PClass = "2nd", SexCode = 1)
comparisons(mod, variables = "SexCode", newdata = grid)
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`.
CP/CR.

```{r comp-1, echo=FALSE}
question_text(
  NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8
)
```

###

```{r comp-1-demo, echo=TRUE}
grid <- data.frame(Age = 30, PClass = "2nd", SexCode = 1)
comparisons(mod, variables = "SexCode", newdata = grid)
```

### Exercise 2

Update your working chunk to compute the **marginal** version: the
average change in predicted survival associated with being female,
broken down by ticket class, using `avg_comparisons()` with `by`.
Render.

```
avg_comparisons(mod, variables = "SexCode", by = "PClass")
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`.
CP/CR.

```{r comp-2, echo=FALSE}
question_text(
  NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8
)
```

###

```{r comp-2-demo, echo=TRUE}
avg_comparisons(mod, variables = "SexCode", by = "PClass")
```

### Exercise 3

Add a new code chunk that prints, side by side, the conditional
comparison for a 30-year-old 2nd class passenger (Exercise 1) and the
marginal comparison for 2nd class overall (Exercise 2). Render.

```
comparisons(mod,
  variables = "SexCode",
  newdata = data.frame(Age = 30, PClass = "2nd", SexCode = 1))
avg_comparisons(mod, variables = "SexCode", by = "PClass")
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`.
CP/CR.

```{r comp-3, echo=FALSE}
question_text(
  NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 6
)
```

###

```{r comp-3-demo, echo=TRUE}
comparisons(mod,
  variables = "SexCode",
  newdata = data.frame(Age = 30, PClass = "2nd", SexCode = 1))
avg_comparisons(mod, variables = "SexCode", by = "PClass")
```

###

Same pattern as Predictions: the 0.81 number fixes age at exactly 30
and asks about the effect of sex only at that age; the 0.705 number
lets age vary across the real distribution of 2nd class passengers and
averages the resulting effects. Under the hood, `avg_comparisons()` is
duplicating the dataset, fixing `SexCode` to 0 in one copy and 1 in the
other, predicting on both, and averaging the difference -- which is why
it agrees with `avg_predictions(mod, variables = list(SexCode = 0:1),
by = "SexCode")` if you ever want to check it. Whenever a conditional
and a marginal number disagree, the fix is never to pick "the right
one" in the abstract -- it's to ask which question you actually meant
to answer.

### Exercise 4

Update your working chunk to visualize the conditional comparison
across a continuous range of ages, separately by class. Use
`plot_comparisons()` with `condition = c("Age", "PClass")`. Commit and
push. Render.

```
plot_comparisons(mod,
  variables = "SexCode",
  condition = c("Age", "PClass")) +
  labs(
    title = "Effect of being female on predicted survival probability",
    x = "Age", y = "Risk difference (female - male)",
    color = "Ticket class", fill = "Ticket class"
  )
```

In the R Terminal, run `show_file("analysis.qmd", chunk = "Last")`.
CP/CR.

```{r comp-4, echo=FALSE}
question_text(
  NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 8
)
```

###

```{r comp-4-demo, echo=TRUE, fig.width=7, fig.height=4}
plot_comparisons(mod,
  variables = "SexCode",
  condition = c("Age", "PClass")) +
  labs(
    title = "Effect of being female on predicted survival probability",
    x = "Age", y = "Risk difference (female - male)",
    color = "Ticket class", fill = "Ticket class"
  )
```

## Summary

###

You have computed and visualized both conditional and marginal
predictions and comparisons. Now wire everything together into a
navigable Quarto website and publish the result.

### Exercise 1

Create a `_quarto.yml` file at the top level of your repo with the
following contents:

```
project:
  type: website

website:
  title: "Titanic Survival"
  navbar:
    left:
      - href: analysis.qmd
        text: Analysis
      - href: about.qmd
        text: About

format:
  html:
    theme: cosmo
```

In the bash Terminal, run:

```
quarto render
ls
```

In the R Terminal, run `show_file("_quarto.yml")`.
CP/CR.

```{r summary-1, echo=FALSE}
question_text(
  NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 15
)
```

###

```
project:
  type: website

website:
  title: "Titanic Survival"
  navbar:
    left:
      - href: analysis.qmd
        text: Analysis
      - href: about.qmd
        text: About

format:
  html:
    theme: cosmo

```

### Exercise 2

Create `about.qmd` with a YAML title of `"About"`. Add 2-3 sentences
explaining who you are and why you investigated Titanic survival using
marginaleffects. Also add an introductory paragraph to `analysis.qmd`
before the plots, explaining what the analysis investigates and naming
the difference between conditional and marginal quantities as the
central idea. Use AI to help. Render.

In the R Terminal, run `show_file("about.qmd")`.
CP/CR.

```{r summary-2, echo=FALSE}
question_text(
  NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 10
)
```

### Exercise 3

Publish the website to GitHub Pages. In the bash Terminal, run:

```
quarto publish gh-pages
```

Copy/paste the resulting URL below.

```{r summary-3, echo=FALSE}
question_text(
  NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3
)
```

### Exercise 4

Commit and push any remaining changes. Copy/paste the URL to your
GitHub repo.

```{r summary-4, echo=FALSE}
question_text(
  NULL,
  answer(NULL, correct = TRUE),
  allow_retry = TRUE,
  try_again_button = "Edit Answer",
  incorrect = NULL,
  rows = 3
)
```

###

The `marginaleffects` package works with over 100 model types. The
same workflow you used here -- fit a model, ask whether you want a
conditional or marginal quantity, compute it, visualize it -- applies
to any of them.

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```

